---
title: "Data"
subtitle: "Rural Observatory Surveys, georeferencing, and variable construction"
---

```{r}
#| label: setup
source("R/load_data.R")
source("R/plot_themes.R")
source("R/gsynth_helpers.R")
```

This chapter summarises the data construction pipeline. The full reproducible workflow — georeferencing, PA overlap analysis, and variable harmonisation — is documented in [Appendix A](A1_data_pipeline.qmd). The underlying survey data are described in a companion data paper [@ChabeFerret2024]; supplementary materials and replication code are available in the [associated repository](https://doi.org/10.1038/s41597-024-03879-9).

# Survey data {#sec-survey-data}

The Rural Observatory System (ROS) surveyed rural households across Madagascar from 1995 to 2015. We use {{< var n_observatories >}} observatories comprising approximately {{< var n_household_years >}} household-year observations. Each observatory tracks ~500 households per year in a rotating panel design, with random replacement of attrited households preserving cross-sectional representativeness.

```{r}
#| label: fig-obs-map
#| fig-cap: "Location of the 11 rural observatories used in this study."
#| fig-height: 7
#| fig-width: 6

ggplot() +
  geom_sf(data = madagascar, fill = "grey95", color = "grey60", linewidth = 0.3) +
  geom_point(
    data = obs_locations,
    aes(x = longitude, y = latitude, color = role, shape = role),
    size = 3.5, stroke = 1
  ) +
  geom_label_repel(
    data = obs_locations,
    aes(x = longitude, y = latitude, label = paste0(name, " (", code, ")")),
    size = 2.5, fontface = "bold", box.padding = 0.5,
    min.segment.length = 0, max.overlaps = 20
  ) +
  scale_color_manual(
    values = c("Marovoay (strict, 2003)" = "#D32F2F",
               "Alaotra (multipurpose, 2008)" = "#1565C0",
               "New treated (2006\u20132007)" = "#FF8F00",
               "Clean donor pool" = "#4CAF50"),
    name = NULL
  ) +
  scale_shape_manual(
    values = c("Marovoay (strict, 2003)" = 17,
               "Alaotra (multipurpose, 2008)" = 17,
               "New treated (2006\u20132007)" = 18,
               "Clean donor pool" = 16),
    name = NULL
  ) +
  labs(title = "Rural Observatories in the Analysis") +
  theme_void(base_size = 11) +
  theme(legend.position = "bottom")
```

```{r}
#| label: fig-time-grid
#| fig-cap: "Year-by-year data coverage across observatories and survey sites."
#| fig-height: 7
#| fig-width: 10

obs_names_df <- obs_locations |> st_drop_geometry() |> select(code, name)
# obs 12 (Antsohihy) is in ALL_DONOR_OBS but not in obs_locations
obs_names_df <- bind_rows(obs_names_df, tibble(code = "12", name = "Antsohihy"))

coverage <- household_consolidated |>
  filter(obs %in% c("03", "21", ALL_DONOR_OBS)) |>
  mutate(
    unit_label = case_when(
      site_id == "031" ~ "Marovoay: Bepako",
      site_id == "032" ~ "Marovoay: Madiromiongana",
      site_id == "033" ~ "Marovoay: Ampijoroa",
      site_id == "034" ~ "Marovoay: Maroala",
      obs == "21" & site_id == "211" ~ "Alaotra: Avaradrano / Feramanga A.",
      obs == "21" & site_id == "212" ~ "Alaotra: Mangabe",
      obs == "21" & site_id == "213" ~ "Alaotra: Analamiranga area",
      obs == "21" ~ paste0("Alaotra: site ", site_id),
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(unit_label) | obs %in% ALL_DONOR_OBS)

treated_cov <- coverage |>
  filter(!is.na(unit_label)) |>
  distinct(unit_label, year, obs) |>
  mutate(
    status = case_when(
      obs == "03" & grepl("Bepako|Madiromiongana", unit_label) &
        year >= TREAT_YEAR_MARO ~ "Treated (post)",
      obs == "03" & grepl("Bepako|Madiromiongana", unit_label) ~ "Treated (pre)",
      obs == "03" ~ "Maro. control",
      obs == "21" & year >= TREAT_YEAR_ALA ~ "Treated (post)",
      obs == "21" ~ "Treated (pre)"
    ),
    group = if_else(obs == "03", "Marovoay", "Alaotra")
  )

donor_cov <- coverage |>
  filter(obs %in% ALL_DONOR_OBS) |>
  distinct(obs, year) |>
  left_join(obs_names_df, by = c("obs" = "code")) |>
  mutate(
    unit_label = name,
    status = "Donor",
    group = "Donor pool"
  )

all_cov <- bind_rows(treated_cov, donor_cov)

level_order <- c(
  donor_cov |> distinct(unit_label) |> arrange(unit_label) |> pull(unit_label),
  "Alaotra: Analamiranga area",
  "Alaotra: Mangabe",
  "Alaotra: Avaradrano / Feramanga A.",
  "Marovoay: Maroala",
  "Marovoay: Ampijoroa",
  "Marovoay: Bepako",
  "Marovoay: Madiromiongana"
)

all_cov <- all_cov |>
  mutate(unit_label = factor(unit_label, levels = level_order))

ggplot(all_cov, aes(x = factor(year), y = unit_label, fill = status)) +
  geom_tile(color = "white", linewidth = 0.3) +
  scale_fill_manual(
    values = c(
      "Treated (post)" = "#EF9A9A",
      "Treated (pre)"  = "#A5D6A7",
      "Maro. control"  = "#90CAF9",
      "Donor"          = "#E0E0E0"
    ),
    name = NULL
  ) +
  labs(x = NULL, y = NULL, title = "Panel Coverage: Sites \u00d7 Years") +
  theme_minimal(base_size = 10) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    axis.text.y = element_text(size = 8),
    legend.position = "bottom",
    panel.grid = element_blank()
  )
```

# Georeferencing {#sec-georef}

The original survey data lacked systematic georeferencing. Locality names were recorded as free text, with inconsistent spellings across years and languages (Malagasy and French). We linked each observation to official administrative boundaries using the Common Operational Datasets (COD), maintained by OCHA and Madagascar's BNGRC.

The matching procedure is hierarchical:

1. **String normalisation** — lowercase, remove diacritics and qualifiers (*centre*, *haut*, *bas*), standardise bilingual toponyms, convert Roman to Arabic numerals.
2. **Fuzzy matching** — Jaro-Winkler distance against the COD gazetteer, filtered by observatory-level district membership, with a maximum distance threshold of 0.25.
3. **Cascading spatial levels** — match first at the commune (ADM3) level, then refine to fokontany (ADM4) where possible.

The result is a geocoded correspondence table linking each ROS household identifier (`j5`) to its P-coded commune and fokontany. This enables spatial overlay with protected area boundaries.

# Protected area overlaps {#sec-pa-overlaps}

We overlay the geocoded ROS fokontany with the dynamic WDPA layer (see [Appendix B](A2_donor_validity.qmd) for the temporal PA boundary construction). Three temporal configurations emerge:

| Configuration | Observatories | Use |
|---|---|---|
| PA predates ROS surveys (no baseline) | e.g., Ranomafana, Andringitra | Excluded |
| PA created *during* ROS (1999–2014) | Marovoay (2003), Alaotra (2008) | **Primary treatment** |
| PA created *after* ROS ended | Farafangana, Toliara, Fénérive (2006–07) | **Extension treatment** |

: Overview of observatory–PA temporal relationships. {#tbl-pa-timing}

```{r}
#| label: fig-pa-overlaps
#| fig-cap: "Protected areas and rural observatories: spatial and temporal relationships."
#| fig-height: 6
#| fig-width: 9

# --- PA creation events from spatial overlaps ---
study_obs_codes <- as.integer(c("02", "03", "12", "13", "15", "16",
                                "21", "23", "24", "25", "41"))

pa_overlaps <- st_join(
  sel_fokontany |> filter(OBS_CODE %in% study_obs_codes),
  wdpa |> select(NAME, STATUS_YR, IUCN_CAT),
  left = FALSE
) |>
  st_drop_geometry() |>
  mutate(STATUS_YR = as.integer(STATUS_YR)) |>
  filter(STATUS_YR >= 1995, STATUS_YR <= 2015) |>
  # Keep earliest event per PA per observatory
  summarize(
    pa_year = min(STATUS_YR),
    IUCN_CAT = first(IUCN_CAT),
    .by = c(OBS_CODE, NAME)
  )

# Map OBS_CODE to name
obs_name_map <- bind_rows(
  obs_locations |> st_drop_geometry() |>
    mutate(OBS_CODE = as.integer(code)) |>
    select(OBS_CODE, obs_name = name),
  tibble(OBS_CODE = 12L, obs_name = "Antsohihy")
)

pa_overlaps <- pa_overlaps |>
  left_join(obs_name_map, by = "OBS_CODE") |>
  mutate(label = paste0(obs_name, " \u2013 ", NAME),
         type = "PA creation")

# --- Survey year coverage ---
survey_cov <- household_consolidated |>
  filter(obs %in% c("03", "21", ALL_DONOR_OBS)) |>
  distinct(obs, year) |>
  left_join(
    bind_rows(
      obs_locations |> st_drop_geometry() |> select(code, name),
      tibble(code = "12", name = "Antsohihy")
    ),
    by = c("obs" = "code")
  ) |>
  filter(!is.na(name))

# Build combined data: one row per obs, showing survey span + PA events
# Panel A: survey spans by observatory
obs_spans <- survey_cov |>
  summarize(first_yr = min(year), last_yr = max(year), .by = name) |>
  arrange(first_yr)

# Only include observatories that have PA overlaps (plus both treated)
obs_with_pa <- pa_overlaps |> distinct(obs_name) |> pull(obs_name)
obs_of_interest <- unique(c("Marovoay", "Alaotra", obs_with_pa))

# Survey years for observatories of interest
survey_pts <- survey_cov |>
  filter(name %in% obs_of_interest) |>
  mutate(type = "ROS survey")

# PA creation events
pa_pts <- pa_overlaps |>
  filter(obs_name %in% obs_of_interest) |>
  select(obs_name, year = pa_year, NAME, type)

# Combined for plotting — use observatory–PA label for y-axis
# For survey years, create one row per obs-PA pair
pa_labels <- pa_overlaps |>
  filter(obs_name %in% obs_of_interest) |>
  distinct(obs_name, label)

# Expand survey years across PA labels for each observatory
survey_expanded <- survey_pts |>
  inner_join(pa_labels, by = c("name" = "obs_name"),
             relationship = "many-to-many") |>
  select(label, year, type)

pa_events <- pa_pts |>
  inner_join(pa_labels, by = c("obs_name"),
             relationship = "many-to-many") |>
  select(label, year, type)

timeline_data <- bind_rows(survey_expanded, pa_events)

# Order y-axis by PA creation year
label_order <- pa_overlaps |>
  filter(obs_name %in% obs_of_interest) |>
  arrange(pa_year) |>
  pull(label) |>
  unique()

timeline_data <- timeline_data |>
  mutate(label = factor(label, levels = rev(label_order)))

ggplot(timeline_data, aes(x = year, y = label, color = type)) +
  geom_point(aes(size = type, shape = type)) +
  scale_color_manual(
    values = c("ROS survey" = "#1565C0", "PA creation" = "#D32F2F"),
    name = NULL
  ) +
  scale_size_manual(values = c("ROS survey" = 1.5, "PA creation" = 4),
                    guide = "none") +
  scale_shape_manual(values = c("ROS survey" = 16, "PA creation" = 18),
                     guide = "none") +
  scale_x_continuous(breaks = seq(1995, 2015, 5)) +
  labs(title = "Timing of ROS Surveys Relative to PA Creation") +
  theme_minimal(base_size = 11) +
  theme(
    axis.title = element_blank(),
    legend.position = "top",
    panel.grid.minor = element_blank()
  )
```

The two primary pairs — **Marovoay / Ankarafantsika** (strict, IUCN II, 2002) and **Alaotra / Lac Alaotra** (multipurpose, IUCN V, 2007) — offer the longest pre- and post-treatment windows. Three additional pairs provide shorter but informative extensions ([Chapter 5](05_extensions.qmd)).

# Outcome variable {#sec-outcome}

The primary outcome is **log equivalised household income**. Construction proceeds as follows:

1. **Total income** (`revtot`) — the sum of current income components: rice revenue (`rev_riz`), other crop revenue (`rev_cu`), livestock revenue (`revel`), fishing revenue (`revpeche`), principal activity income (`revppal`), and secondary activity income (`revsec`). Missing components are set to zero; the total is imputed from available sub-components when `revtot` itself is missing.

2. **Equivalisation** — using the OECD-modified scale: $1 + 0.5 \times (\text{additional adults}) + 0.3 \times (\text{children under 14})$, computed from individual-level demographic records.

3. **Winsorisation** — at the 1st and 99th percentiles within each year to limit the influence of extreme values.

4. **Log transformation** — $y_{it} = \log(\text{equivalised income}_{it} + 1)$.

Cross-year harmonisation is non-trivial: variable names, questionnaire modules, and income definitions changed across survey waves. The full details of these adjustments are in [Appendix A](A1_data_pipeline.qmd).

```{r}
#| label: fig-income-trends
#| fig-cap: "Mean log equivalised household income over time, by observatory."
#| fig-height: 5
#| fig-width: 12

maro_panel <- household_consolidated |>
  filter(obs == "03") |>
  mutate(
    treated  = site_id %in% c("031", "032"),
    revtot_eq = revtot / oecd_equiv,
    revtot_w  = winsorize(revtot_eq),
    ln_rev    = log(pmax(revtot_w, 1)),
    .by = year
  )

maro_trends <- maro_panel |>
  summarize(
    mean_ln = mean(ln_rev, na.rm = TRUE),
    se_ln   = sd(ln_rev, na.rm = TRUE) / sqrt(n()),
    .by = c(year, treated)
  ) |>
  mutate(group = if_else(treated, "East bank (treated)", "West bank (donor)"))

p1 <- ggplot(maro_trends, aes(year, mean_ln, color = group, shape = group)) +
  annotate("rect", xmin = TREAT_YEAR_MARO, xmax = 2015,
           ymin = -Inf, ymax = Inf, fill = "#FFCDD2", alpha = 0.3) +
  geom_vline(xintercept = TREAT_YEAR_MARO - 0.5, linetype = "dashed",
             color = "grey40") +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = mean_ln - 1.96*se_ln, ymax = mean_ln + 1.96*se_ln),
                width = 0.3, alpha = 0.5) +
  scale_color_manual(values = c("East bank (treated)" = "#D32F2F",
                                "West bank (donor)" = "#1565C0")) +
  labs(title = "Ankarafantsika (strict, 2003)",
       x = "Year", y = "Mean log(equiv. income)",
       color = NULL, shape = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

ala_panel <- household_consolidated |>
  filter(obs == "21") |>
  mutate(
    revtot_eq = revtot / oecd_equiv,
    revtot_w  = winsorize(revtot_eq),
    ln_rev    = log(pmax(revtot_w, 1)),
    .by = year
  )

ala_trends <- ala_panel |>
  summarize(
    mean_ln = mean(ln_rev, na.rm = TRUE),
    se_ln   = sd(ln_rev, na.rm = TRUE) / sqrt(n()),
    .by = year
  )

p2 <- ggplot(ala_trends, aes(year, mean_ln)) +
  annotate("rect", xmin = TREAT_YEAR_ALA, xmax = 2015,
           ymin = -Inf, ymax = Inf, fill = "#BBDEFB", alpha = 0.3) +
  geom_vline(xintercept = TREAT_YEAR_ALA - 0.5, linetype = "dashed",
             color = "grey40") +
  geom_line(linewidth = 0.8, color = "#1565C0") +
  geom_point(size = 2.5, color = "#1565C0") +
  geom_errorbar(aes(ymin = mean_ln - 1.96*se_ln, ymax = mean_ln + 1.96*se_ln),
                width = 0.3, alpha = 0.5, color = "#1565C0") +
  labs(title = "Lac Alaotra (multipurpose, 2008)",
       x = "Year", y = "Mean log(equiv. income)") +
  theme_minimal(base_size = 11)

p1 + p2 +
  plot_annotation(
    title = "Income Trends by Observatory",
    subtitle = "Shaded regions = post-treatment"
  )
```

# Descriptive statistics {#sec-descriptives}

```{r}
#| label: tbl-descriptives
#| tbl-cap: "Summary statistics for the analysis sample."

analysis_sample <- household_consolidated |>
  filter(
    obs %in% c("03", "21", ALL_DONOR_OBS),
    year >= STUDY_START & year <= STUDY_END
  ) |>
  mutate(
    group = case_when(
      obs == "03" & site_id %in% c("031", "032") ~ "Marovoay treated (east)",
      obs == "03" ~ "Marovoay control (west)",
      obs == "21" ~ "Alaotra (treated)",
      TRUE ~ "Donor pool"
    ),
    revtot_eq = revtot / oecd_equiv
  )

analysis_sample |>
  summarize(
    `N household-years` = n(),
    `Mean income (Ar)` = round(mean(revtot, na.rm = TRUE)),
    `Mean equiv. income` = round(mean(revtot_eq, na.rm = TRUE)),
    `Mean HH size` = round(mean(n_membres, na.rm = TRUE), 1),
    `Mean years (max)` = round(mean(max_years_edu, na.rm = TRUE), 1),
    `% literate adult` = round(mean(any_literate_adult, na.rm = TRUE) * 100, 1),
    .by = group
  ) |>
  arrange(match(group, c("Marovoay treated (east)", "Marovoay control (west)",
                         "Alaotra (treated)", "Donor pool"))) |>
  gt() |>
  tab_header(
    title = "Sample Descriptive Statistics",
    subtitle = paste0("Analysis period: ", STUDY_START, "\u2013", STUDY_END)
  ) |>
  fmt_number(columns = c(`Mean income (Ar)`, `Mean equiv. income`),
             use_seps = TRUE, decimals = 0)
```
