---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Overlaps

```{r load_libraries}
library(tidyverse)    # A series of packages for data manipulation
library(haven)
library(sf)           # for spatial data handling
library(tmap)         # for mapping
library(readxl)       # Read data frames to Excel format
library(writexl)      # Write data frames to Excel format
library(gt)           # for nicely formatted tables
library(cowplot)      # to combine plots
library(gtsummary)    # to produce nice summary tables
library(janitor)      # to simply add rowsums
```

## Inventory of ROS data

```{r}
# Select appropriate folder as data source
data_path <- "data/ROS_MDG_microdata_1995-2015_dta_format/"

# Define a function to load and count surveys per observatory for a given year
load_and_count <- function(year, factorize = FALSE) {
  # Define file path
  file_path <- paste0(data_path, year, "/res_deb.dta")
  
  # Load data
  data <- read_dta(file_path)
  
  # Extract label and convert to factors if option
  if (factorize) {
    data <- data %>%
      mutate(across(everything(), as.character),
             across(where(is.labelled), ~ as.character(as_factor(.))))
  }
  
  # Count surveys per observatory
  count_data <- data %>%
    group_by(j0) %>%
    summarise(survey_count = n()) %>%
    ungroup() %>%
    mutate(year = year)  # Add year column
  
  return(count_data)
}

# Generate a list of years
years <- 1995:2015

# Use purrr::map_df to loop through each year and bind results
obs_count <- map_df(years, load_and_count) %>%
  # Remove rows with observatory "7 " and "NA", which are errors
  filter(j0 != 7 & !is.na(j0) & survey_count > 1) %>%
  rename(observatory = j0)

# Read observatory names
observatory_names <- readxl::read_xlsx("data/observatory_names.xlsx") %>%
  select(code, observatory_name = name)

# PAss it to wide.
obs_count <- obs_count %>%
  left_join(observatory_names, by = c("observatory" = "code")) %>%
  group_by(observatory_name, year) %>%
  summarise(survey_count = sum(survey_count))
```


```{r}
obs_communes <- st_read(
  "data/Observatoires_ROS_communes_COD_v3.gpkg",
  quiet = TRUE)
Vahatra <- st_read("data/Vahatra/Shapefiles/AP_Vahatra.shp", 
                   quiet = TRUE) %>%
  st_make_valid() %>%
  mutate(label = "Protected area__")
tmap_mode("view")
obs_communes %>%
  filter(!is.na(OBS_NAME)) %>%
  tm_shape() +
  tm_polygons(col = "OBS_NAME") +
  tm_shape(Vahatra) +
  tm_fill(col = "darkgreen", alpha = 0.3) +
  tm_scale_bar()
```

## Temporal overlaps

We have the following time frame for observatories neighbouring Protected areas:

```{r}
#| label: fig-ror-pa-years
#| fig-cap: "Timing of ROR surveys and neigbouring PA creations"

Vahatra <- Vahatra %>%
  mutate(nom = ifelse(str_starts(nom, "Corridor Forestier Ambositra-Vondrozo"),
                      "Corridor Forestier Ambositra-Vondrozo", nom))

obs_PA <- tibble::tribble(
       ~Observatory_name,                                ~PA_Name,
              "Marovoay",                        "Ankarafantsika",
               "Alaotra",                           "Lac Alaotra",
               "Alaotra",          "Corridor Ankeniheny Zahamena",
           "Menabe-Belo",                       "Menabe Antimena",
         "Ambohimahasoa",                            "Ranomafana",
         "Ambohimahasoa", "Corridor Forestier Ambositra-Vondrozo",
          "Fianarantsoa", "Corridor Forestier Ambositra-Vondrozo",
             "Tanandava", "Complexe Zones Humides Mangoky Ihotry",
          "Toliara Nord",                                 "Mikea",
      "Toliara littoral",                       "Tsimanampesotsa",
              "Tolanaro",                            "Andohahela",
               "Tsivory",                  "Vohidava-Betsimalaho")



# Combine the PA name and observatory name into a single label for faceting
obs_PA <- obs_PA %>%
  mutate(facet_label = paste(PA_Name, "-", Observatory_name))

# Fetch Survey Years and add as a new column in obs_PA
obs_PA1 <- obs_PA %>%
  rowwise() %>%
  mutate(year = Vahatra %>%
      filter(nom == PA_Name) %>%
      pull(an_crtn) %>%
      unique(),
    year_type = "PA creation year",
    dot_size = 3) %>%
  ungroup()

obs_PA2 <- obs_PA %>%
  rowwise() %>%
  mutate(year = list(
    obs_count %>%
      filter(observatory_name == Observatory_name) %>%
      pull(year)),
    year_type = "ROR survey year",
    dot_size = 2) %>%
  ungroup() %>%
  unnest(year)

obs_PA <- bind_rows(obs_PA1, obs_PA2)

# Create the facet plot
ggplot(obs_PA, aes(x = year, y = facet_label)) +
  geom_point(aes(color = year_type, size = dot_size)) +
  scale_color_manual(
    values = c("ROR survey year" = "blue", "PA creation year" = "red"),
    name = "",
    labels = c("ROR survey year", "PA creation year")
  )+
  scale_size(
    range = c(2, 4),
    guide = "none"  # This removes the size scale from the legend
  ) +
  scale_x_continuous(limits = c(1990, 2015)) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "top")
```

As it can be seen here, the Ranomanfana protected area that was created well before the ROR surveys. Two protected areas where created during ROR survey years: Ankarafantsika protected area (Marovoay observatory) and Mikea (Toliara Nord observatory). 8 other protected areas were created after the ROR surveys had stoped: Vohidava, Tsimanampesotsoa, Menabe Antinmena, Lac Alaotra, Corridor Forestier Abositra-Vodronzo, Corridor Ankeniheny Zahamena, Complexe Zones Humides Mangoky Ihotry and Andohahela.

