---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Overlaps between ROS data and PAs

From the georeferencing work implemented in the previous section, we map the sites surveyed by the rural observatories against the protected areas referenced in the SAPM (système des aires protégées de Madagascar).

```{r load_libraries}
library(tidyverse)    # A series of packages for data manipulation
library(haven)
library(sf)           # for spatial data handling
library(tmap)         # for mapping
library(readxl)       # Read data frames to Excel format


selected_communes <- read_rds("output/selected_communes.rds")
selected_fokontany <- read_rds("output/selected_fokontany.rds")
Vahatra <- st_read("data/Vahatra/Shapefiles/AP_Vahatra.shp", 
                   quiet = TRUE) %>%
  st_make_valid()


tmap_mode("view")
tm_shape(selected_communes) + 
  tm_polygons(fill = "OBS_NAME", col = NULL,
              fill.legend = tm_legend_hide(),
          popup.vars = c("Observatory" = "OBS_NAME",
                         "Data collection years"  = "years",
                         "District" = "ADM2_EN",
                         "Region" = "ADM1_EN")) +
  tm_shape(selected_fokontany) + 
  tm_polygons(fill = "label", palette = c("black"), fill_alpha = 0.6, 
              fill.legend = tm_legend_hide(),
          popup.vars = c("Observatory" = "OBS_NAME",
                         "Data collection years"  = "years",
                         "Commune" = "ADM3_EN",
                         "District" = "ADM2_EN",
                         "Region" = "ADM1_EN")) +
  tm_shape(Vahatra) + 
  tm_polygons(fill = "darkgreen", fill_alpha = 0.4,
          popup.vars = c("Year creation" = "an_crtn",
                         "IUCN category" = "cat_icn",
                         "Manager" = "gest_1",
                         "Manager 2" = "gest_2")) +
  tm_scalebar()

```

## Inventory of ROS data

```{r}
# Select appropriate folder as data source
data_path <- "data/ROS_MDG_microdata/"

# Define a function to load and count surveys per observatory for a given year
load_and_count <- function(year, factorize = FALSE) {
  # Define file path
  file_path <- paste0(data_path, year, "/res_deb.dta")
  
  # Load data
  data <- read_dta(file_path)
  
  # Extract label and convert to factors if option
  if (factorize) {
    data <- data %>%
      mutate(across(everything(), as.character),
             across(where(is.labelled), ~ as.character(as_factor(.))))
  }
  
  # Count surveys per observatory
  count_data <- data %>%
    group_by(j0) %>%
    summarise(survey_count = n()) %>%
    ungroup() %>%
    mutate(year = year)  # Add year column
  
  return(count_data)
}

# Generate a list of years
years <- 1995:2015

# Use purrr::map_df to loop through each year and bind results
obs_count <- map_df(years, load_and_count) %>%
  # Remove rows with observatory "7 " and "NA", which are errors
  filter(j0 != 7 & !is.na(j0) & survey_count > 1) %>%
  rename(observatory = j0)

# Read observatory names
observatory_names <- readxl::read_xlsx("data/observatory_names.xlsx") %>%
  select(code, observatory_name = name)

# PAss it to wide.
obs_count <- obs_count %>%
  left_join(observatory_names, by = c("observatory" = "code")) %>%
  group_by(observatory_name, year) %>%
  summarise(survey_count = sum(survey_count))
```

## Temporal overlaps

We have the following time frame for observatories neighbouring Protected areas:

```{r}
#| label: fig-ror-pa-years
#| fig-cap: "Timing of ROR surveys and neigbouring PA creations"

Vahatra <- Vahatra %>%
  mutate(nom = ifelse(str_starts(nom, "Corridor Forestier Ambositra-Vondrozo"),
                      "Corridor Forestier Ambositra-Vondrozo", nom))

obs_PA <- tibble::tribble(
       ~Observatory_name,                                ~PA_Name,
              "Marovoay",                        "Ankarafantsika",
               "Alaotra",                           "Lac Alaotra",
               "Alaotra",          "Corridor Ankeniheny Zahamena",
           "Menabe-Belo",                       "Menabe Antimena",
         "Ambohimahasoa",                            "Ranomafana",
         "Ambohimahasoa", "Corridor Forestier Ambositra-Vondrozo",
          "Fianarantsoa", "Corridor Forestier Ambositra-Vondrozo",
             "Tanandava", "Complexe Zones Humides Mangoky Ihotry",
         "Toliara North",                                 "Mikea",
       "Toliara coastal",                       "Tsimanampesotsa",
              "Tolanaro",                            "Andohahela",
               "Tsivory",                  "Vohidava-Betsimalaho",
         "Toliara North",                           "Ranobe PK32")



# Combine the PA name and observatory name into a single label for faceting
obs_PA <- obs_PA %>%
  mutate(facet_label = paste(PA_Name, "-", Observatory_name))

# Fetch Survey Years and add as a new column in obs_PA
obs_PA1 <- obs_PA %>%
  rowwise() %>%
  mutate(year = Vahatra %>%
      filter(nom == PA_Name) %>%
      pull(an_crtn) %>%
      unique(),
    year_type = "PA creation year",
    dot_size = 3) %>%
  ungroup()

obs_PA2 <- obs_PA %>%
  rowwise() %>%
  mutate(year = list(
    obs_count %>%
      filter(observatory_name == Observatory_name) %>%
      pull(year)),
    year_type = "ROR survey year",
    dot_size = 2) %>%
  ungroup() %>%
  unnest(year)

obs_PA <- bind_rows(obs_PA1, obs_PA2)

# Create the facet plot
ggplot(obs_PA, aes(x = year, y = facet_label)) +
  geom_point(aes(color = year_type, size = dot_size)) +
  scale_color_manual(
    values = c("ROR survey year" = "blue", "PA creation year" = "red"),
    name = "",
    labels = c("PA creation year", "ROR survey year")
  )+
  scale_size(
    range = c(2, 4),
    guide = "none"  # This removes the size scale from the legend
  ) +
  scale_x_continuous(limits = c(1990, 2015)) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "top")
```

As it can be seen here, the Ranomanfana protected area that was created well before the ROR surveys. Two protected areas where created during ROR survey years: Ankarafantsika protected area (Marovoay observatory) and Mikea (Toliara Nord observatory). 8 other protected areas were created after the ROR surveys had stoped: Vohidava, Tsimanampesotsoa, Menabe Antinmena, Lac Alaotra, Corridor Forestier Abositra-Vodronzo, Corridor Ankeniheny Zahamena, Complexe Zones Humides Mangoky Ihotry and Andohahela.


## Check décrets provisoires

On checke sur https://cnlegis.gov.mg avec les noms des AP

Vohadava-Betsimalaho

```{r}
obs_PA1_revu <- obs_PA1 %>%
  bind_rows(
    filter(obs_PA1, str_detect(PA_Name, "Zahamena")) %>%
      mutate(year = 2005),
    filter(obs_PA1, str_detect(PA_Name, "Alaotra")) %>%
      mutate(year = 2007),
    filter(obs_PA1, str_detect(PA_Name, "Mangoky")) %>%
      mutate(year = 2008),
    filter(obs_PA1, str_detect(PA_Name, "Vondrozo")) %>%
      mutate(year = 2008),
    filter(obs_PA1, str_detect(PA_Name, "Antimena")) %>%
      mutate(year = 2006),
    filter(obs_PA1, str_detect(PA_Name, "Mikea")) %>%
      mutate(year = 2007),
    filter(obs_PA1, str_detect(PA_Name, "Ranobe")) %>%
      mutate(year = 2008))

obs_PA_revu <- bind_rows(obs_PA1_revu, obs_PA2) %>%
  filter(!(PA_Name %in% c("Tsimanampesotsa", "Ranomafana")))

# Create the facet plot
ggplot(obs_PA_revu, aes(x = year, y = facet_label)) +
  geom_point(aes(color = year_type, size = dot_size)) +
  scale_color_manual(
    values = c("ROR survey year" = "blue", "PA creation year" = "red"),
    name = "",
    labels = c("PA creation year", "ROR survey year")
  )+
  scale_size(
    range = c(2, 4),
    guide = "none"  # This removes the size scale from the legend
  ) +
  scale_x_continuous(limits = c(1990, 2015)) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "top")
```

On fait une carte avec uniquement ces AP

```{r}

selected_communes2 <- selected_communes %>%
  filter(OBS_NAME %in% obs_PA_revu$Observatory_name)
selected_fokontany2 <- selected_fokontany %>%
  st_filter(selected_communes2, .predicate = st_overlaps)
Vahatra2 <- Vahatra %>%
  filter(nom %in% obs_PA_revu$PA_Name) %>%
  mutate(temp_decree_year = case_when(
    str_detect(nom, "Zahamena") ~ 2005,
    str_detect(nom, "Alaotra") ~ 2007,
    str_detect(nom, "Mangoky") ~ 2008,
    str_detect(nom, "Vondrozo") ~ 2008,
    str_detect(nom, "Antimena") ~ 2006,
    str_detect(nom, "Mikea") ~ 2007,
    str_detect(nom, "Ranobe") ~ 2008,
    .default = NA),
    an_crtn = as.character(an_crtn),
    an_crtn = case_when(
      str_detect(nom, "Ankarafantsika") ~ "1927-2002",
      .default = an_crtn))

Vahatra2_buffer <- Vahatra2 %>%
  st_transform(crs = 29701) %>%
  st_buffer(dist = 10000) %>%
  st_transform(crs = 4326)

# library(wdpar)
# wdpa <- wdpar::wdpa_read("Madagascar")
# write_rds(wdpa, "data/wdpa/wdpa_202406.rds")
wdpa <- read_rds("data/wdpa/wdpa_202406.rds")
   
tmap_mode("view")
tm_shape(selected_communes2) + 
  tm_fill(col = "OBS_NAME", palette = "Set1", title = "Observatory",
          id = "ADM3_EN", legend.show = FALSE,
          popup.vars = c("Observatory" = "OBS_NAME",
                         "Data collection years"  = "years",
                         "District" = "ADM2_EN",
                         "Region" = "ADM1_EN")) +
  tm_shape(selected_fokontany2) + 
  tm_fill(col = "label", palette = c("black"), alpha = 0.6, title = "Fokontany", 
          id = "ADM4_EN", legend.show = FALSE,
          popup.vars = c("Observatory" = "OBS_NAME",
                         "Data collection years"  = "years",
                         "Commune" = "ADM3_EN",
                         "District" = "ADM2_EN",
                         "Region" = "ADM1_EN")) +
  tm_shape(Vahatra2) + 
  tm_fill(col = "darkgreen", alpha = 0,
          popup.vars = c("Year temporal decree" = "temp_decree_year",
                         "Year creation" = "an_crtn",
                         "IUCN category" = "cat_icn",
                         "Manager" = "gest_1",
                         "Manager 2" = "gest_2")) +
  tm_borders(col = "darkgreen") +
  tm_shape(Vahatra2_buffer) +
  tm_borders(col = "lightgreen") +
 # tm_shape(wdpa) + 
 # tm_fill(col = "darkgreen", alpha = 0.4) +
  tm_scale_bar()
```

Selected list by order of priorities: 

- Alaotra - lac Alaotra and Corridor forestier Ankeniheny Zahamena
- Marovoay - Ankarafantsika
- Tolanaro - Andohahela
- Toliara North - Mikea
- Fenerive East - Tamporo
