---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Marovoay-Ankarafantsika

We will now have a closer look at the data from the Marovoay rural observatory, which surveyed a panel of household every year from 1995 to 2014, except in 2009 due to the political crisis that rocked the country on that year. This timeline coincides with the creation of the Ankarafantsika protected area in 2007. We will have a closer link 

```{r}
library(tidyverse)    # A series of packages for data manipulation
library(haven)        # Required for reading STATA files (.dta)
library(labelled)     # To work with labelled data from STATA
library(sf)           # for spatial data handling
library(stringdist)   # for string distance and matching
library(tmap)         # for mapping
library(fuzzyjoin)    # for fuzzy joining
library(readxl)       # Read data frames to Excel format
library(writexl)      # Write data frames to Excel format
library(gt)           # for nicely formatted tables
```

## Marovoay observatory data

### Assessing geocoding completeness

```{r}
observatory_names <- read_xlsx("data/observatory_names.xlsx")
obs_communes <- st_read(
  "data/Observatoires_ROS_communes_COD_v3.gpkg",
  quiet = TRUE) %>%
  left_join(select(observatory_names, name, code), 
            by = c("OBS_NAME" = "name")) %>%
  rename(OBS_CODE = code) %>%
  st_drop_geometry() %>%
  filter(OBS_NAME == "Marovoay")

obs_fknt <- st_read(
  "data/OCHA_BNGRC admin boundaries/mdg_admbnda_adm4_BNGRC_OCHA_20181031.shp",
  quiet = TRUE) %>%
  filter(ADM3_PCODE %in% obs_communes$ADM3_PCODE) %>%
  st_drop_geometry() 

```


Extract the faulty municipalities

```{r}

correspondence_obs <- read_dta("output/correspondence_table3.dta") %>%
  mutate(j0 = as.numeric(j0)) %>%
  left_join(observatory_names, by = c("j0" = "code")) %>%
  rename(OBS_NAME = name) %>%
  relocate(OBS_NAME, .before = j0)

correspondence_marovoay <- correspondence_obs %>%
  filter(OBS_NAME == "Marovoay")

corr_na_marovoay <- correspondence_marovoay %>%
  filter(method_village == "")

write_xlsx(list(missing = corr_na_marovoay,
                all_matched = correspondence_marovoay,
                all_communes = obs_communes,
                all_fokontany = obs_fknt), 
           "output/corr_na_marovoay.xlsx")

correspondence_marovoay %>%
  group_by(method_village) %>%
  summarise(n = sum(obs_count)) %>%
  gt()

```

We see that the village is missing for 521 observations and present in 9,202 observations for this observatory. Let's have a look at the missing observations. 

```{r}
corr_na_marovoay %>%
  select(Municipality = j42, Village = j4, `Years of occurrence` = years,
         `Number of observations` = obs_count) %>%
  gt()
```

