---
title: "Results"
subtitle: "Strict vs. multipurpose conservation impacts on household income"
---

```{r}
#| label: setup
source("R/load_data.R")
source("R/gsynth_helpers.R")
source("R/plot_themes.R")
```

```{r}
#| label: load-results

# Cached gsynth results (paper14 = clean 5-donor pool)
gs_maro     <- read_rds("data/paper14_gsynth_maro_oecd.rds")
gs_ala      <- read_rds("data/paper14_gsynth_ala_oecd.rds")

# Extended panel (1999–2025) — only available with the full donor pool
gs_maro_ext <- read_rds("data/paper12_gsynth_maro_ext_oecd.rds")

# Pre-computed extended panel for trajectory plot
maro_gs_ext <- read_rds("data/maro_gs_ext_panel.rds")
```

# Ankarafantsika: strict conservation (IUCN II) {#sec-maro}

## Panel construction

The Marovoay model comprises 2 exposed sites (Bepako and Madiromiongana, east bank) and donor sites from the 5 clean observatories identified in [Appendix B](A2_donor_validity.qmd) (Antsirabe, Tsiroanomandidy, Ambovombe, Mahanoro, Itasy). The exposure period begins in 2003, providing 4 pre-exposure years (1999–2002) and up to 12 post-exposure years (2003–2014).

## gsynth ATT path {#sec-maro-att}

```{r}
#| label: fig-maro-att
#| fig-cap: "Ankarafantsika: gsynth ATT path for east-bank sites relative to exposure onset (2003). Outcome: log OECD-equivalised income. Shaded band: 95% parametric bootstrap CI."
#| fig-height: 5
#| fig-width: 8

att_maro_df <- tibble(
  rel_time = as.numeric(rownames(gs_maro$est_att)),
  att      = gs_maro$est_att[, "ATT"],
  ci_lo    = gs_maro$est_att[, "CI.lower"],
  ci_hi    = gs_maro$est_att[, "CI.upper"]
)

ggplot(att_maro_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#D32F2F", alpha = 0.15) +
  geom_line(color = "#D32F2F", linewidth = 0.8) +
  geom_point(color = "#D32F2F", size = 2.5) +
  labs(
    x = "Years relative to exposure (2003)",
    y = "ATT (log equiv. income)",
    title = "Ankarafantsika: gsynth ATT Path",
    subtitle = paste0("Site-level exposure vs ", gs_maro$N_co,
                      " donor sites. r* = ", gs_maro$r_cv, " factor(s).")
  ) +
  theme_minimal(base_size = 12)
```

The gsynth estimator finds a persistent negative effect of PA exposure on equivalised household income at the east-bank Marovoay sites. The average post-exposure ATT is approximately {{< var att_maro >}} log points (95% CI: [{{< var att_maro_ci_lo >}}, {{< var att_maro_ci_hi >}}], *p* = {{< var att_maro_p >}}), corresponding to a roughly {{< var att_maro_pct >}} decline in equivalised income relative to the synthetic counterfactual. The effect emerges shortly after the 2002 extension and does not attenuate over the observation period.

## Unit-specific effects

```{r}
#| label: fig-maro-units
#| fig-cap: "Unit-specific exposure effects for the two east-bank fokontany."
#| fig-height: 5
#| fig-width: 9

eff_mat <- gs_maro$eff
years   <- as.numeric(rownames(eff_mat))

unit_effects <- bind_rows(
  tibble(year = years, effect = eff_mat[, "031"], village = "Bepako (east bank)"),
  tibble(year = years, effect = eff_mat[, "032"], village = "Madiromiongana (east bank)")
) |>
  filter(year >= TREAT_YEAR_MARO)

ggplot(unit_effects, aes(x = factor(year), y = effect, fill = village)) +
  geom_col(position = "dodge") +
  geom_hline(yintercept = 0, color = "grey60") +
  scale_fill_manual(values = c(
    "Bepako (east bank)" = "#fb6a4a",
    "Madiromiongana (east bank)" = "#cb181d"
  )) +
  facet_wrap(~village) +
  labs(
    x = NULL, y = "Exposure effect (log equiv. income)",
    title = "Unit-Specific Post-Exposure Effects"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))
```

Both exposed fokontany show negative effects, though with different magnitudes. Bepako (4.1 km from the park boundary) shows a larger and more consistent effect than Madiromiongana (11.7 km), consistent with a distance gradient in enforcement intensity.

# Lac Alaotra: multipurpose conservation (IUCN VI) {#sec-ala}

## Panel construction

The Alaotra model has 1 exposed unit (the aggregated observatory) and donor sites from 5 clean observatories. The exposure period begins in 2008, providing 9 pre-exposure years (1999–2007) and 7 post-exposure years (2008–2014).

## gsynth ATT path {#sec-ala-att}

```{r}
#| label: fig-ala-att
#| fig-cap: "Lac Alaotra: gsynth ATT path for the aggregated observatory relative to exposure onset (2008). Outcome: log OECD-equivalised income. Shaded band: 95% parametric bootstrap CI."
#| fig-height: 5
#| fig-width: 8

att_ala_df <- tibble(
  rel_time = as.numeric(rownames(gs_ala$est_att)),
  att      = gs_ala$est_att[, "ATT"],
  ci_lo    = gs_ala$est_att[, "CI.lower"],
  ci_hi    = gs_ala$est_att[, "CI.upper"]
)

ggplot(att_ala_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#1565C0", alpha = 0.15) +
  geom_line(color = "#1565C0", linewidth = 0.8) +
  geom_point(color = "#1565C0", size = 2.5) +
  labs(
    x = "Years relative to exposure (2008)",
    y = "ATT (log equiv. income)",
    title = "Lac Alaotra: gsynth ATT Path",
    subtitle = paste0("Observatory-level exposure vs ", gs_ala$N_co,
                      " donor sites. r* = ", gs_ala$r_cv, " factor(s).")
  ) +
  theme_minimal(base_size = 12)
```

The Alaotra gsynth estimates an average post-exposure ATT of approximately {{< var att_ala >}} log points (*p* = {{< var att_ala_p >}}) — close to zero and not statistically significant. The confidence interval comfortably spans zero. Year-by-year estimates are noisy with no clear trend away from the counterfactual.

The null result is consistent with several interpretations: participatory conservation under CBNRM may impose lower costs on aggregate income; the 7-year post-exposure window may be insufficient to detect a gradual effect; and the single-unit design limits statistical power.

# Comparison {#sec-comparison}

```{r}
#| label: tbl-comparison
#| tbl-cap: "Summary of gsynth results: strict (Ankarafantsika) vs multipurpose (Alaotra) conservation."

get_ci <- function(gs, side) {
  if (!is.null(gs$est_avg)) gs$est_avg[1, side] else NA_real_
}
get_se <- function(gs) {
  if (!is.null(gs$est_avg)) gs$est_avg[1, "S.E."] else NA_real_
}
get_pval <- function(gs) {
  if (!is.null(gs$est_avg)) gs$est_avg[1, "p.value"] else NA_real_
}

tibble(
  PA = c("Ankarafantsika PN", "Lac Alaotra PHP"),
  `IUCN Category` = c("II (strict)", "V (multipurpose)"),
  `Exposure year` = c(TREAT_YEAR_MARO, TREAT_YEAR_ALA),
  `Exposed unit(s)` = c("Madiromiongana & Bepako (east bank)",
                        "Aggregated observatory"),
  `ATT (log pts)` = round(c(gs_maro$att_avg, gs_ala$att_avg), 3),
  `SE` = round(c(get_se(gs_maro), get_se(gs_ala)), 3),
  `95% CI` = c(
    paste0("[", round(get_ci(gs_maro, "CI.lower"), 3), ", ",
           round(get_ci(gs_maro, "CI.upper"), 3), "]"),
    paste0("[", round(get_ci(gs_ala, "CI.lower"), 3), ", ",
           round(get_ci(gs_ala, "CI.upper"), 3), "]")
  ),
  `ATT (pct)` = paste0(
    round((exp(c(gs_maro$att_avg, gs_ala$att_avg)) - 1) * 100, 1), "%"),
  `p-value` = round(c(get_pval(gs_maro), get_pval(gs_ala)), 3),
  `r* (CV)` = c(gs_maro$r_cv, gs_ala$r_cv),
  `N donors` = c(gs_maro$N_co, gs_ala$N_co)
) |>
  gt() |>
  tab_header(
    title = "Comparative gsynth Results",
    subtitle = "Separate estimation for each PA (1999–2014)"
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(rows = 1)
  ) |>
  tab_style(
    style = cell_fill(color = "#BBDEFB"),
    locations = cells_body(rows = 2)
  )
```

```{r}
#| label: fig-comparison
#| fig-cap: "Side-by-side ATT paths. Left: Ankarafantsika (strict, from 2003) shows a persistent negative effect. Right: Lac Alaotra (multipurpose, from 2008) shows no significant effect."
#| fig-height: 5
#| fig-width: 12

p_maro_att <- ggplot(att_maro_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#D32F2F", alpha = 0.15) +
  geom_line(color = "#D32F2F", linewidth = 0.8) +
  geom_point(color = "#D32F2F", size = 2) +
  labs(x = "Relative time", y = "ATT (log equiv. income)",
       title = "Ankarafantsika (strict)") +
  theme_minimal(base_size = 11)

p_ala_att <- ggplot(att_ala_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#1565C0", alpha = 0.15) +
  geom_line(color = "#1565C0", linewidth = 0.8) +
  geom_point(color = "#1565C0", size = 2) +
  labs(x = "Relative time", y = "ATT (log equiv. income)",
       title = "Lac Alaotra (multipurpose)") +
  theme_minimal(base_size = 11)

p_maro_att + p_ala_att +
  plot_annotation(title = "ATT Paths: Strict vs Multipurpose Conservation")
```

The contrast is striking. Strict enforcement at Ankarafantsika produces a large, persistent negative effect on household income. Participatory conservation at Lac Alaotra produces no detectable aggregate effect. This divergence is consistent with the different institutional arrangements: exclusionary management restricting access to forest resources versus community-based management integrating local livelihoods.

# Long-run assessment: 2025 resurvey {#sec-longrun}

A 2025 resurvey of Marovoay (1,026 households across the same fokontany) extends the observation window to 22 years post-exposure.

## Extended gsynth {#sec-ext-gsynth}

```{r}
#| label: fig-ext-att
#| fig-cap: "Extended ATT path for Ankarafantsika east-bank sites (1999–2025). The 2025 point, 22 years after the park extension, shows the largest single-period effect."
#| fig-height: 5.5
#| fig-width: 10

att_ext_df <- tibble(
  rel_time = as.numeric(rownames(gs_maro_ext$est_att)),
  att      = gs_maro_ext$est_att[, "ATT"],
  ci_lo    = gs_maro_ext$est_att[, "CI.lower"],
  ci_hi    = gs_maro_ext$est_att[, "CI.upper"]
)

att_2025 <- tail(att_ext_df, 1)

ggplot(att_ext_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#D32F2F", alpha = 0.12) +
  geom_line(color = "#D32F2F", linewidth = 0.8) +
  geom_point(color = "#D32F2F", size = 2.5) +
  geom_point(data = att_2025, color = "#D32F2F", size = 4, shape = 18) +
  annotate("text", x = att_2025$rel_time - 0.3, y = att_2025$att + 0.12,
           label = "2025", fontface = "bold", color = "#D32F2F", hjust = 1,
           size = 3.5) +
  labs(
    x = "Periods relative to exposure (2003 = 0)",
    y = "ATT (log equiv. income)",
    title = "Ankarafantsika: Extended ATT Path (1999–2025)",
    subtitle = paste0(
      "Average post-exposure ATT = ", round(gs_maro_ext$att_avg, 3),
      " (p = ", format.pval(gs_maro_ext$est_avg[1, "p.value"], digits = 3), ")",
      ". 2025 ATT = ", round(att_2025$att, 3)
    )
  ) +
  theme_minimal(base_size = 12)
```

Including 2025 in the panel, the average post-exposure ATT deepens substantially. The 2025-specific ATT is roughly twice the 1999–2014 average, indicating that income losses compound rather than dissipate over time. The widening gap is visible in raw site-level trajectories:

```{r}
#| label: fig-ext-traj
#| fig-cap: |
#|   Site-level income trajectories for the four Marovoay fokontany
#|   (1999–2025). East-bank (exposed, red) and west-bank (control, blue)
#|   tracked together through 2003, then diverge persistently.
#| fig-height: 5.5
#| fig-width: 9

traj_data <- maro_gs_ext |>
  filter(site_id %in% c("031", "032", "033", "034")) |>
  mutate(
    village = case_when(
      site_id == "031" ~ "Bepako (E)",
      site_id == "032" ~ "Madiromiongana (E)",
      site_id == "033" ~ "Ampijoroa (W)",
      site_id == "034" ~ "Maroala (W)"
    ),
    bank = if_else(site_id %in% c("031", "032"),
                   "Exposed (East)", "Control (West)")
  )

ggplot(traj_data, aes(x = year, y = ln_revtot_w,
                      color = village, linetype = bank)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_vline(xintercept = 2003, linetype = "dashed", alpha = 0.5) +
  annotate("text", x = 2003.5,
           y = min(traj_data$ln_revtot_w, na.rm = TRUE) + 0.3,
           label = "PA extension\n(2003)",
           hjust = 0, size = 3, color = "grey40") +
  scale_color_manual(values = c(
    "Ampijoroa (W)" = "#2171b5", "Maroala (W)" = "#6baed6",
    "Madiromiongana (E)" = "#cb181d", "Bepako (E)" = "#fb6a4a"
  )) +
  labs(
    title = "Site-Level Income Trajectories: Marovoay 1999–2025",
    x = NULL, y = "Mean log(equiv. income)",
    color = "Village", linetype = "Exposure group"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

## Robustness of the 2025 gap {#sec-ext-robust}

The east-west income gap in 2025 is not driven by outliers:

- The gap is visible at *every* percentile of the income distribution.
- IQR-based outlier removal *widens* the gap (outliers disproportionately favour the east bank).
- The west-bank advantage spans 5 of 7 major income components.
- Wilcoxon rank-sum $p = 2.1 \times 10^{-10}$; Welch $t$-test $p = 1.8 \times 10^{-5}$.

## Alaotra long-run assessment {#sec-ext-ala}

For Alaotra, a 2025 gsynth extension is not feasible: all Alaotra sites are exposed and no other observatory was resurveyed, so there are no 2025 control observations. The Alaotra estimate remains based on the 1999–2014 panel (ATT ≈ {{< var att_ala >}}, $p$ = {{< var att_ala_p >}}).
