---
title: "Results"
subtitle: "Strict vs. multipurpose conservation impacts on household income"
---

```{r}
#| label: setup
source("R/load_data.R")
source("R/gsynth_helpers.R")
source("R/plot_themes.R")
```

```{r}
#| label: load-results
# Cached gsynth results (paper14 = clean 6-donor pool)
gs_maro <- read_rds("data/paper14_gsynth_maro_oecd.rds")
gs_ala <- read_rds("data/paper14_gsynth_ala_oecd.rds")

# Extended panel (1999–2025) — only available with the full donor pool
gs_maro_ext <- read_rds("data/paper12_gsynth_maro_ext_oecd.rds")

# Pre-computed extended panel for trajectory plot
maro_gs_ext <- read_rds("data/maro_gs_ext_panel.rds")
```

# Ankarafantsika: strict conservation (IUCN II) {#sec-maro}

## Panel construction

The Marovoay model comprises 2 exposed sites (Bepako and Madiromiongana, east bank) and donor sites from the 6 clean observatories identified in [Appendix B](A2_donor_validity.qmd) (Antsirabe, Antsohihy, Tsiroanomandidy, Ambovombe, Mahanoro, Itasy). The exposure period begins in 2003, providing 4 pre-exposure years (1999–2002) and up to 12 post-exposure years (2003–2014).

## gsynth ATT path {#sec-maro-att}

```{r}
#| label: fig-maro-att
#| fig-cap: "Ankarafantsika: gsynth ATT path for east-bank sites relative to exposure onset (2003). Outcome: log OECD-equivalised income. Shaded band: 95% parametric bootstrap CI."
#| fig-height: 5
#| fig-width: 8
att_maro_df <- tibble(
  rel_time = as.numeric(rownames(gs_maro$est_att)),
  att = gs_maro$est_att[, "ATT"],
  ci_lo = gs_maro$est_att[, "CI.lower"],
  ci_hi = gs_maro$est_att[, "CI.upper"]
)

ggplot(att_maro_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#D32F2F", alpha = 0.15) +
  geom_line(color = "#D32F2F", linewidth = 0.8) +
  geom_point(color = "#D32F2F", size = 2.5) +
  labs(
    x = "Years relative to exposure (2003)",
    y = "ATT (log equiv. income)",
    title = "Ankarafantsika: gsynth ATT Path",
    subtitle = paste0(
      "Site-level exposure vs ",
      gs_maro$N_co,
      " donor sites. r* = ",
      gs_maro$r_cv,
      " factor(s)."
    )
  ) +
  theme_minimal(base_size = 12)
```

The generalized synthetic control model finds a persistent negative gap between the east-bank Marovoay sites and their synthetic counterfactual. The average post-exposure average treatment effect (ATT) is approximately {{< var att_maro >}} log points (95% CI: [{{< var att_maro_ci_lo >}}, {{< var att_maro_ci_hi >}}], *p* = {{< var att_maro_p >}}), corresponding to a roughly {{< var att_maro_pct >}} difference relative to the projected counterfactual. The gap opens shortly after the 2002 park extension and does not close over the observation period.

How much of this gap is attributable to the park itself — rather than to the broader economic trajectory of the Marovoay region — is a question that the data alone cannot fully resolve. Marovoay was historically Madagascar's most productive irrigated rice zone, often called the *grenier à blé* of the country. From the 1980s onward, the region experienced a well-documented decline driven by the silting of the Betsiboka River, the degradation of colonial-era irrigation infrastructure, and the collapse of state agricultural support. This regional trajectory predates the 2002 park extension and could plausibly account for part of the divergence that gsynth attributes to conservation exposure, if the donor observatories on the highland plateau or in the southern drylands were not similarly affected by the deterioration of irrigated lowland agriculture.

### Within-Marovoay household DiD check {#sec-maro-rcs}

The household-level repeated cross-section design described in @sec-complementary applies the Callaway–Sant'Anna estimator to the 1999–2014 + 2025 Marovoay panel, contrasting east-bank (Bepako and Madiromiongana) against west-bank fokontany (Ampijoroa and Maroala).

```{r}
#| label: tbl-maro-rcs
#| tbl-cap: "Within-Marovoay household DiD (Callaway–Sant'Anna, panel = FALSE), east versus west bank. Two standard errors are reported: design-based (influence function, no clustering) and site-clustered."
sec_did <- read_rds("data/secondary_did_results.rds")
sec_did$maro$summary |>
  transmute(
    Design = "East bank vs west bank",
    Estimate = estimate,
    `SE (IF)` = se_if,
    `SE (cluster)` = se_cluster,
    `% income` = sprintf("%+.1f%%", pct_income)
  ) |>
  gt() |>
  fmt_number(columns = c(Estimate, `SE (IF)`, `SE (cluster)`), decimals = 3) |>
  tab_header(title = "Marovoay: within-observatory DiD")
```

The within-Marovoay point estimate is close to zero with a tight cluster-robust standard error — strikingly different in magnitude from the gsynth gap of {{< var att_maro >}} log points reported above. This contrast is a key diagnostic, and it cannot simply be dismissed as the two methods answering different questions.

The gsynth compares the east-bank villages against donor observatories located in the highland plateau or the southern drylands — places with different agronomic systems, different exposure to Betsiboka siltation, and different relationships with irrigated rice. If the Marovoay region as a whole has declined relative to those places for reasons unconnected to the park — the collapse of irrigation infrastructure, the retreat of the state from the rice economy, cycles of flooding and soil erosion — then gsynth will attribute that regional divergence to the PA, because the exposure timing coincides with the beginning of the observable panel. The within-Marovoay design controls for this regional shock: because east-bank and west-bank villages share the same river, the same downstream rice economy, and the same local market, a regional shock hits both sides simultaneously and cancels out of the east-west comparison.

The fact that east-bank and west-bank incomes tracked each other so closely throughout the study period is therefore a genuine warning. It suggests that the bulk of the divergence that gsynth identifies is a regional phenomenon — the decline of a once-prosperous agricultural region — rather than an effect that can be confidently attributed to the park boundary. A real park effect may still exist: east-bank communities face the park fence directly, while west-bank communities do not. But the within-observatory evidence places a much tighter bound on the plausible conservation-specific component than the gsynth estimate alone would suggest. The two results should be read together, with neither overruling the other.

## Unit-specific effects

```{r}
#| label: fig-maro-units
#| fig-cap: "Unit-specific exposure effects for the two east-bank fokontany."
#| fig-height: 5
#| fig-width: 9
eff_mat <- gs_maro$eff
years <- as.numeric(rownames(eff_mat))

unit_effects <- bind_rows(
  tibble(
    year = years,
    effect = eff_mat[, "031"],
    village = "Bepako (east bank)"
  ),
  tibble(
    year = years,
    effect = eff_mat[, "032"],
    village = "Madiromiongana (east bank)"
  )
) |>
  filter(year >= TREAT_YEAR_MARO)

ggplot(unit_effects, aes(x = factor(year), y = effect, fill = village)) +
  geom_col(position = "dodge") +
  geom_hline(yintercept = 0, color = "grey60") +
  scale_fill_manual(
    values = c(
      "Bepako (east bank)" = "#fb6a4a",
      "Madiromiongana (east bank)" = "#cb181d"
    )
  ) +
  facet_wrap(~village) +
  labs(
    x = NULL,
    y = "Exposure effect (log equiv. income)",
    title = "Unit-Specific Post-Exposure Effects"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

Both exposed fokontany show negative effects, though with different magnitudes. Bepako (4.1 km from the park boundary) shows a larger and more consistent effect than Madiromiongana (11.7 km), consistent with a distance gradient in enforcement intensity.

# Lac Alaotra: multipurpose conservation (IUCN VI) {#sec-ala}

## Panel construction

The Alaotra model has 1 exposed unit (the aggregated observatory) and donor sites from 6 clean observatories. The exposure period begins in 2008, providing 9 pre-exposure years (1999–2007) and 7 post-exposure years (2008–2014).

## gsynth ATT path {#sec-ala-att}

```{r}
#| label: fig-ala-att
#| fig-cap: "Lac Alaotra: gsynth ATT path for the aggregated observatory relative to exposure onset (2008). Outcome: log OECD-equivalised income. Shaded band: 95% parametric bootstrap CI."
#| fig-height: 5
#| fig-width: 8
att_ala_df <- tibble(
  rel_time = as.numeric(rownames(gs_ala$est_att)),
  att = gs_ala$est_att[, "ATT"],
  ci_lo = gs_ala$est_att[, "CI.lower"],
  ci_hi = gs_ala$est_att[, "CI.upper"]
)

ggplot(att_ala_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#1565C0", alpha = 0.15) +
  geom_line(color = "#1565C0", linewidth = 0.8) +
  geom_point(color = "#1565C0", size = 2.5) +
  labs(
    x = "Years relative to exposure (2008)",
    y = "ATT (log equiv. income)",
    title = "Lac Alaotra: gsynth ATT Path",
    subtitle = paste0(
      "Observatory-level exposure vs ",
      gs_ala$N_co,
      " donor sites. r* = ",
      gs_ala$r_cv,
      " factor(s)."
    )
  ) +
  theme_minimal(base_size = 12)
```

The Alaotra gsynth estimates an average post-exposure ATT of approximately {{< var att_ala >}} log points (95% CI: [{{< var att_ala_ci_lo >}}, {{< var att_ala_ci_hi >}}], *p* = {{< var att_ala_p >}}) — close to zero and not statistically significant. Year-by-year estimates are noisy with no clear trend away from the counterfactual.

The null result is consistent with several interpretations: participatory conservation under CBNRM may impose lower costs on aggregate income; the 7-year post-exposure window may be insufficient to detect a gradual effect; and the single-unit design limits statistical power.

### Within-Alaotra household DiD check {#sec-ala-rcs}

The household-level repeated cross-section design described in @sec-complementary applies the Callaway–Sant'Anna estimator to the 1999–2014 + 2025 Alaotra panel under two within-observatory exposure proxies: distance to the nearest PA boundary (≤ 10 km) and the 2025 share of households declaring PA-dependent extractive uses (≥ 12.5%).

```{r}
#| label: tbl-ala-rcs
#| tbl-cap: "Within-Alaotra household DiD (Callaway–Sant'Anna, panel = FALSE) on the merged seven-fokontany panel, under two exposure proxies. Two standard errors are reported: design-based (influence function, no clustering) and site-clustered."
sec_did_ala <- read_rds("data/secondary_did_results.rds")$ala
sec_did_ala$summary |>
  transmute(
    Design = recode(
      design,
      dist = "Distance ≤ 10 km",
      ext = "Extractive usage ≥ 12.5%"
    ),
    Estimate = estimate,
    `SE (IF)` = se_if,
    `SE (cluster)` = se_cluster,
    `% income` = sprintf("%+.1f%%", pct_income)
  ) |>
  gt() |>
  fmt_number(columns = c(Estimate, `SE (IF)`, `SE (cluster)`), decimals = 3) |>
  tab_header(title = "Alaotra: within-observatory DiD")
```

Both within-Alaotra designs yield small point estimates and wide cluster-robust standard errors that comfortably include zero. Their conclusion agrees with gsynth along a different identifying contrast: not against an outside donor pool, but against the less exposed merged fokontany of the same observatory. The convergence narrows the range of plausible interpretations of the Alaotra null: the absence of an aggregate income effect of the *Paysage Harmonieux Protégé* is robust to whether the counterfactual is built from outside observatories (gsynth) or from within-observatory exposure heterogeneity (within-observatory DiD).

# Comparison {#sec-comparison}

```{r}
#| label: tbl-comparison
#| tbl-cap: "Summary of gsynth results: strict (Ankarafantsika) vs multipurpose (Alaotra) conservation."
get_ci <- function(gs, side) {
  if (!is.null(gs$est_avg)) gs$est_avg[1, side] else NA_real_
}
get_se <- function(gs) {
  if (!is.null(gs$est_avg)) gs$est_avg[1, "S.E."] else NA_real_
}
get_pval <- function(gs) {
  if (!is.null(gs$est_avg)) gs$est_avg[1, "p.value"] else NA_real_
}

tibble(
  PA = c("Ankarafantsika PN", "Lac Alaotra PHP"),
  `IUCN Category` = c("II (strict)", "V (multipurpose)"),
  `Exposure year` = c(TREAT_YEAR_MARO, TREAT_YEAR_ALA),
  `Exposed unit(s)` = c(
    "Madiromiongana & Bepako (east bank)",
    "Aggregated observatory"
  ),
  `ATT (log pts)` = round(c(gs_maro$att_avg, gs_ala$att_avg), 3),
  `SE` = round(c(get_se(gs_maro), get_se(gs_ala)), 3),
  `95% CI` = c(
    paste0(
      "[",
      round(get_ci(gs_maro, "CI.lower"), 3),
      ", ",
      round(get_ci(gs_maro, "CI.upper"), 3),
      "]"
    ),
    paste0(
      "[",
      round(get_ci(gs_ala, "CI.lower"), 3),
      ", ",
      round(get_ci(gs_ala, "CI.upper"), 3),
      "]"
    )
  ),
  `ATT (pct)` = paste0(
    round((exp(c(gs_maro$att_avg, gs_ala$att_avg)) - 1) * 100, 1),
    "%"
  ),
  `p-value` = round(c(get_pval(gs_maro), get_pval(gs_ala)), 3),
  `r* (CV)` = c(gs_maro$r_cv, gs_ala$r_cv),
  `N donors` = c(gs_maro$N_co, gs_ala$N_co)
) |>
  gt() |>
  tab_header(
    title = "Comparative gsynth Results",
    subtitle = "Separate estimation for each PA (1999–2014)"
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(rows = 1)
  ) |>
  tab_style(
    style = cell_fill(color = "#BBDEFB"),
    locations = cells_body(rows = 2)
  )
```

```{r}
#| label: fig-comparison
#| fig-cap: "Overlaid ATT paths. Ankarafantsika (strict, red) shows a persistent negative effect from 2003; Lac Alaotra (multipurpose, blue) shows no significant effect from 2008. Horizontal scale normalized to each PA's exposure onset."
#| fig-height: 5
#| fig-width: 9
bind_rows(
  att_maro_df |> mutate(case = "Ankarafantsika (strict)"),
  att_ala_df  |> mutate(case = "Lac Alaotra (multipurpose)")
) |>
  ggplot(aes(rel_time, att, color = case, fill = case)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.12, color = NA) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c("Ankarafantsika (strict)" = "#D32F2F",
               "Lac Alaotra (multipurpose)" = "#1565C0"),
    name = NULL
  ) +
  scale_fill_manual(
    values = c("Ankarafantsika (strict)" = "#D32F2F",
               "Lac Alaotra (multipurpose)" = "#1565C0"),
    name = NULL
  ) +
  labs(
    x = "Years relative to exposure",
    y = "ATT (log equiv. income)",
    title = "ATT Paths: Strict vs Multipurpose Conservation"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

The contrast between the two estimated ATT paths is sharp on paper: a large persistent negative gap at Ankarafantsika, and a near-zero gap at Alaotra. Before drawing governance conclusions, however, some caution is warranted. At Marovoay, the within-observatory comparison between east and west bank finds no detectable difference, raising the possibility that the gsynth estimate is capturing regional economic decline rather than conservation effects. At Alaotra, the null result may reflect genuinely benign governance or insufficient statistical power. These two caveats temper but do not dissolve the contrast: the evidence is suggestive that strict exclusionary management carries higher livelihood costs than participatory co-management, but the data cannot definitively establish the causal mechanism.

# Long-run assessment: 2025 resurvey {#sec-longrun}

```{r}
#| label: longrun-setup
#| include: false
# 2025 resurvey: Marovoay-only household count
res_deb_25 <- haven::read_dta("data/ROS_MDG_microdata/2025/res_deb.dta")
n_hh_maro_25 <- res_deb_25 |>
  dplyr::filter(obs == "01") |> # Marovoay observatory
  nrow()

# Extended gsynth summary numbers (avoid hard-coding)
att_avg_ext <- gs_maro_ext$att_avg
att_avg_main <- gs_maro$att_avg
att_2025_val <- tail(gs_maro_ext$est_att[, "ATT"], 1)
att_ratio_ext <- att_2025_val / att_avg_main
n_donors_ext <- gs_maro_ext$N_co
```

A 2025 resurvey of Marovoay (`r n_hh_maro_25` households across the four
fokontany: Bepako and Madiromiongana on the east bank, Ampijoroa and Maroala on
the west bank) extends the observation window to 22 years post-exposure.

## Panel construction (extended model)

```{r}
#| label: longrun-donors
#| include: false
donor_cols <- setdiff(colnames(gs_maro_ext$eff), c("031", "032"))
donor_obs <- sort(unique(substr(donor_cols, 1, 2)))
also_exp <- intersect(donor_obs, c("15", "21", "23", "24"))
clean_donor <- intersect(donor_obs, c("02", "12", "13", "16", "25", "41"))
maro_wbank <- intersect(donor_cols, c("033", "034"))
years_2025 <- sum(!is.na(gs_maro_ext$eff["2025", ]))
```

The extended model uses the full donor pool (all ROS sites with a
complete 1999–2014 income series) rather than the clean 6-donor pool
of @sec-maro. The exposed units are unchanged — Bepako (site 031) and
Madiromiongana (site 032), east of the Betsiboka — but the
`r n_donors_ext` donor sites span `r length(donor_obs)` observatories:

- the `r length(clean_donor)` clean donors used in @sec-maro
  (observatories `r paste(clean_donor, collapse = ", ")`);
- the `r length(also_exp)` "also-exposed" observatories flagged in
  [Appendix B](A2_donor_validity.qmd) (`r paste(also_exp, collapse = ", ")`);
- the `r length(maro_wbank)` Marovoay west-bank fokontany (Ampijoroa 033 and
  Maroala 034), which are in the same observatory as the treated units and
  therefore share local shocks.

This is a weaker identification setting than the main specification: three of
the donor observatories were themselves subject to PA expansions during the
period, and two donor sites are one kilometre away from the exposed units. The
results below should therefore be read as descriptive corroboration of the
@sec-maro-att estimates rather than an independent test.

The 2025 resurvey only covered Marovoay (and Alaotra, which is not part of
this panel). Accordingly, in the extended panel, 2025 income data exist for
exactly `r years_2025` sites — the 4 Marovoay fokontany (031, 032, 033, 034) —
and are NA for all other donors. The 2025 counterfactual is therefore
extrapolated from pre-2015 donor dynamics under the gsynth factor model.

## Extended gsynth {#sec-ext-gsynth}

```{r}
#| label: fig-ext-att
#| fig-cap: "Extended ATT path for Ankarafantsika east-bank sites (1999–2025). The 2025 point, 22 years after the park extension, shows the largest single-period effect."
#| fig-height: 5.5
#| fig-width: 10
att_ext_df <- tibble(
  rel_time = as.numeric(rownames(gs_maro_ext$est_att)),
  att = gs_maro_ext$est_att[, "ATT"],
  ci_lo = gs_maro_ext$est_att[, "CI.lower"],
  ci_hi = gs_maro_ext$est_att[, "CI.upper"]
)

att_2025 <- tail(att_ext_df, 1)

ggplot(att_ext_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#D32F2F", alpha = 0.12) +
  geom_line(color = "#D32F2F", linewidth = 0.8) +
  geom_point(color = "#D32F2F", size = 2.5) +
  geom_point(data = att_2025, color = "#D32F2F", size = 4, shape = 18) +
  annotate(
    "text",
    x = att_2025$rel_time - 0.3,
    y = att_2025$att + 0.12,
    label = "2025",
    fontface = "bold",
    color = "#D32F2F",
    hjust = 1,
    size = 3.5
  ) +
  labs(
    x = "Periods relative to exposure (2003 = 0)",
    y = "ATT (log equiv. income)",
    title = "Ankarafantsika: Extended ATT Path (1999–2025)",
    subtitle = paste0(
      "Average post-exposure ATT = ",
      round(gs_maro_ext$att_avg, 3),
      " (p = ",
      format.pval(gs_maro_ext$est_avg[1, "p.value"], digits = 3),
      ")",
      ". 2025 ATT = ",
      round(att_2025$att, 3)
    )
  ) +
  theme_minimal(base_size = 12)
```

Including 2025 in the panel, the average post-exposure ATT deepens
substantially: `r round(att_avg_ext, 3)` log points in the extended model vs
`r round(att_avg_main, 3)` in the 1999–2014 main specification. The
2025-specific ATT (`r round(att_2025_val, 3)` log points) is about
`r sprintf("%.1f", att_ratio_ext)`× the 1999–2014 average, indicating that
income losses compound rather than dissipate over time. The widening gap is
visible in raw site-level trajectories:

```{r}
#| label: fig-ext-traj
#| fig-cap: |
#|   Site-level income trajectories for the four Marovoay fokontany
#|   (1999–2025). East-bank (exposed, red) and west-bank (control, blue)
#|   tracked together through 2003, then diverge persistently.
#| fig-height: 5.5
#| fig-width: 9
traj_data <- maro_gs_ext |>
  filter(site_id %in% c("031", "032", "033", "034")) |>
  mutate(
    village = case_when(
      site_id == "031" ~ "Bepako (E)",
      site_id == "032" ~ "Madiromiongana (E)",
      site_id == "033" ~ "Ampijoroa (W)",
      site_id == "034" ~ "Maroala (W)"
    ),
    bank = if_else(
      site_id %in% c("031", "032"),
      "Exposed (East)",
      "Control (West)"
    )
  )

ggplot(
  traj_data,
  aes(x = year, y = ln_revtot_w, color = village, linetype = bank)
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_vline(xintercept = 2003, linetype = "dashed", alpha = 0.5) +
  annotate(
    "text",
    x = 2003.5,
    y = min(traj_data$ln_revtot_w, na.rm = TRUE) + 0.3,
    label = "PA extension\n(2003)",
    hjust = 0,
    size = 3,
    color = "grey40"
  ) +
  scale_color_manual(
    values = c(
      "Ampijoroa (W)" = "#2171b5",
      "Maroala (W)" = "#6baed6",
      "Madiromiongana (E)" = "#cb181d",
      "Bepako (E)" = "#fb6a4a"
    )
  ) +
  scale_y_continuous(limits = c(0, NA)) +
  guides(
    color = guide_legend(nrow = 1, byrow = TRUE, order = 1),
    linetype = guide_legend(nrow = 1, order = 2)
  ) +
  labs(
    title = "Site-Level Income Trajectories: Marovoay 1999–2025",
    x = NULL,
    y = "Mean log(equiv. income)",
    color = "Village",
    linetype = "Exposure group"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    legend.spacing.y = unit(2, "pt")
  )
```

## Robustness of the 2025 gap {#sec-ext-robust}

The east-west income gap in 2025 is not driven by outliers: it is visible at every percentile of the income distribution, IQR-based outlier removal widens rather than narrows the gap, and the west-bank advantage appears across five of seven major income components. In statistical terms, the gap is large enough to pass standard tests (Wilcoxon rank-sum $p = 2.1 \times 10^{-10}$; Welch $t$-test $p = 1.8 \times 10^{-5}$).

What remains unresolved is whether this gap reflects cumulative park exposure or the long-term effects of the same regional decline discussed in @sec-maro-rcs. By 2025, the Marovoay irrigation system has deteriorated further, rural-to-urban migration has reshaped communities, and the region's rice economy has never recovered its former position. The 2025 gap is real and robust; its attribution to conservation policy specifically requires caution.

## Alaotra long-run assessment {#sec-ala-longrun}

For Alaotra, a 2025 gsynth extension is not feasible: all Alaotra sites are exposed and no other observatory was resurveyed, so there are no 2025 control observations. The Alaotra estimate remains based on the 1999–2014 panel (ATT ≈ {{< var att_ala >}}, $p$ = {{< var att_ala_p >}}).
