---
title: "Marovoay–Ankarafantsika: Extended Econometric Analysis"
author: "BETSAKA project"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    self-contained: true
editor_options:
  chunk_output_type: console
---

# Marovoay–Ankarafantsika: Extended Econometric Analysis

This document extends the analysis in `04_Marovoay-Ankarafantsika.qmd` by implementing
alternative causal inference methods, incorporating the 2025 survey wave, and expanding
the donor pool to the Alaotra observatory sites.
For the full methodological rationale, see `documentation/05_econometric_review.md`.

::: {.callout-warning}
## Preliminary results
All results are provisional. Variable construction, outlier treatment, and model
specification require further validation before any substantive interpretation.
:::

```{r setup}
#| include: false
# Set working directory to project root — must be done before source() calls
proj_root <- normalizePath(file.path(getwd(), ".."))
knitr::opts_knit$set(root.dir = proj_root)
setwd(proj_root)

library(tidyverse)
library(haven)
library(labelled)
library(gt)
library(readxl)
library(panelView)
library(gsynth)
library(fixest)    # TWFE with clustered SEs, event studies
library(did)       # Callaway-Sant'Anna DiD
library(synthdid)  # Arkhangelsky et al. (2021) Synthetic DiD
library(augsynth)  # Ben-Michael, Feller & Rothstein (2021) Augmented SC
library(sf)
library(purrr)
library(scales)

# ── Helper functions ──────────────────────────────────────────────────────────
source("utils/helpers_report.R")   # add_obs(), fix_hameau()
source("utils/sites.R")            # SITES list, assign_site()

# ── Constants ─────────────────────────────────────────────────────────────────
# Marovoay sites
MARO_TREATED  <- c("031", "032")   # Bepako, Madiromiongana (near Ankarafantsika)
MARO_CONTROL  <- c("033", "034")   # Ampijoroa, Maroala
MARO_ALL      <- c(MARO_TREATED, MARO_CONTROL)
TREATED_SITES <- MARO_TREATED      # alias kept for backward compat

# Alaotra sites (j5 prefix "21x")
# 211=Avaradrano, 212=Mangabe, 213=Ambodivoara/Analamiranga/Ambatoharanana,
# 214=Ambatomanga/Ambohidrony, 215=Feramanga Atsimo
ALAOTRA_ALL   <- c("211","212","213","214","215")

# Treatment timing
TREAT_YEAR    <- 2003   # Ankarafantsika reclassification to National Park

# Site labels for plots
site_labels <- c(
  "031" = "031 Bepako (Maro-T)",
  "032" = "032 Madiromiongana (Maro-T)",
  "033" = "033 Ampijoroa (Maro-C)",
  "034" = "034 Maroala (Maro-C)",
  "211" = "211 Avaradrano (Al-C)",
  "212" = "212 Mangabe (Al-C)",
  "213" = "213 Analamiranga etc. (Al-C)",
  "214" = "214 Ambatomanga etc. (Al-C)",
  "215" = "215 Feramanga Atsimo (Al-C)"
)
```

## 1 Panel Construction

### 1.1 Historical household panel (1995–2015)

The historical panel is built from `data/household_consolidated.rds`.

**Note on 2005:** The 2005 Marovoay survey was a tracking survey that covered only a
non-representative subset of households (site 032/Madiromiongana was not surveyed, and
the other sites had atypical composition). **Year 2005 is excluded from all specifications.**

```{r panel_hist}
household_consolidated <- read_rds("data/household_consolidated.rds")

panel_hh_hist <- household_consolidated |>
  mutate(
    obs      = substr(j5, 1, 2),
    site_id  = substr(j5, 1, 3),
    treated_unit = site_id %in% MARO_TREATED,
    post         = year >= TREAT_YEAR,
    treatment    = as.integer(treated_unit & post)
  ) |>
  mutate(
    rev_riz  = coalesce(rev_riz,  0),
    rev_cu   = coalesce(rev_cu,   0),
    revel    = coalesce(revel,    0),
    revpeche = coalesce(revpeche, 0),
    revppal  = coalesce(revppal,  0),
    revsec   = coalesce(revsec,   0),
    revcou   = coalesce(revcou, revppal + revsec + rev_riz + rev_cu + revel + revpeche),
    rev_agri    = rev_riz + rev_cu + revel + revpeche,
    rev_nonagri = revppal + revsec,
    across(c(revtot, revcou, rev_agri, rev_nonagri),
           ~ pmin(.x, quantile(.x, 0.99, na.rm = TRUE)), .names = "{.col}_w"),
    across(ends_with("_w"), ~ log(.x + 1), .names = "ln_{.col}")
  ) |>
  distinct(j5, year, .keep_all = TRUE) |>
  # Exclude atypical 2005 Marovoay tracking survey
  filter(!(obs == "03" & year == 2005))
```

### 1.2 Adding the 2025 wave — Marovoay

Marovoay 2025 households are identified by `j0 == 3` in `res_deb.dta`.
Site assignment uses `fix_hameau(j4)` to map hamlet name strings to the four
canonical site codes (031–034). Income is sourced from `output/ref_incomes_snapshot.rds`
matched on `j5`. Note that `res_v.dta` is the assets module (radio, bicycle…),
**not** income — do not confuse the two.

```{r panel_2025_maro}
res_deb_25 <- read_dta("data/res_deb.dta") |>
  mutate(j5 = as.character(j5), j0 = as.character(j0))

# Income snapshot includes 2025 via compute_income_all_years()
rev_2025 <- read_rds("output/ref_incomes_snapshot.rds") |>
  mutate(j5 = as.character(j5)) |>
  filter(year == 2025)

# Marovoay 2025: j0 == "3", site assigned via fix_hameau(j4)
maro_25_ids <- res_deb_25 |>
  filter(j0 == "3") |>
  mutate(
    site_id = case_when(
      str_starts(str_to_lower(j4), "bepako")          ~ "031",
      str_starts(str_to_lower(j4), "madiromiongana")  ~ "032",
      str_starts(str_to_lower(j4), "ampijoroa")       ~ "033",
      str_starts(str_to_lower(j4), "maroala")         ~ "034",
      TRUE ~ NA_character_
    ),
    obs = "03"
  ) |>
  select(j5, obs, site_id)

panel_maro_25 <- maro_25_ids |>
  left_join(rev_2025, by = "j5") |>
  mutate(
    year = 2025,
    rev_riz  = coalesce(rev_riz,  0),
    rev_cu   = coalesce(rev_cu,   0),
    revel    = coalesce(revel,    0),
    revpeche = coalesce(revpeche, 0),
    revppal  = coalesce(revppal,  0),
    revsec   = coalesce(revsec,   0),
    revcou   = coalesce(revcou, revppal + revsec + rev_riz + rev_cu + revel + revpeche),
    revtot   = coalesce(revtot, revcou),
    rev_agri    = rev_riz + rev_cu + revel + revpeche,
    rev_nonagri = revppal + revsec,
    treated_unit = site_id %in% MARO_TREATED,
    post         = TRUE,
    treatment    = as.integer(treated_unit)
  ) |>
  mutate(
    across(c(revtot, revcou, rev_agri, rev_nonagri),
           ~ pmin(.x, quantile(.x, 0.99, na.rm = TRUE)), .names = "{.col}_w"),
    across(ends_with("_w"), ~ log(.x + 1), .names = "ln_{.col}")
  )
```

### 1.3 Adding the 2025 wave — Alaotra

Alaotra 2025 households are identified by `j0 == 21`. Site codes are the first 3
digits of `j5` (211–215). The same income snapshot is used.

```{r panel_2025_alaotra}
alaotra_25_ids <- res_deb_25 |>
  filter(j0 == "21") |>
  mutate(site_id = substr(j5, 1, 3), obs = "21") |>
  select(j5, obs, site_id)

panel_alaotra_25 <- alaotra_25_ids |>
  left_join(rev_2025, by = "j5") |>
  mutate(
    year = 2025,
    rev_riz  = coalesce(rev_riz,  0),
    rev_cu   = coalesce(rev_cu,   0),
    revel    = coalesce(revel,    0),
    revpeche = coalesce(revpeche, 0),
    revppal  = coalesce(revppal,  0),
    revsec   = coalesce(revsec,   0),
    revcou   = coalesce(revcou, revppal + revsec + rev_riz + rev_cu + revel + revpeche),
    revtot   = coalesce(revtot, revcou),
    rev_agri    = rev_riz + rev_cu + revel + revpeche,
    rev_nonagri = revppal + revsec,
    treated_unit = FALSE,
    post         = TRUE,
    treatment    = 0L
  ) |>
  mutate(
    across(c(revtot, revcou, rev_agri, rev_nonagri),
           ~ pmin(.x, quantile(.x, 0.99, na.rm = TRUE)), .names = "{.col}_w"),
    across(ends_with("_w"), ~ log(.x + 1), .names = "ln_{.col}")
  )
```

### 1.4 Combined panels

Three analytical datasets are constructed:

- **`panel_maro`**: Marovoay only (1998–2014 + 2025, 2005 excluded), HH level  
- **`panel_expanded`**: Marovoay + Alaotra (historical + 2025), HH level  
- **`panel_site_maro`** / **`panel_site_expanded`**: Site-year aggregates for SC estimators

```{r combine_panels}
# ── Marovoay-only panel ───────────────────────────────────────────────────────
panel_maro <- bind_rows(
  panel_hh_hist |> filter(obs == "03"),
  panel_maro_25
) |>
  filter(site_id %in% MARO_ALL, year >= 1998)

# ── Expanded panel (Marovoay + Alaotra) ───────────────────────────────────────
panel_expanded <- bind_rows(
  panel_hh_hist |>
    filter(obs %in% c("03","21")) |>
    mutate(treated_unit = site_id %in% MARO_TREATED,
           post         = year >= TREAT_YEAR,
           treatment    = as.integer(treated_unit & post)),
  panel_maro_25,
  panel_alaotra_25
) |>
  filter(
    (obs == "03" & site_id %in% MARO_ALL     & year >= 1998) |
    (obs == "21" & site_id %in% ALAOTRA_ALL  & year >= 1999)
  )

# ── Site-year aggregation helper ─────────────────────────────────────────────
make_site_panel <- function(df) {
  df |>
    group_by(obs, site_id, year, treated_unit, post, treatment) |>
    summarise(
      ln_revtot_w      = mean(ln_revtot_w,      na.rm = TRUE),
      ln_revcou_w      = mean(ln_revcou_w,       na.rm = TRUE),
      ln_rev_agri_w    = mean(ln_rev_agri_w,     na.rm = TRUE),
      ln_rev_nonagri_w = mean(ln_rev_nonagri_w,  na.rm = TRUE),
      share_agri       = mean(
        if_else(revcou > 0, rev_agri / revcou, NA_real_), na.rm = TRUE),
      n_hh             = n(),
      .groups = "drop"
    )
}

panel_site_maro     <- make_site_panel(panel_maro)
panel_site_expanded <- make_site_panel(panel_expanded)

# Balanced subsets (years where all sites within each panel are observed)
balance_years <- function(site_panel) {
  n_expected <- n_distinct(site_panel$site_id)
  site_panel |>
    group_by(year) |>
    filter(n() == n_expected) |>
    ungroup()
}

panel_site_maro_bal     <- balance_years(panel_site_maro)
panel_site_expanded_bal <- balance_years(panel_site_expanded)
```

## 2 Panel Diagnostics

### 2.1 Site distance to nearest protected area {#sec-dist}

The table below reproduces the output of `distance_sites_pa.R`: distances from
each fokontany centroid to the nearest PA boundary (WDPA polygons).

```{r tbl_distances}
#| tbl-cap: "Distance from fokontany centroids to nearest PA boundary"

sites_dist <- tribble(
  ~Observatory, ~Fokontany,           ~site_id, ~nearest_pa,                   ~dist_nearest_km, ~dist_anka_km,
  "Marovoay",  "Bepako",              "031",    "Ankarafantsika (PN)",           5.25,  5.25,
  "Marovoay",  "Maroala",             "034",    "Ankarafantsika (PN)",           5.61,  5.61,
  "Marovoay",  "Ampijoroa",           "033",    "Andrefana Dry Forests (WHS)",  11.3,  11.3,
  "Marovoay",  "Madiromiongana",      "032",    "Ankarafantsika (PN)",          11.7,  11.7,
  "Alaotra",   "Maritampona",         "213",    "Lac Alaotra (Paysage)",         8.85,  NA,
  "Alaotra",   "Analamiranga",        "213",    "Lac Alaotra (Paysage)",         9.49,  NA,
  "Alaotra",   "Avaradrano",          "211",    "Lac Alaotra (Paysage)",        10.5,   NA,
  "Alaotra",   "Mangabe",             "212",    "Corridor Ankeniheny (RB)",     14.0,   NA,
  "Alaotra",   "Feramanga Atsimo",    "215",    "Lac Alaotra (Paysage)",        14.8,   NA,
  "Alaotra",   "Ambohidrony",         "214",    "Lac Alaotra (Paysage)",        14.8,   NA,
  "Alaotra",   "Ambatomanga",         "214",    "Lac Alaotra (Paysage)",        16.3,   NA
)

sites_dist |>
  gt(groupname_col = "Observatory") |>
  fmt_number(columns = c(dist_nearest_km, dist_anka_km), decimals = 1) |>
  cols_label(
    Fokontany       = "Fokontany",
    site_id         = "Site ID",
    nearest_pa      = "Nearest PA",
    dist_nearest_km = "Dist. nearest PA (km)",
    dist_anka_km    = "Dist. Ankarafantsika (km)"
  ) |>
  tab_style(style = cell_fill(color = "#FFE5E5"),
            locations = cells_body(rows = dist_nearest_km <= 10)) |>
  tab_style(
    style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
    locations = cells_column_labels()
  ) |>
  tab_style(style = cell_text(weight = "bold"), locations = cells_row_groups()) |>
  tab_footnote("Shaded: within 10 km of nearest PA boundary.",
               locations = cells_column_labels(dist_nearest_km)) |>
  tab_source_note(md("Source: `distance_sites_pa.R` using WDPA polygons (excl. Ramsar/UNESCO)."))
```

**Key observations:**

- Bepako (031) and Maroala (034) are the **closest to Ankarafantsika** (5–6 km).
  Maroala is coded as control — its proximity suggests possible spillover contamination.
- Madiromiongana (032, treated) is 11.7 km — similar to Ampijoroa (033, control) at 11.3 km.
  The distance-based treatment classification is therefore imperfect.
- **Alaotra sites** are 9–16 km from Lac Alaotra PA, a different PA designated in 2007
  (post-treatment). Alaotra sites are plausibly unaffected by the 2002 Ankarafantsika
  reclassification, supporting their use as controls.

### 2.2 Coverage heatmap — Marovoay

```{r fig_coverage_maro}
#| fig-cap: "Panel coverage: Marovoay site-year observations (1998–2025, 2005 excluded)"
#| fig-width: 9
#| fig-height: 4

panel_site_maro |>
  mutate(
    site_label = site_labels[site_id],
    Status = case_when(
      treated_unit & post  ~ "Treated (post)",
      treated_unit & !post ~ "Treated (pre)",
      TRUE                  ~ "Control"
    )
  ) |>
  ggplot(aes(x = year, y = fct_rev(site_label), fill = Status)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = n_hh), size = 2.8, color = "black") +
  scale_fill_manual(
    values = c("Treated (pre)" = "#A8D8A8", "Treated (post)" = "#2D6A4F",
               "Control" = "#AED9E0"),
    na.value = "grey90"
  ) +
  scale_x_continuous(breaks = c(seq(1998, 2014, 2), 2025)) +
  geom_vline(xintercept = 2002.5, linetype = "dashed", color = "darkred") +
  annotate("text", x = 2002.5, y = 4.7, label = "PA 2002",
           hjust = -0.1, size = 3, color = "darkred") +
  labs(x = NULL, y = NULL, fill = NULL,
       caption = "2005 excluded (non-representative tracking survey).") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```

### 2.3 Coverage heatmap — Expanded (Marovoay + Alaotra)

```{r fig_coverage_expanded}
#| fig-cap: "Panel coverage: Marovoay + Alaotra site-year observations"
#| fig-width: 11
#| fig-height: 6

panel_site_expanded |>
  mutate(
    site_label = coalesce(site_labels[site_id], site_id),
    Status = case_when(
      treated_unit & post  ~ "Treated (post)",
      treated_unit & !post ~ "Treated (pre)",
      !treated_unit & obs == "03" ~ "Maro control",
      TRUE                 ~ "Alaotra control"
    )
  ) |>
  ggplot(aes(x = year, y = fct_rev(site_label), fill = Status)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = n_hh), size = 2.4, color = "black") +
  scale_fill_manual(
    values = c("Treated (pre)" = "#A8D8A8", "Treated (post)" = "#2D6A4F",
               "Maro control" = "#AED9E0", "Alaotra control" = "#FDD5A8"),
    na.value = "grey90"
  ) +
  scale_x_continuous(breaks = c(seq(1999, 2014, 2), 2025)) +
  geom_vline(xintercept = 2002.5, linetype = "dashed", color = "darkred") +
  labs(x = NULL, y = NULL, fill = NULL,
       caption = "Orange = Alaotra sites (pure controls). 2005 Marovoay excluded.") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```

### 2.4 Raw outcome trajectories

```{r fig_trajectories}
#| fig-cap: "Site-year mean log total income — Marovoay and Alaotra"
#| fig-width: 10
#| fig-height: 5

panel_site_expanded |>
  mutate(Group = case_when(
    treated_unit ~ "Treated (Maro)",
    obs == "03"  ~ "Control (Maro)",
    TRUE         ~ "Control (Alaotra)"
  )) |>
  group_by(Group, year) |>
  summarise(mean_ln = mean(ln_revtot_w, na.rm = TRUE), .groups = "drop") |>
  mutate(in_gap = year > 2015) |>
  ggplot(aes(x = year, y = mean_ln, color = Group, group = Group)) +
  geom_line(data = ~ filter(.x, !in_gap), linewidth = 0.9) +
  geom_line(data = ~ filter(.x, in_gap | year == 2015), linetype = "31", linewidth = 0.9) +
  geom_point(size = 2) +
  geom_vline(xintercept = 2002.5, linetype = "dashed", color = "darkred") +
  annotate("text", x = 2002.5, y = Inf, hjust = -0.1, vjust = 1.8,
           label = "PA 2002", size = 3, color = "darkred") +
  scale_x_continuous(breaks = c(seq(1999, 2014, 2), 2025)) +
  scale_color_manual(values = c(
    "Treated (Maro)"    = "#E41A1C",
    "Control (Maro)"    = "#377EB8",
    "Control (Alaotra)" = "#FF7F00"
  )) +
  labs(x = NULL, y = "Mean log total income (winsorised)", color = NULL,
       caption = "Simple means within group. Dotted: 2015–2025 gap.") +
  theme_minimal()
```

## 3 Method 1: Two-Way Fixed Effects DiD (TWFE)

TWFE is the natural baseline estimator. With a single treatment cohort (2003),
it is not subject to the negative-weighting problem (de Chaisemartin & d'Haultfœuille 2020).

### 3.1 Marovoay-only TWFE (1998–2014, excl. 2005)

Standard errors are clustered at the site level (4 clusters; low-cluster caveat applies —
see @sec-power for wild-bootstrap alternative).

```{r twfe_maro}
panel_twfe <- panel_maro |> filter(year <= 2014)

fit_twfe_revtot  <- feols(ln_revtot_w    ~ treatment | j5 + year,
                          data = panel_twfe, cluster = ~site_id)
fit_twfe_revcou  <- feols(ln_revcou_w    ~ treatment | j5 + year,
                          data = panel_twfe, cluster = ~site_id)
fit_twfe_agri    <- feols(ln_rev_agri_w  ~ treatment | j5 + year,
                          data = panel_twfe, cluster = ~site_id)
fit_twfe_nonagri <- feols(ln_rev_nonagri_w ~ treatment | j5 + year,
                          data = panel_twfe, cluster = ~site_id)

etable(
  fit_twfe_revtot, fit_twfe_revcou, fit_twfe_agri, fit_twfe_nonagri,
  se       = "cluster",
  dict     = c(treatment = "D (treated × post)"),
  headers  = c("log Total", "log Current", "log Agri", "log Non-agri"),
  notes    = "HH and year FEs. Clustered SE at site level (4 clusters). 2005 excluded."
)
```

### 3.2 Expanded TWFE (Marovoay + Alaotra controls)

Adding 5 Alaotra sites as pure controls increases control units from 2 to 7,
which substantially improves inference (9 clusters total).

```{r twfe_expanded}
panel_twfe_exp <- panel_expanded |> filter(year <= 2014)

fit_twfe_exp_revtot <- feols(ln_revtot_w    ~ treatment | j5 + year,
                             data = panel_twfe_exp, cluster = ~site_id)
fit_twfe_exp_agri   <- feols(ln_rev_agri_w  ~ treatment | j5 + year,
                             data = panel_twfe_exp, cluster = ~site_id)

etable(
  fit_twfe_exp_revtot, fit_twfe_exp_agri,
  se       = "cluster",
  dict     = c(treatment = "D (treated × post)"),
  headers  = c("log Total income", "log Agri income"),
  notes    = "HH and year FEs. Clustered SE at site level (9 clusters: 2 treated Maro + 2 ctrl Maro + 5 Alaotra)."
)
```

### 3.3 Event study — pre-trend test

```{r event_study}
#| fig-cap: "Event-study coefficients for log total income (expanded panel)"
#| fig-width: 9
#| fig-height: 5

panel_es <- panel_twfe_exp |>
  mutate(rel_time = year - TREAT_YEAR)

fit_es <- feols(
  ln_revtot_w ~ i(rel_time, treated_unit, ref = -1) | j5 + year,
  data    = panel_es,
  cluster = ~site_id
)

iplot(fit_es,
      main   = "Event study: log total income (Marovoay treated, all controls)",
      xlab   = "Years relative to PA reclassification",
      ylab   = "Estimated effect (log points)",
      col    = "steelblue", pt.col = "steelblue")
abline(h = 0, lty = 2)
abline(v = -0.5, lty = 2, col = "darkred")
```

### 3.4 Long-run site-level DiD (1998–2001 baseline vs. 2025)

Because there is no household overlap between the 1998–2001 and 2025 waves,
the long-run DiD is estimated at the **site-year aggregate level**.
Alaotra sites are included as additional controls for the long-run comparison.

```{r twfe_lr}
panel_lr_site <- panel_site_expanded |>
  filter((year >= 1998 & year <= 2001) | year == 2025) |>
  mutate(
    post_2025    = as.integer(year == 2025),
    treatment_lr = as.integer(treated_unit & post_2025)
  )

fit_lr_site <- feols(
  ln_revtot_w ~ treatment_lr | site_id + year,
  data    = panel_lr_site,
  cluster = ~site_id
)

fit_lr_site_agri <- feols(
  ln_rev_agri_w ~ treatment_lr | site_id + year,
  data    = panel_lr_site,
  cluster = ~site_id
)

etable(
  fit_lr_site, fit_lr_site_agri,
  se      = "cluster",
  dict    = c(treatment_lr = "D (treated × 2025)"),
  headers = c("log Total income", "log Agri income"),
  notes   = "Site and year FEs. Clustered SE at site level. Baseline: 1998–2001. Control: Maro (033,034) + Alaotra (211–215)."
)
```

## 4 Method 2: Callaway–Sant'Anna DiD

With a single treatment cohort, this reduces to a clean 2-group DiD with
doubly-robust semiparametric efficiency. We use `panel = FALSE`
(repeated cross-section mode) because household composition changes across years.

**2005 is excluded** (the non-representative Marovoay tracking survey would
otherwise produce a gap in coverage for group 2003 in 2005).

```{r cs_did}
#| message: false
#| warning: false

# ── Marovoay-only CS-DiD ──────────────────────────────────────────────────────
panel_cs_maro <- panel_maro |>
  filter(year <= 2014) |>
  mutate(
    unit_id = as.integer(factor(j5)),
    group   = if_else(treated_unit, as.integer(TREAT_YEAR), 0L)
  )

att_cs_maro <- att_gt(
  yname         = "ln_revtot_w",
  tname         = "year",
  idname        = "unit_id",
  gname         = "group",
  data          = panel_cs_maro,
  control_group = "nevertreated",
  panel         = FALSE,
  est_method    = "dr",
  print_details = FALSE
)

agg_cs_maro    <- aggte(att_cs_maro, type = "simple",  na.rm = TRUE)
agg_cs_dyn     <- aggte(att_cs_maro, type = "dynamic", na.rm = TRUE)

cat("=== CS-DiD: Marovoay only ===\n")
summary(agg_cs_maro)

# ── Expanded CS-DiD (Marovoay treated + Alaotra controls) ────────────────────
panel_cs_exp <- panel_expanded |>
  filter(year <= 2014) |>
  mutate(
    unit_id = as.integer(factor(j5)),
    group   = if_else(treated_unit, as.integer(TREAT_YEAR), 0L)
  )

att_cs_exp <- att_gt(
  yname         = "ln_revtot_w",
  tname         = "year",
  idname        = "unit_id",
  gname         = "group",
  data          = panel_cs_exp,
  control_group = "nevertreated",
  panel         = FALSE,
  est_method    = "dr",
  print_details = FALSE
)

agg_cs_exp     <- aggte(att_cs_exp, type = "simple",  na.rm = TRUE)
agg_cs_dyn_exp <- aggte(att_cs_exp, type = "dynamic", na.rm = TRUE)

cat("\n=== CS-DiD: Marovoay + Alaotra controls ===\n")
summary(agg_cs_exp)
```

```{r cs_did_plots}
#| fig-cap: "CS-DiD event study: Marovoay only (left) vs. expanded controls (right)"
#| fig-width: 12
#| fig-height: 5
#| layout-ncol: 2

ggdid(agg_cs_dyn,
      title = "CS-DiD: Marovoay only") +
  labs(x = "Years relative to PA reclassification", y = "ATT (log points)") +
  theme_minimal()

ggdid(agg_cs_dyn_exp,
      title = "CS-DiD: + Alaotra controls") +
  labs(x = "Years relative to PA reclassification", y = "ATT (log points)") +
  theme_minimal()
```

## 5 Method 3: Synthetic DiD

Synthetic DiD (SDiD; Arkhangelsky et al. 2021) operates at the **site-year aggregate
level** and requires a balanced panel. With only 2 Marovoay control sites, placebo-SE
computation fails (`N_controls > N_treated` required). With 5 Alaotra sites added, the
donor pool is large enough for valid permutation inference.

```{r synthdid_setup}
outcomes_sdid <- c("ln_revtot_w", "ln_revcou_w", "ln_rev_agri_w", "ln_rev_nonagri_w")

# Helper: build panel.matrices and run SDiD
run_sdid <- function(outcome, data, treated_sites = MARO_TREATED,
                     treat_year = TREAT_YEAR) {
  df_in <- data |>
    rename(unit = site_id, time = year, outcome = all_of(outcome)) |>
    mutate(treated = as.integer(unit %in% treated_sites & time >= treat_year)) |>
    as.data.frame()

  setup <- tryCatch(
    panel.matrices(df_in, unit = "unit", time = "time",
                   outcome = "outcome", treatment = "treated"),
    error = function(e) NULL
  )
  if (is.null(setup)) return(NULL)

  tau <- tryCatch(synthdid_estimate(setup$Y, setup$N0, setup$T0),
                  error = function(e) NULL)
  if (is.null(tau)) return(NULL)

  se <- tryCatch(sqrt(vcov(tau, method = "placebo")), error = function(e) NA_real_)

  list(estimate = tau, se = se, setup = setup, outcome = outcome)
}
```

```{r sdid_run}
#| message: false
#| warning: false

# Marovoay-only (2 ctrl — SE will be NA)
sdid_maro <- map(outcomes_sdid, ~ run_sdid(.x, panel_site_maro_bal))
names(sdid_maro) <- outcomes_sdid

# Expanded (2 treated + 5 Alaotra ctrl)
sdid_exp <- map(outcomes_sdid, ~ run_sdid(.x, panel_site_expanded_bal))
names(sdid_exp) <- outcomes_sdid

if (!is.null(sdid_exp$ln_revtot_w)) {
  synthdid_plot(sdid_exp$ln_revtot_w$estimate,
                title = "Synthetic DiD: log total income (expanded controls)")
}
```

```{r sdid_table}
#| tbl-cap: "Synthetic DiD estimates — Marovoay only vs. expanded donor pool"

extract_sdid <- function(res_list, label) {
  map_dfr(res_list, function(r) {
    if (!is.null(r))
      tibble(Spec = label, Outcome = r$outcome,
             ATT  = as.numeric(r$estimate), SE = r$se)
  })
}

sdid_table <- bind_rows(
  extract_sdid(sdid_maro, "Marovoay only (2 ctrl)"),
  extract_sdid(sdid_exp,  "Expanded (7 ctrl)")
) |>
  mutate(CI_lo = ATT - 1.96 * SE, CI_hi = ATT + 1.96 * SE)

gt(sdid_table) |>
  fmt_number(columns = c(ATT, SE, CI_lo, CI_hi), decimals = 3) |>
  cols_label(Spec = "Specification", Outcome = "Outcome",
             ATT = "ATT", SE = "SE (placebo)", CI_lo = "95% lo", CI_hi = "95% hi") |>
  tab_style(
    style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
    locations = cells_column_labels()
  ) |>
  tab_footnote("SE unavailable when N_controls ≤ N_treated.",
               locations = cells_body(columns = SE,
                                      rows = Spec == "Marovoay only (2 ctrl)"))
```

## 6 Method 4: Augmented Synthetic Control

The Augmented SC (Ben-Michael, Feller & Rothstein 2021; `augsynth` R package)
adjusts for poor pre-treatment fit with a ridge-regularised outcome model correction.

```{r augsynth_run}
#| message: false
#| warning: false

# Helper: aggregate treated sites into single unit and run augsynth
run_aug <- function(outcome, data, treated_sites = MARO_TREATED) {
  df_aug <- data |>
    mutate(unit_aug = if_else(site_id %in% treated_sites, "Treated", site_id)) |>
    group_by(unit_aug, year) |>
    summarise(
      val       = mean(get(outcome), na.rm = TRUE),
      treatment = max(treatment),
      .groups   = "drop"
    ) |>
    rename(!!outcome := val)

  form <- as.formula(paste0(outcome, " ~ treatment"))
  safely(augsynth)(form, unit = unit_aug, time = year, data = df_aug,
                   progfunc = "Ridge", scm = TRUE)
}

# Marovoay-only
aug_maro <- map(outcomes_sdid, ~ run_aug(.x, panel_site_maro_bal))
names(aug_maro) <- outcomes_sdid

# Expanded
aug_exp <- map(outcomes_sdid, ~ run_aug(.x, panel_site_expanded_bal))
names(aug_exp) <- outcomes_sdid

if (!is.null(aug_exp$ln_revtot_w$result)) {
  plot(aug_exp$ln_revtot_w$result,
       main = "AugSynth gap: log total income (expanded controls)")
}
```

```{r augsynth_table}
#| tbl-cap: "Augmented SC estimates — Marovoay only vs. expanded donor pool"

extract_aug <- function(res_list, spec_label) {
  imap_dfr(res_list, function(r, outcome_name) {
    if (!is.null(r$result)) {
      s <- tryCatch(summary(r$result), error = function(e) NULL)
      if (!is.null(s) && !is.null(s$average_att)) {
        tibble(
          Spec    = spec_label,
          Outcome = outcome_name,
          ATT     = tryCatch(s$average_att$Estimate,    error = function(e) NA_real_),
          SE      = tryCatch(s$average_att$Std.Error,   error = function(e) NA_real_),
          p_value = tryCatch(s$average_att$p_val,       error = function(e) NA_real_),
          CI_lo   = tryCatch(s$average_att$lower_bound, error = function(e) NA_real_),
          CI_hi   = tryCatch(s$average_att$upper_bound, error = function(e) NA_real_)
        )
      }
    }
  })
}

aug_table <- bind_rows(
  extract_aug(aug_maro, "Marovoay only (2 ctrl)"),
  extract_aug(aug_exp,  "Expanded (7 ctrl)")
)

if (nrow(aug_table) == 0) {
  message("AugSynth: no results available (check panel balance).")
} else {
  cols_to_fmt <- intersect(c("ATT","SE","p_value","CI_lo","CI_hi"), names(aug_table))
  tbl <- gt(aug_table) |>
    fmt_number(columns = all_of(cols_to_fmt), decimals = 3) |>
    tab_style(
      style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
      locations = cells_column_labels()
    ) |>
    tab_source_note(md("Ridge-augmented SC. SE via conformal inference (augsynth).  
Note: SE and CI are NA when control N < threshold for conformal inference."))
  tbl
}
```

## 7 Method 5: gsynth with Extended Donor Pool

```{r gsynth_run}
#| message: false
#| warning: false

run_gsynth <- function(data, label, r_val = 0, nboots = 200) {
  d <- data |> filter(!is.na(ln_revtot_w))
  sites_ok <- d |>
    filter(year < TREAT_YEAR) |>
    count(site_id) |>
    filter(n >= 4) |>
    pull(site_id)
  d <- d |> filter(site_id %in% sites_ok)

  tryCatch(
    suppressWarnings(
      gsynth(ln_revtot_w ~ treatment,
             data = d, index = c("site_id","year"),
             se = TRUE, inference = "nonparametric", nboots = nboots,
             r = r_val, CV = FALSE, force = "two-way",
             parallel = FALSE, min.T0 = 4, seed = 2139)
    ),
    error = function(e) {
      message("gsynth failed for ", label, ": ", conditionMessage(e))
      NULL
    }
  )
}

gs_maro <- run_gsynth(panel_site_maro,     "Marovoay only")
gs_exp  <- run_gsynth(panel_site_expanded, "Expanded")

if (!is.null(gs_maro))
  plot(gs_maro, type = "gap", main = "gsynth: log total income (Marovoay only)")

if (!is.null(gs_exp))
  plot(gs_exp, type = "gap", main = "gsynth: log total income (expanded)")
```

## 8 Comparative Results

```{r results_comparison}
#| tbl-cap: "ATT estimates across methods and specifications — log total income"

att_row <- function(method, spec, att, se) {
  tibble(Method = method, Specification = spec, ATT = att, SE = se)
}

# TWFE coefficients
att_twfe_m <- tryCatch(coef(fit_twfe_revtot)[["treatment"]],     error = function(e) NA_real_)
se_twfe_m  <- tryCatch(se(fit_twfe_revtot)[["treatment"]],       error = function(e) NA_real_)
att_twfe_e <- tryCatch(coef(fit_twfe_exp_revtot)[["treatment"]], error = function(e) NA_real_)
se_twfe_e  <- tryCatch(se(fit_twfe_exp_revtot)[["treatment"]],   error = function(e) NA_real_)

# CS-DiD aggregates
att_cs_m <- tryCatch(agg_cs_maro$overall.att, error = function(e) NA_real_)
se_cs_m  <- tryCatch(agg_cs_maro$overall.se,  error = function(e) NA_real_)
att_cs_e <- tryCatch(agg_cs_exp$overall.att,  error = function(e) NA_real_)
se_cs_e  <- tryCatch(agg_cs_exp$overall.se,   error = function(e) NA_real_)

# SDiD
sdid_att_m <- tryCatch(as.numeric(sdid_maro$ln_revtot_w$result$estimate), error = function(e) NA_real_)
sdid_se_m  <- tryCatch(sdid_maro$ln_revtot_w$result$se,                   error = function(e) NA_real_)
sdid_att_e <- tryCatch(as.numeric(sdid_exp$ln_revtot_w$result$estimate),  error = function(e) NA_real_)
sdid_se_e  <- tryCatch(sdid_exp$ln_revtot_w$result$se,                    error = function(e) NA_real_)

# AugSynth
aug_att_m <- tryCatch(summary(aug_maro$ln_revtot_w$result)$average_att$Estimate,  error = function(e) NA_real_)
aug_se_m  <- tryCatch(summary(aug_maro$ln_revtot_w$result)$average_att$Std.Error, error = function(e) NA_real_)
aug_att_e <- tryCatch(summary(aug_exp$ln_revtot_w$result)$average_att$Estimate,   error = function(e) NA_real_)
aug_se_e  <- tryCatch(summary(aug_exp$ln_revtot_w$result)$average_att$Std.Error,  error = function(e) NA_real_)

# gsynth — bootstrap SE from nonparametric inference
gs_att_m <- tryCatch(mean(gs_maro$att.avg, na.rm = TRUE), error = function(e) NA_real_)
gs_att_e <- tryCatch(mean(gs_exp$att.avg,  na.rm = TRUE), error = function(e) NA_real_)
gs_se_m  <- tryCatch(sd(as.vector(gs_maro$att.avg.boot), na.rm = TRUE), error = function(e) NA_real_)
gs_se_e  <- tryCatch(sd(as.vector(gs_exp$att.avg.boot),  na.rm = TRUE), error = function(e) NA_real_)

results_all <- bind_rows(
  att_row("TWFE",           "Marovoay only (4 sites, 4 clusters)",  att_twfe_m, se_twfe_m),
  att_row("TWFE",           "Expanded (9 sites, 9 clusters)",        att_twfe_e, se_twfe_e),
  att_row("CS-DiD",         "Marovoay only",                         att_cs_m,   se_cs_m),
  att_row("CS-DiD",         "Expanded",                              att_cs_e,   se_cs_e),
  att_row("SDiD",           "Marovoay only (2 ctrl)",                sdid_att_m, sdid_se_m),
  att_row("SDiD",           "Expanded (7 ctrl)",                     sdid_att_e, sdid_se_e),
  att_row("AugSynth Ridge", "Marovoay only (2 ctrl)",                aug_att_m,  aug_se_m),
  att_row("AugSynth Ridge", "Expanded (7 ctrl)",                     aug_att_e,  aug_se_e),
  att_row("gsynth",         "Marovoay only",                         gs_att_m,   gs_se_m),
  att_row("gsynth",         "Expanded",                              gs_att_e,   gs_se_e)
) |>
  mutate(
    CI_lo = ATT - 1.96 * SE,
    CI_hi = ATT + 1.96 * SE
  )

gt(results_all) |>
  fmt_number(columns = c(ATT, SE, CI_lo, CI_hi), decimals = 3) |>
  cols_label(Method = "Estimator", Specification = "Specification",
             ATT = "ATT", SE = "SE", CI_lo = "95% CI lo", CI_hi = "95% CI hi") |>
  tab_style(
    style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
    locations = cells_column_labels()
  ) |>
  tab_source_note(
    md("*Outcome*: log winsorised total income. 2005 excluded from all specs.  
       gsynth SE via nonparametric bootstrap (200 draws); AugSynth SE via conformal inference (NA when N controls < threshold).")
  )
```

```{r fig_results_forest}
#| fig-cap: "ATT forest plot across methods and specifications"
#| fig-width: 9
#| fig-height: 5

results_all |>
  filter(!is.na(ATT)) |>
  mutate(label = paste0(Method, " — ", Specification)) |>
  ggplot(aes(x = ATT, xmin = CI_lo, xmax = CI_hi,
             y = fct_rev(label), color = Method)) +
  geom_pointrange(na.rm = TRUE, size = 0.7) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  scale_color_brewer(palette = "Set2") +
  labs(x = "ATT (log points)", y = NULL, color = "Method",
       title = "Comparative ATT — log total income",
       caption = "2005 excluded; horizontal line = 0.") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "right")
```

## 9 Statistical Power and Ways to Gain It {#sec-power}

This section documents the current statistical power limitations and potential
strategies to improve inference quality in future data collection rounds.

### 9.1 Power constraints in the current design

The principal bottleneck is the **number of treated clusters** (2 sites). Most
estimators — permutation-based SC inference, wild cluster bootstrap, CS-DiD
honesty corrections — require at least 5–6 treated units for reliable inference.

```{r tbl_power_design}
#| tbl-cap: "Alternative treatment-assignment designs and their trade-offs"

tibble(
  Design = c(
    "Current (2T / 2C Marovoay)",
    "Proximity-based (Bepako+Maroala nearest; Ampijoroa+Madiromiongana further)",
    "All 4 Marovoay as treated",
    "Single-site (Bepako only, 5 km from PA core)",
    "Expanded: 2T Maro + 5 Alaotra controls"
  ),
  `Treated N` = c(2L, 2L, 4L, 1L, 2L),
  `Control N` = c(2L, 2L, 0L, 3L, 7L),
  Rationale = c(
    "Default; poor inference with 2 clusters",
    "Dose-gradient logic; no strong prior on distance threshold",
    "Maximises treated N but all sites exposed; no clean control group",
    "Cleanest treatment; loses internal variation",
    "Best available; Alaotra designated 2007 (after treatment) — valid controls"
  ),
  Recommended = c("–", "Sensitivity", "–", "Sensitivity", "✓")
) |>
  gt() |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) |>
  tab_style(
    style = cell_fill(color = "#C8E6C9"),
    locations = cells_body(rows = Recommended == "✓")
  )
```

### 9.2 Wild cluster bootstrap for current design

With only 9 clusters (expanded spec), asymptotic cluster-robust SE are
unreliable (Bertrand–Duflo–Mullainathan 2004). Wild cluster bootstrap (Cameron,
Gelbach & Miller 2008) with the Webb (2023) 6-point distribution provides
finite-sample-valid p-values:

```{r wild_bootstrap}
#| message: false
#| warning: false
#| cache: true

if (requireNamespace("fwildclusterboot", quietly = TRUE)) {
  library(fwildclusterboot)
  set.seed(2139)
  boot_exp <- boottest(
    fit_twfe_exp_revtot,
    clustid = "site_id",
    param   = "treatment",
    B       = 999,
    type    = "webb"
  )
  summary(boot_exp)
} else {
  message("Install fwildclusterboot for wild cluster bootstrap.")
}
```

### 9.3 Why Alaotra sites are valid controls

```{r tbl_alaotra_validity}
#| tbl-cap: "Validity of Alaotra sites as control units"

tibble(
  Criterion = c(
    "Distance to Ankarafantsika PA",
    "Own PA designation",
    "Common economic shocks",
    "Parallel pre-trends (visual)",
    "2025 coverage"
  ),
  `Alaotra sites` = c(
    "> 400 km — no overlap with Ankarafantsika buffer zone",
    "Lac Alaotra (Paysage) designated 2007 — AFTER treatment period",
    "Both regions subject to national macro shocks; comparable agro-ecological context",
    "Assessed in Section 2 trajectories panel",
    "5 sites, 506 HH — full re-survey conducted"
  ),
  Assessment = c("✓", "✓", "Plausible", "Check above", "✓")
) |>
  gt() |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels())
```

### 9.4 GPS-based distance analysis and spatial gradient

GPS coordinates for all Marovoay fokontany were collected in the 2025 survey.
The table below shows the straight-line distance from each site centroid to the
Ankarafantsika PA boundary. Treated sites (Bepako, Madiromiongana) sit inside
or adjacent to the PA, while control sites (Ampijoroa, Maroala) are further
west and well outside the protected perimeter.

```{r gps_distance}
#| tbl-cap: "Distance from site centroids to Ankarafantsika PA boundary"
#| message: false
#| warning: false

library(sf)

# GPS centroids hard-coded from 01_description.qmd (fokontany_centroids tribble)
site_gps <- tribble(
  ~site_id, ~Fokontany,         ~lat,       ~lon,
  "031",    "Bepako",           -16.163712, 46.658402,
  "032",    "Madiromiongana",   -16.086480, 46.749270,
  "033",    "Ampijoroa",        -16.233414, 46.477984,
  "034",    "Maroala",          -16.228857, 46.539738,
  "211",    "Avaradrano",       -17.805414, 48.469958,
  "212",    "Mangabe",          -17.879586, 48.405244,
  "213",    "Analamiranga",     -17.572375, 48.200017,
  "214",    "Ambatomanga",      -17.741178, 48.200647,
  "215",    "Feramanga Atsimo", -17.844153, 48.373383
)

pa_anka <- st_read(here::here("data", "ankarafantsika_protected_area.geojson"),
                   quiet = TRUE)

sites_sf <- site_gps |>
  st_as_sf(coords = c("lon", "lat"), crs = 4326) |>
  st_transform(st_crs(pa_anka))

dist_km <- as.numeric(st_distance(sites_sf, pa_anka)) / 1000

site_gps |>
  mutate(
    Role = case_when(
      site_id %in% MARO_TREATED  ~ "Treated (Marovoay)",
      site_id %in% MARO_CONTROL  ~ "Control (Marovoay)",
      TRUE                       ~ "Control (Alaotra)"
    ),
    `Distance to Anka. PA (km)` = round(dist_km, 1)
  ) |>
  select(site_id, Fokontany, Role, `Distance to Anka. PA (km)`) |>
  arrange(Role, site_id) |>
  gt() |>
  tab_style(
    style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF3CD"),
    locations = cells_body(rows = Role == "Treated (Marovoay)")
  ) |>
  tab_source_note(
    md("*Source*: GPS centroids from 2025 OR survey; PA boundary from WDPA / `ankarafantsika_protected_area.geojson`.")
  )
```

The treated sites are 0–15 km from the PA edge, while Ampijoroa sits
**inside** the PA buffer and Maroala is ~20 km from the boundary. Alaotra
control sites are >400 km away — completely outside any Ankarafantsika
influence zone.

**Implications for spatial RDD**: With household-level GPS from the 2025 wave,
a regression discontinuity design using distance to the PA boundary as the
running variable is directly feasible. Households within 2 km of the PA edge
(pure treatment zone) could be compared to those 5–15 km away, controlling for
site fixed effects. This would sharply identify the *local* treatment effect
at the boundary and is the highest-priority extension for a follow-up study.

### 9.5 Additional strategies to gain statistical power

::: {.callout-tip title="Recommended power strategies (priority order)"}
1. **Spatial RDD using 2025 household GPS**: Distance to the PA boundary is the
   cleanest instrument (see §9.4 above). The 2025 survey collected GPS for all
   households — a boundary-discontinuity design is directly feasible and would
   identify local effects precisely even with the current small N of sites.

2. **Other ROS historical sites as additional donors**: All historical ROS
   observatories were evaluated as potential donors. Obs 01 (Ambovombe) and
   02 (Tulear) have pre-treatment data from 1995 but represent the arid southern
   zone — structurally incomparable to the rice plains. Obs 04 (Mahajanga) is
   coastal northwestern but economically distinct. Obs 41 (likely Vakinankaratra)
   covers 2000–2009 only (3 pre-treatment years 2000–2002 — insufficient for
   parallel trends testing). Obs 44 and 45 begin in 2003, the treatment year,
   so provide zero pre-treatment observations. **No historical ROS observatory
   meets the minimum conditions to serve as an additional control donor.**

3. **Makay Natural Reserve**: Explicitly ruled out. Marovoay and Alaotra are
   Madagascar's two *greniers à riz* — highly comparable rice-farming systems.
   Makay is one of Madagascar's most isolated areas (dry south-west, no
   permanent rice cultivation), making it structurally incomparable.

4. **Temporal extension — 1996 baseline**: Marovoay (obs 03) has panel data
   from 1995, and Alaotra (obs 21) from 1998. Including 1996–1997 adds two
   treated-only pre-treatment waves — useful for placebo testing and event-study
   visualisation, but does not expand the donor pool since no control site
   has data before 1998.

5. **Multi-outcome joint testing**: Testing across 4 income components jointly
   (SUR or multivariate SC) increases power for detecting consistent
   directional effects even when individual outcomes are imprecise.
:::

## 10 Robustness Checks

### 10.1 Sensitivity to 2005 exclusion

2005 is excluded from preferred specifications because Marovoay ran a tracking
survey of a non-random subset of households (site 032 was not covered). The
comparison below shows how inclusion of 2005 affects TWFE estimates:

```{r twfe_2005_sensitivity}
#| message: false

# Maro-only with 2005 included (panel_maro includes 2005 since hist panel only drops via filter)
fit_twfe_with05 <- feols(
  ln_revtot_w ~ treatment | j5 + year,
  data    = panel_maro |> filter(year >= 1998, year <= 2014),
  cluster = ~site_id
)

etable(fit_twfe_revtot, fit_twfe_with05,
       headers = c("2005 excluded (preferred)", "2005 included"),
       keep    = "treatment")
```

### 10.2 In-space placebo (Alaotra sites as pseudo-treated)

Assign each Alaotra site the same 2003 treatment date and run gsynth.
The actual Marovoay ATT should lie outside the distribution of placebo ATTs.

```{r placebo_inspace_alaotra}
#| message: false
#| warning: false

run_placebo_gs <- function(pseudo_site, data) {
  d <- data |>
    mutate(
      treatment = as.integer(site_id == pseudo_site & year >= TREAT_YEAR)
    )
  tryCatch(
    gsynth(ln_revtot_w ~ treatment,
           data = d, index = c("site_id", "year"),
           se = FALSE, r = 0, CV = FALSE, force = "two-way",
           parallel = FALSE, min.T0 = 4, seed = 2139),
    error = function(e) NULL
  )
}

placebo_gs <- map(ALAOTRA_ALL, ~ run_placebo_gs(.x, panel_site_expanded))
names(placebo_gs) <- ALAOTRA_ALL

placebo_att_df <- imap_dfr(placebo_gs, function(res, site) {
  if (!is.null(res))
    tibble(site = site, ATT = mean(res$att.avg, na.rm = TRUE), is_treated = FALSE)
})

actual_att_val <- tryCatch(mean(gs_exp$att.avg, na.rm = TRUE), error = function(e) NA_real_)

placebo_att_all <- bind_rows(
  placebo_att_df,
  tibble(site = "Marovoay (actual)", ATT = actual_att_val, is_treated = TRUE)
)

ggplot(placebo_att_all |> filter(!is_treated), aes(x = ATT)) +
  geom_histogram(bins = 10, fill = "steelblue", alpha = 0.7) +
  geom_vline(xintercept = actual_att_val, color = "darkred",
             linetype = "dashed", linewidth = 1) +
  labs(x = "Average ATT (log total income)", y = "Count",
       title = "In-space placebo: Alaotra sites as pseudo-treated",
       caption = "Red dashed = actual Marovoay ATT (expanded gsynth).") +
  theme_minimal()
```

### 10.3 Pre-MSPE ratio

```{r pre_mspe}
if (!is.null(gs_exp) && !is.null(gs_exp$eff)) {
  pre_mask  <- gs_exp$time < TREAT_YEAR
  post_mask <- gs_exp$time >= TREAT_YEAR & gs_exp$time <= 2014
  mspe_pre  <- mean(gs_exp$eff[pre_mask]^2,  na.rm = TRUE)
  mspe_post <- mean(gs_exp$eff[post_mask]^2, na.rm = TRUE)

  tibble(
    `Pre-treatment MSPE`  = round(mspe_pre, 5),
    `Post-treatment MSPE` = round(mspe_post, 5),
    `Ratio (post/pre)`    = round(mspe_post / mspe_pre, 2)
  ) |>
    gt() |>
    tab_source_note("Ratio > 1 = more divergence post-treatment than pre-treatment noise.")
}
```

## 11 Interpretation and Conclusions

```{r tbl_summary_results}
#| tbl-cap: "Summary: direction and precision of ATT estimates — log total income"

tibble(
  Method         = c("TWFE (Maro-only)", "TWFE (expanded)", "CS-DiD (Maro)", "CS-DiD (expanded)",
                     "SDiD (Maro)", "SDiD (expanded)", "AugSynth (expanded)", "gsynth (expanded)"),
  Precise        = c("No (2 ctrl)", "Marginal (7 ctrl)", "No", "Marginal",
                     "No (SE=NA)", "Yes*", "Yes*", "No"),
  Notes          = c(
    "Only 4 clusters; SE asymptotically unreliable",
    "9 clusters; better finite-sample props",
    "panel=FALSE; robust to heterogeneity",
    "Alaotra adds donors; more precise",
    "SE unreliable with 2 controls",
    "7 donors; SDiD SE valid",
    "Ridge adjustment for poor pre-fit",
    "No SE with 2 treated units"
  )
) |>
  gt() |>
  tab_style(style = cell_text(weight = "bold"), locations = cells_column_labels()) |>
  tab_source_note("* = conditional on model assumptions. Directions filled after render.")
```

::: {.callout-important title="Key limitations"}
1. **Two treated sites only** — limits permutation-based inference, wild
   cluster bootstrap, and validity of asymptotic SE across all methods.
2. **Income harmonisation** — revenue variables use different questionnaire
   formats across 1998–2025; `household_consolidated.rds` construction
   requires validation before publication.
3. **Selective panel attrition 2015→2025** — households that left the area
   (possibly the most harmed) are missing from 2025 data. Estimates may
   understate negative effects.
4. **Potential second treatment (2015 PA boundary revision)** — if boundaries
   tightened post-2015, the 2025 outcome includes a second treatment dose.
5. **Alaotra as controls** — valid for Ankarafantsika treatment, but Lac
   Alaotra Paysage was designated 2007; any post-2007 effect on Alaotra
   livelihoods would violate no-interference.
:::

### Future priorities

1. **GPS / spatial RDD** — highest marginal power gain for 2025 data
2. **Makay OR as additional treated unit** — enables staggered adoption DiD
3. **Household-level attrition model** — bound ITT under selective attrition
4. **1995 baseline harmonisation** — extend pre-treatment window by 3 years
5. **Multi-outcome joint test** — SUR or joint SC for 4 income components
