---
title: "Marovoay–Ankarafantsika: Household-Level and Long-Run Analysis"
author: BETSAKA project
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    self-contained: true
editor_options:
  chunk_output_type: console
---

```{r setup}
#| include: false
proj_root <- normalizePath(file.path(getwd(), ".."))
knitr::opts_knit$set(root.dir = proj_root)
setwd(proj_root)

library(tidyverse)
library(haven)
library(gt)
library(fixest)
library(gsynth)
library(sf)
library(scales)
library(purrr)

source("utils/helpers_report.R")
source("utils/sites.R")

MARO_TREATED  <- c("031", "032")
MARO_CONTROL  <- c("033", "034")
MARO_ALL      <- c(MARO_TREATED, MARO_CONTROL)
# Alaotra sites: 211 = Ambato (Avaradrano/Feramanga Atsimo/Mangabe),
#                212 = Amparafa (Ambatoharanana/Ambodivoara/Analamiranga),
#                213 = Morarano (Ambatomanga/Ambohidrony) — enters 2004
ALAOTRA_ALL   <- c("211", "212", "213")
TREAT_YEAR    <- 2003
PRE_YEARS     <- 1999:2002    # Alaotra starts 1999; 4-year pre-treatment window
HIST_YEARS    <- 1999:2014    # full historical window (excl. 2005 for Marovoay)

# All 7 sites (2 treated + 5 control) for the long-run analysis
ALL_SITES <- c(MARO_ALL, ALAOTRA_ALL)

site_labels <- c(
  "031" = "031 Bepako (T)",
  "032" = "032 Madiromiongana (T)",
  "033" = "033 Ampijoroa (C)",
  "034" = "034 Maroala (C)",
  "211" = "211 Ambato (C)",
  "212" = "212 Amparafa (C)",
  "213" = "213 Morarano (C)"
)

# Load income computation pipeline (handles 1995-2015 and 2025)
source("utils/calc_incomes_all_years.R")
```

## 1 Panel Design: The Rotating-Panel Constraint

### 1.1 What the attrition data tell us

The Madagascar ROS panel is a **rotating panel**, not a strict re-interview
panel. This is documented in detail by Chabé-Ferret et al. (2024,
*Sci. Data* 11: 1085; doi:10.1038/s41597-024-03879-9):

- Annual attrition rate ≈ **15%**, cumulative ≈ **80% over 10 years**
- Attriting households are replaced by new randomly drawn ones to maintain
  n ≈ 500 per observatory per wave
- Attrition is **non-random**: the dominant cause is out-migration (65% of
  traceable attritors), which correlates with income and shock exposure
- Household identifiers (`j5`) are **not fully stable** across all years:
  six observatory-years had attrition rates >75% attributable to ID code
  reshuffles; the 2025 `j5` codes are from a new enumeration and carry no
  continuity with the 1995–2014 series

### 1.2 Implications for estimation strategy

::: {.callout-important title="Household fixed effects are not appropriate here"}
The standard panel DiD with **household fixed effects** (`| j5 + year`) would
require the same households to be observed in both pre- and post-treatment
periods. With 15% annual exit, only ~44% of 1998 households remain in 2003,
and ~6% survive to 2014. Since treatment is assigned at the *site* level (not
the household level), household FE absorbs within-household variance without
providing the identification gains expected from a true balanced panel. The
2025 `j5` codes are entirely new and cannot be linked to any historical
household.
:::

**Valid approaches given a rotating panel:**

| Approach | Unit | FE structure | Requires HH tracking? |
|---|---|---|---|
| Repeated cross-sections DiD | Household | Site + year | No |
| Event study (RC) | Household | Site + year | No |
| Long-run 2×2 DiD | Household / site | Site + period | No |
| Synthetic control (SC/SDiD/gsynth) | Site-year | — | No |
| Spatial RDD (2025 cross-section) | Household | None | No |

All approaches below use one of these valid designs.

---

## 2 Panel Construction

```{r panel_build}
#| message: false

household_consolidated <- read_rds("data/household_consolidated.rds")

# ── Historical panel (1998–2014, 2005 Marovoay excluded) ─────────────────────
mk_income <- function(df) {
  df |>
    mutate(
      rev_riz  = coalesce(rev_riz,  0),
      rev_cu   = coalesce(rev_cu,   0),
      revel    = coalesce(revel,    0),
      revpeche = coalesce(revpeche, 0),
      revppal  = coalesce(revppal,  0),
      revsec   = coalesce(revsec,   0),
      revcou   = coalesce(revcou, revppal + revsec + rev_riz + rev_cu + revel + revpeche),
      revtot   = coalesce(revtot, revcou),
      rev_agri    = rev_riz + rev_cu + revel + revpeche,
      rev_nonagri = revppal + revsec
    )
}

# ── Historical panel: Marovoay 1999–2014 (2005 excluded) ────────────────────
panel_hist_maro <- household_consolidated |>
  filter(obs == "03", site_id %in% MARO_ALL,
         year %in% HIST_YEARS, !(year == 2005)) |>
  mk_income() |>
  mutate(
    treated_unit = site_id %in% MARO_TREATED,
    post         = year >= TREAT_YEAR,
    treatment    = as.integer(treated_unit & post),
    rel_time     = year - TREAT_YEAR
  ) |>
  # Winsorise at 99th percentile pooled across all site-years
  mutate(across(c(revtot, revcou, rev_agri, rev_nonagri),
                ~ pmin(.x, quantile(.x, 0.99, na.rm = TRUE)),
                .names = "{.col}_w")) |>
  mutate(across(ends_with("_w"), ~ log(.x + 1), .names = "ln_{.col}")) |>
  distinct(j5, year, .keep_all = TRUE)

# ── 2025 income (both observatories) via the standard pipeline ───────────────
# The historical snapshot (output/ref_incomes_snapshot.rds) ends at 2015.
# For 2025 we use compute_income_year() from utils/calc_incomes_all_years.R,
# which reads raw DTA files from data/ROS_MDG_microdata/2025/.
rev_2025_raw <- compute_income_year(2025) |>
  mutate(j5 = as.character(j5))

# ── 2025 wave — Marovoay (j0=3; fresh enumeration) ───────────────────────────
res_deb_25 <- read_dta("data/res_deb.dta") |>
  mutate(j5 = as.character(j5), j0 = as.character(as.numeric(j0)))

maro_25_ids <- res_deb_25 |>
  filter(j0 == "3") |>
  mutate(
    site_id = case_when(
      str_starts(str_to_lower(j4), "bepako")         ~ "031",
      str_starts(str_to_lower(j4), "madiromiongana") ~ "032",
      str_starts(str_to_lower(j4), "ampijoroa")      ~ "033",
      str_starts(str_to_lower(j4), "maroala")        ~ "034",
      TRUE ~ NA_character_
    ),
    obs = "03"
  ) |>
  select(j5, obs, site_id)

panel_2025_maro <- maro_25_ids |>
  left_join(rev_2025_raw, by = "j5") |>
  mutate(year = 2025) |>
  mk_income() |>
  mutate(
    treated_unit = site_id %in% MARO_TREATED,
    post = TRUE, treatment = as.integer(treated_unit),
    rel_time = 2025 - TREAT_YEAR
  ) |>
  mutate(across(c(revtot, revcou, rev_agri, rev_nonagri),
                ~ pmin(.x, quantile(.x, 0.99, na.rm = TRUE)),
                .names = "{.col}_w")) |>
  mutate(across(ends_with("_w"), ~ log(.x + 1), .names = "ln_{.col}"))

# ── 2025 wave — Alaotra (j0=21; control sites, fresh enumeration) ─────────────
# Site mapping: 211=Ambato (Avaradrano/Feramanga Atsimo/Mangabe),
#              212=Amparafa (Ambatoharanana/Ambodivoara/Analamiranga),
#              213=Morarano (Ambatomanga/Ambohidrony)
alaotra_25_ids <- res_deb_25 |>
  filter(j0 == "21") |>
  mutate(
    site_id = case_when(
      j4 %in% c("Avaradrano", "Feramanga Atsimo", "Mangabe")        ~ "211",
      j4 %in% c("Ambatoharanana", "Ambodivoara", "Analamiranga")   ~ "212",
      j4 %in% c("Ambatomanga", "Ambohidrony")                       ~ "213",
      TRUE ~ NA_character_
    ),
    obs = "21"
  ) |>
  select(j5, obs, site_id)

panel_2025_alaotra <- alaotra_25_ids |>
  filter(!is.na(site_id)) |>
  left_join(rev_2025_raw, by = "j5") |>
  mutate(year = 2025) |>
  mk_income() |>
  mutate(
    treated_unit = FALSE,
    post = TRUE, treatment = 0L,
    rel_time = 2025 - TREAT_YEAR
  ) |>
  mutate(across(c(revtot, revcou, rev_agri, rev_nonagri),
                ~ pmin(.x, quantile(.x, 0.99, na.rm = TRUE)),
                .names = "{.col}_w")) |>
  mutate(across(ends_with("_w"), ~ log(.x + 1), .names = "ln_{.col}"))

# ── Historical panel: Alaotra 1999–2014 (control only) ──────────────────────
panel_hist_alaotra <- household_consolidated |>
  filter(obs == "21", site_id %in% ALAOTRA_ALL, year %in% HIST_YEARS) |>
  mk_income() |>
  mutate(
    treated_unit = FALSE,
    post         = year >= TREAT_YEAR,
    treatment    = 0L,
    rel_time     = year - TREAT_YEAR
  ) |>
  mutate(across(c(revtot, revcou, rev_agri, rev_nonagri),
                ~ pmin(.x, quantile(.x, 0.99, na.rm = TRUE)),
                .names = "{.col}_w")) |>
  mutate(across(ends_with("_w"), ~ log(.x + 1), .names = "ln_{.col}")) |>
  distinct(j5, year, .keep_all = TRUE)

# ── Full RC DiD panel: Marovoay only (1999-2014 + 2025) ──────────────────────
# Historical Alaotra is NOT included here — parallel trends would need to be
# defended across two geographically distinct observatories for the medium run.
panel_maro_rc <- bind_rows(panel_hist_maro, panel_2025_maro) |>
  filter(!is.na(site_id))

# ── Long-run panel: Marovoay + Alaotra (1999-2002 pre vs 2025 post) ──────────
# Alaotra adds 3 control sites for the long-run comparison, boosting power.
# Valid because: (1) no PA in Alaotra until 2007 (temporary) / 2015 (permanent),
# so 1999-2002 pre-period is unaffected; (2) the Alaotra PA designation postdates
# the Ankarafantsika treatment by 4-12 years, so Alaotra is a clean control
# for the 2025 endpoint ONLY if we assume the Alaotra PA had no income effects
# by 2025 — a maintained assumption to be discussed in the text.
panel_lr_full <- bind_rows(
  panel_hist_maro  |> filter(year %in% PRE_YEARS),
  panel_hist_alaotra |> filter(year %in% PRE_YEARS),
  panel_2025_maro,
  panel_2025_alaotra
) |>
  filter(!is.na(site_id)) |>
  mutate(
    post_lr = as.integer(year == 2025),
    did_lr  = as.integer(treated_unit & post_lr)
  )

# Keep Marovoay-only long-run panel for comparison (original spec)
panel_lr2 <- bind_rows(
  panel_hist_maro |> filter(year %in% PRE_YEARS),
  panel_2025_maro
) |>
  filter(!is.na(site_id)) |>
  mutate(
    post_lr = as.integer(year == 2025),
    did_lr  = as.integer(treated_unit & post_lr)
  )

# ── Site-year aggregation (for gsynth comparison) ─────────────────────────────
panel_site_maro <- panel_hist_maro |>
  group_by(site_id, year, treated_unit, post, treatment) |>
  summarise(across(starts_with("ln_"), mean, na.rm = TRUE),
            n_hh = n(), .groups = "drop")
```

---

## 3 Attrition Diagnostics

### 3.1 Sample size per site per year

```{r fig_sample_size}
#| fig-cap: "Households surveyed per site per year (Marovoay, historical panel)"
#| fig-width: 9
#| fig-height: 4

panel_hist_maro |>
  count(site_id, year) |>
  mutate(site_label = site_labels[site_id]) |>
  ggplot(aes(x = year, y = n, color = site_label, linetype = site_label)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_vline(xintercept = 2002.5, linetype = "dashed", color = "darkred", alpha = 0.6) +
  annotate("text", x = 2002.5, y = Inf, label = " PA 2002",
           hjust = 0, vjust = 1.3, size = 3, color = "darkred") +
  scale_color_manual(values = c("#2D6A4F","#52B788","#1E6091","#74B3CE")) +
  labs(x = NULL, y = "Households surveyed", color = NULL, linetype = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

### 3.2 Distribution of observation spells (1998–2014)

How many times is each `j5` household identifier observed across the 16
historical waves? This approximates the re-interview rate — though note that
`j5` reassignments cause under-counting of true re-interviews.

```{r tbl_spells}
#| tbl-cap: "Distribution of household observation spells (Marovoay, 1998–2014)"

spells <- panel_hist_maro |>
  count(j5, name = "n_waves")

spell_summary <- spells |>
  count(n_waves, name = "n_households") |>
  mutate(pct = round(100 * n_households / sum(n_households), 1),
         cum_pct = cumsum(pct))

gt(spell_summary) |>
  cols_label(n_waves = "Waves observed", n_households = "Households",
             pct = "%", cum_pct = "Cumulative %") |>
  tab_style(style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
            locations = cells_column_labels()) |>
  tab_source_note(md(
    "Each `j5` code is counted across 1998–2014 (2005 excluded). Because `j5` codes
    are not fully stable across all years (Chabé-Ferret et al. 2024), this
    **overestimates** the share of single-observation households and **underestimates**
    true re-interview rates."
  ))
```

### 3.3 Differential attrition test

If attrition is correlated with treatment status, year-over-year sample
composition may diverge between treated and control sites. We test this by
comparing (a) annual sample counts and (b) mean household size over time
across treated vs. control sites.

```{r fig_diff_attrition}
#| fig-cap: "Year-over-year N and mean household size by treatment group"
#| fig-width: 10
#| fig-height: 4
#| layout-ncol: 2

panel_hist_maro |>
  mutate(Group = if_else(treated_unit, "Treated", "Control")) |>
  summarise(n = n(), mean_hh_size = mean(n_membres, na.rm = TRUE),
            .by = c(Group, year)) |>
  pivot_longer(c(n, mean_hh_size), names_to = "metric", values_to = "value") |>
  mutate(metric = recode(metric,
                         n = "Annual N (households)",
                         mean_hh_size = "Mean household size")) |>
  ggplot(aes(x = year, y = value, color = Group)) +
  geom_line(linewidth = 0.8) +
  geom_vline(xintercept = 2002.5, linetype = "dashed", color = "darkred", alpha = 0.5) +
  facet_wrap(~metric, scales = "free_y") +
  scale_color_manual(values = c("Treated" = "#2D6A4F", "Control" = "#1E6091")) +
  labs(x = NULL, y = NULL, color = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

```{r test_diff_attrition}
#| tbl-cap: "Test for differential trends in household size (composition check)"

# Linear trend in household size post-treatment, by group
# If PA restriction caused differential migration, household size should diverge
fit_hh_size <- feols(
  n_membres ~ treatment + i(year, treated_unit, ref = 2002) | site_id + year,
  data    = panel_hist_maro,
  cluster = ~site_id
)

# Simplified: just compare pre vs. post mean size by group
panel_hist_maro |>
  mutate(period = if_else(post, "Post-2002", "Pre-2002"),
         Group  = if_else(treated_unit, "Treated", "Control")) |>
  summarise(mean_size = round(mean(n_membres, na.rm = TRUE), 2),
            sd_size   = round(sd(n_membres,   na.rm = TRUE), 2),
            n         = n(),
            .by = c(Group, period)) |>
  arrange(Group, period) |>
  gt() |>
  cols_label(Group = "Group", period = "Period",
             mean_size = "Mean HH size", sd_size = "SD", n = "N obs") |>
  tab_style(style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
            locations = cells_column_labels()) |>
  tab_source_note(md(
    "If mean household size diverges between treated and control groups after 2002,
    this suggests differential out-migration — a potential source of composition bias."
  ))
```

---

## 4 Repeated Cross-Sections DiD

The repeated cross-sections (RC) DiD uses site + year fixed effects to identify
the treatment effect. Since treatment is assigned at the site level, this is
equivalent to the site-level TWFE but applied to household-level data — giving
tighter standard errors by using the full within-site variation in outcomes,
while still clustering at the site level to account for within-site correlation.

$$
\ln y_{ist} = \alpha_s + \lambda_t + \delta \cdot D_{st} + \varepsilon_{ist}
$$

where $D_{st} = \mathbf{1}[\text{site } s \in \text{Treated}] \times \mathbf{1}[t \geq 2003]$,
$\alpha_s$ are site fixed effects, and $\lambda_t$ year fixed effects.
Standard errors are clustered at the site level ($S = 4$).

### 4.1 Main specification

```{r rc_did_main}
#| tbl-cap: "Repeated cross-sections DiD — log income outcomes (Marovoay, 1998–2014)"
#| message: false

# Historical panel only (excluding 2025 — different composition, instrument changes)
panel_rc_hist <- panel_hist_maro |>
  filter(year %in% HIST_YEARS)  # 1999–2014

fit_rc_revtot    <- feols(ln_revtot_w    ~ treatment | site_id + year,
                          data = panel_rc_hist, cluster = ~site_id)
fit_rc_revcou    <- feols(ln_revcou_w    ~ treatment | site_id + year,
                          data = panel_rc_hist, cluster = ~site_id)
fit_rc_agri      <- feols(ln_rev_agri_w  ~ treatment | site_id + year,
                          data = panel_rc_hist, cluster = ~site_id)
fit_rc_nonagri   <- feols(ln_rev_nonagri_w ~ treatment | site_id + year,
                          data = panel_rc_hist, cluster = ~site_id)

etable(fit_rc_revtot, fit_rc_revcou, fit_rc_agri, fit_rc_nonagri,
       headers = c("Total", "Coupled", "Agri", "Non-agri"),
       keep    = "treatment",
       se.below = TRUE,
       notes   = "Site + year FE. Clustered SE (4 clusters). 2005 excluded.")
```

### 4.2 With household-size control

Household size may shift between treated and control sites if the PA
restriction caused differential out-migration. Adding `n_membres` controls
for this potential composition effect.

```{r rc_did_cov}
#| message: false

fit_rc_cov <- feols(
  ln_revtot_w ~ treatment + n_membres | site_id + year,
  data = panel_rc_hist, cluster = ~site_id
)

etable(fit_rc_revtot, fit_rc_cov,
       headers = c("No controls", "+ HH size"),
       keep    = c("treatment", "n_membres"),
       se.below = TRUE,
       notes   = "Site + year FE. Clustered SE (4 clusters).")
```

### 4.3 Event study (repeated cross-sections)

The event study uses site × year FE with relative-time indicators interacted
with the treated-unit dummy. Since treatment is site-level, the pre-treatment
coefficients are identified from the within-site outcome variation relative to
the between-group (treated vs. control) trend.

```{r fig_event_study}
#| fig-cap: "Event study — log total income (RC specification, site + year FE)"
#| fig-width: 9
#| fig-height: 4
#| message: false

panel_es <- panel_rc_hist |>
  mutate(rel_time = pmax(pmin(year - TREAT_YEAR, 11), -4))

fit_es_rc <- feols(
  ln_revtot_w ~ i(rel_time, treated_unit, ref = -1) | site_id + year,
  data    = panel_es,
  cluster = ~site_id
)

iplot(fit_es_rc,
      main  = "Event study: log total income (RC DiD, site+year FE)",
      xlab  = "Years relative to PA reclassification",
      ylab  = "ATT (log points)",
      pt.join = TRUE)
abline(h = 0, lty = 2, col = "grey40")
```

### 4.4 Exact permutation test (4 sites, 6 possible assignments)

With only 4 Marovoay sites and 2 treated, there are $\binom{4}{2} = 6$ possible
assignments of 2 treated sites. We enumerate all of them exhaustively and
compare the actual ATT to the permutation distribution. The minimum achievable
two-sided p-value under this design is $2/6 \approx 0.33$ — a hard floor set
by the number of units, not by the estimation method.

```{r permutation_rc}
#| tbl-cap: "Exact permutation test — RC DiD, log total income"

# All C(4,2) = 6 possible treatment assignments
perm_grid <- combn(MARO_ALL, 2, simplify = FALSE)

perm_atts <- map_dbl(perm_grid, function(treat_sites) {
  d <- panel_rc_hist |>
    mutate(treatment = as.integer(site_id %in% treat_sites & post))
  tryCatch(
    coef(feols(ln_revtot_w ~ treatment | site_id + year,
               data = d, cluster = ~site_id,
               warn = FALSE, notes = FALSE))[["treatment"]],
    error = function(e) NA_real_
  )
})

actual_att_rc <- coef(fit_rc_revtot)[["treatment"]]
p_two_sided   <- mean(abs(perm_atts) >= abs(actual_att_rc), na.rm = TRUE)
p_one_sided   <- mean(perm_atts >= actual_att_rc, na.rm = TRUE)

# Table: all permutations
tibble(
  `Treated sites` = map_chr(perm_grid, paste, collapse = " + "),
  `ATT`           = round(perm_atts, 4),
  `Is actual`     = map_lgl(perm_grid, ~ setequal(.x, MARO_TREATED))
) |>
  arrange(desc(`Is actual`), ATT) |>
  gt() |>
  tab_style(
    style = list(cell_fill(color = "#D4EDDA"), cell_text(weight = "bold")),
    locations = cells_body(rows = `Is actual`)
  ) |>
  tab_style(style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
            locations = cells_column_labels()) |>
  tab_footnote(
    paste0("Two-sided p = ", round(p_two_sided, 3),
           " (min achievable = 0.333). One-sided p = ", round(p_one_sided, 3), "."),
    locations = cells_column_labels(ATT)
  ) |>
  tab_source_note(md("Green row = actual treatment assignment. All 6 possible assignments enumerated."))
```

---

## 5 Long-Run 2×2 DiD: Pre-treatment vs. 2025

The 2025 BETSAKA survey provides a **23-year post-treatment observation**.
A long-run DiD comparing the pre-treatment period (1999–2002) to the 2025
wave bypasses all intermediate panel-gap issues: it requires no tracking and
no interpolation across the 2015–2024 gap. The estimand is the compound ATT
accumulating both the original 2002 Ankarafantsika reclassification and the
2015 boundary extension.

::: {.callout-warning title="What the long-run DiD measures"}
The 2025 endline reflects **23 years of cumulative treatment**, including both
the 2002 PA designation and the 2015 boundary extension (Décret n° 2015-730),
which further expanded the park perimeter. The long-run ATT cannot be
attributed to either event in isolation. Interpret as the total long-run
impact of the PA regime as of 2025.
:::

::: {.callout-note title="Alaotra controls: validity and timing"}
The Alaotra Lac observatory sites (211 Ambato, 212 Amparafa, 213 Morarano)
serve as additional control sites in the extended long-run specification.
Alaotra is subject to a distinct PA process: a **temporary protection decree
(2007)** followed by **permanent national protection (2015)**. Two implications
for validity:

1. **Pre-treatment period (1999–2002):** Alaotra has no PA restrictions.
   Parallel trends in the 1999–2002 baseline is plausible — both observatories
   are rural, rice-farming zones in Madagascar.
2. **Post-period (2025):** The 2015 Alaotra PA decree postdates Ankarafantsika
   by 13 years and may itself have affected Alaotra incomes. If the Alaotra PA
   *reduced* incomes, using Alaotra as a control in 2025 would **attenuate**
   the estimated Ankarafantsika effect (bias toward zero). If it *increased*
   incomes (e.g. through conservation payments or eco-tourism), the bias is
   upward. We treat this as a maintained assumption and report the Alaotra-
   extended spec alongside the Marovoay-only spec for transparency.
:::

### 5.1 Site-level comparison

```{r fig_longrun_trajectories}
#| fig-cap: "Income trajectories with 2025 endpoint — treated vs. control sites"
#| fig-width: 10
#| fig-height: 4

panel_site_maro |>
  bind_rows(
    panel_2025_maro |>
      group_by(site_id, year, treated_unit, post, treatment) |>
      summarise(across(starts_with("ln_"), mean, na.rm = TRUE),
                n_hh = n(), .groups = "drop")
  ) |>
  filter(year >= 1998) |>
  mutate(
    site_label = site_labels[site_id],
    Group = if_else(treated_unit, "Treated (031/032)", "Control (033/034)")
  ) |>
  ggplot(aes(x = year, y = ln_revtot_w, color = site_label, group = site_label)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_vline(xintercept = 2002.5, linetype = "dashed", color = "darkred", alpha = 0.7) +
  geom_vline(xintercept = 2014.5, linetype = "dotted", color = "grey50") +
  annotate("text", x = 2002.5, y = Inf, label = " PA 2002",
           hjust = 0, vjust = 1.4, size = 3, color = "darkred") +
  annotate("text", x = 2014.5, y = Inf, label = " Gap ",
           hjust = 1, vjust = 1.4, size = 3, color = "grey50") +
  scale_color_manual(values = c("#2D6A4F","#52B788","#1E6091","#74B3CE")) +
  scale_x_continuous(breaks = c(seq(1998, 2014, 4), 2025)) +
  labs(x = NULL, y = "Mean log income (winsorised)", color = NULL,
       caption = "Dotted: start of 10-year gap (2015–2024). 2005 excluded.") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

### 5.2 Long-run DiD regression (household-level, site + period FE)

Two specifications are reported side-by-side:

- **Marovoay only (4 sites)**: cleanest control group, tightest geographic
  comparability; minimum permutation p = 0.333.
- **Marovoay + Alaotra (7 sites)**: adds 3 Alaotra control sites, improving
  power; minimum permutation p = 0.095 with $\binom{7}{2} = 21$ allocations.
  Validity requires Alaotra incomes not to have been differentially affected
  by the 2015 Alaotra PA (see callout above).

```{r longrun_did}
#| message: false
#| tbl-cap: "Long-run DiD: log total income (pre-treatment 1999–2002 vs. 2025)"

# Marovoay-only specification (4 sites)
fit_lr <- feols(
  ln_revtot_w ~ did_lr | site_id + post_lr,
  data    = panel_lr2,
  cluster = ~site_id
)
fit_lr_agri <- feols(
  ln_rev_agri_w ~ did_lr | site_id + post_lr,
  data    = panel_lr2, cluster = ~site_id
)
fit_lr_nonagri <- feols(
  ln_rev_nonagri_w ~ did_lr | site_id + post_lr,
  data    = panel_lr2, cluster = ~site_id
)

# Extended specification: Marovoay + Alaotra (7 sites)
fit_lr_ext <- feols(
  ln_revtot_w ~ did_lr | site_id + post_lr,
  data    = panel_lr_full,
  cluster = ~site_id
)
fit_lr_ext_agri <- feols(
  ln_rev_agri_w ~ did_lr | site_id + post_lr,
  data    = panel_lr_full, cluster = ~site_id
)
fit_lr_ext_nonagri <- feols(
  ln_rev_nonagri_w ~ did_lr | site_id + post_lr,
  data    = panel_lr_full, cluster = ~site_id
)

etable(
  fit_lr, fit_lr_agri, fit_lr_nonagri,
  fit_lr_ext, fit_lr_ext_agri, fit_lr_ext_nonagri,
  headers = c("Total (Maro)", "Agri (Maro)", "Non-agri (Maro)",
              "Total (Maro+Al)", "Agri (Maro+Al)", "Non-agri (Maro+Al)"),
  keep    = "did_lr",
  se.below = TRUE,
  notes   = paste0(
    "Pre-period: ", min(PRE_YEARS), "–", max(PRE_YEARS),
    ". Post-period: 2025. Site + period FE. ",
    "Maro: 4 clusters; Maro+Alaotra: 7 clusters."
  ))
```

### 5.3 Exact permutation tests

We run two permutation tests: one for the 4-site Marovoay-only spec
($\binom{4}{2} = 6$ allocations, p-floor = 0.333) and one for the 7-site
extended spec ($\binom{7}{2} = 21$ allocations, p-floor = 0.095).

```{r longrun_perm}
#| tbl-cap: "Exact permutation test — long-run DiD, Marovoay only (4 sites)"

# ── 4-site permutation (Marovoay only) ───────────────────────────────────────
perm_atts_lr <- map_dbl(perm_grid, function(treat_sites) {
  d <- panel_lr2 |>
    mutate(
      treated_unit = site_id %in% treat_sites,
      did_lr       = as.integer(treated_unit & post_lr)
    )
  tryCatch(
    coef(feols(ln_revtot_w ~ did_lr | site_id + post_lr,
               data = d, cluster = ~site_id,
               warn = FALSE, notes = FALSE))[["did_lr"]],
    error = function(e) NA_real_
  )
})

actual_att_lr <- coef(fit_lr)[["did_lr"]]
p_two_lr  <- mean(abs(perm_atts_lr) >= abs(actual_att_lr), na.rm = TRUE)
p_one_lr  <- mean(perm_atts_lr >= actual_att_lr, na.rm = TRUE)

tibble(
  `Treated sites` = map_chr(perm_grid, paste, collapse = " + "),
  `ATT (long-run)` = round(perm_atts_lr, 4),
  `Is actual`      = map_lgl(perm_grid, ~ setequal(.x, MARO_TREATED))
) |>
  arrange(desc(`Is actual`), `ATT (long-run)`) |>
  gt() |>
  tab_style(
    style = list(cell_fill(color = "#D4EDDA"), cell_text(weight = "bold")),
    locations = cells_body(rows = `Is actual`)
  ) |>
  tab_style(style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
            locations = cells_column_labels()) |>
  tab_footnote(
    paste0("Two-sided p = ", round(p_two_lr, 3),
           " (min achievable = 0.333). One-sided p = ", round(p_one_lr, 3), "."),
    locations = cells_column_labels(`ATT (long-run)`)
  ) |>
  tab_source_note(md(
    "Pre-period: 1999–2002. Post-period: 2025. All 6 C(4,2) allocations enumerated."
  ))
```

```{r longrun_perm_ext}
#| tbl-cap: "Exact permutation test — long-run DiD, Marovoay + Alaotra (7 sites)"

# ── 7-site permutation (Marovoay + Alaotra) ──────────────────────────────────
perm_grid_ext <- combn(ALL_SITES, 2, simplify = FALSE)

perm_atts_lr_ext <- map_dbl(perm_grid_ext, function(treat_sites) {
  d <- panel_lr_full |>
    mutate(
      treated_unit = site_id %in% treat_sites,
      did_lr       = as.integer(treated_unit & post_lr)
    )
  tryCatch(
    coef(feols(ln_revtot_w ~ did_lr | site_id + post_lr,
               data = d, cluster = ~site_id,
               warn = FALSE, notes = FALSE))[["did_lr"]],
    error = function(e) NA_real_
  )
})

actual_att_lr_ext <- coef(fit_lr_ext)[["did_lr"]]
p_two_lr_ext <- mean(abs(perm_atts_lr_ext) >= abs(actual_att_lr_ext), na.rm = TRUE)
p_one_lr_ext <- mean(perm_atts_lr_ext >= actual_att_lr_ext, na.rm = TRUE)

tibble(
  `Treated sites` = map_chr(perm_grid_ext, paste, collapse = " + "),
  `ATT (long-run)` = round(perm_atts_lr_ext, 4),
  `Is actual` = map_lgl(perm_grid_ext, ~ setequal(.x, MARO_TREATED))
) |>
  arrange(desc(`Is actual`), `ATT (long-run)`) |>
  gt() |>
  tab_style(
    style = list(cell_fill(color = "#D4EDDA"), cell_text(weight = "bold")),
    locations = cells_body(rows = `Is actual`)
  ) |>
  tab_style(style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
            locations = cells_column_labels()) |>
  tab_footnote(
    paste0("Two-sided p = ", round(p_two_lr_ext, 3),
           " (min achievable = 0.095). One-sided p = ", round(p_one_lr_ext, 3), "."),
    locations = cells_column_labels(`ATT (long-run)`)
  ) |>
  tab_source_note(md(
    "Pre-period: 1999–2002. Post-period: 2025. All 21 C(7,2) allocations enumerated."
  ))
```

---

## 6 Site-Level Synthetic Control (for comparison)

The site-level gsynth result from the companion analysis
(`06_Marovoay-Ankarafantsika-extended.qmd`) is reproduced here for direct
comparison with the RC DiD estimates above.

```{r gsynth_ref}
#| message: false
#| warning: false
#| cache: true

run_gsynth_site <- function(data) {
  d <- data |>
    filter(!is.na(ln_revtot_w))
  suppressWarnings(
    gsynth(ln_revtot_w ~ treatment,
           data      = d,
           index     = c("site_id", "year"),
           se        = TRUE, inference = "nonparametric", nboots = 200,
           r = 0, CV = FALSE, force = "two-way",
           parallel  = FALSE, min.T0 = 4, seed = 2139)
  )
}

gs_maro <- tryCatch(run_gsynth_site(panel_site_maro), error = function(e) NULL)

if (!is.null(gs_maro))
  plot(gs_maro, type = "gap",
       main = "gsynth gap plot — site-level (Marovoay only)",
       ylab = "ATT (log points)", xlab = "Year")
```

---

## 7 Synthesis: All Estimates

```{r tbl_synthesis}
#| tbl-cap: "All ATT estimates — log total income, Marovoay–Ankarafantsika"

att_row <- function(method, spec, estimand, att, se, ci_lo = NA, ci_hi = NA) {
  tibble(Method = method, Specification = spec, Estimand = estimand,
         ATT = att, SE = se, CI_lo = ci_lo, CI_hi = ci_hi)
}

# RC DiD
rc_att  <- tryCatch(coef(fit_rc_revtot)[["treatment"]],  error = function(e) NA_real_)
rc_se   <- tryCatch(se(fit_rc_revtot)[["treatment"]],    error = function(e) NA_real_)

# RC DiD (with HH size control)
rcc_att <- tryCatch(coef(fit_rc_cov)[["treatment"]],     error = function(e) NA_real_)
rcc_se  <- tryCatch(se(fit_rc_cov)[["treatment"]],       error = function(e) NA_real_)

# Long-run DiD (Marovoay only)
lr_att  <- tryCatch(coef(fit_lr)[["did_lr"]],            error = function(e) NA_real_)
lr_se   <- tryCatch(se(fit_lr)[["did_lr"]],              error = function(e) NA_real_)

# Long-run DiD (Marovoay + Alaotra)
lre_att <- tryCatch(coef(fit_lr_ext)[["did_lr"]],        error = function(e) NA_real_)
lre_se  <- tryCatch(se(fit_lr_ext)[["did_lr"]],          error = function(e) NA_real_)

# gsynth site-level
gs_att  <- tryCatch(mean(gs_maro$att.avg, na.rm = TRUE), error = function(e) NA_real_)
gs_se   <- tryCatch(sd(as.vector(gs_maro$att.avg.boot), na.rm = TRUE), error = function(e) NA_real_)

bind_rows(
  att_row("RC DiD",           "Site + year FE",             "Medium-run ATT (1999–2014)", rc_att,  rc_se,
          rc_att  - 1.96 * rc_se,  rc_att  + 1.96 * rc_se),
  att_row("RC DiD + HH size", "Site + year FE + n_membres", "Medium-run ATT (1999–2014)", rcc_att, rcc_se,
          rcc_att - 1.96 * rcc_se, rcc_att + 1.96 * rcc_se),
  att_row("Long-run DiD (Maro)",  "Site + period FE, 4 sites", "Long-run ATT (1999–2002 vs 2025)",
          lr_att,  lr_se,  lr_att  - 1.96 * lr_se,  lr_att  + 1.96 * lr_se),
  att_row("Long-run DiD (Maro+Al)", "Site + period FE, 7 sites", "Long-run ATT (1999–2002 vs 2025)",
          lre_att, lre_se, lre_att - 1.96 * lre_se, lre_att + 1.96 * lre_se),
  att_row("gsynth (site)",    "IFE, nonparam bootstrap",    "Medium-run ATT (1999–2014)", gs_att, gs_se,
          gs_att - 1.96 * gs_se, gs_att + 1.96 * gs_se)
) |>
  gt() |>
  fmt_number(columns = c(ATT, SE, CI_lo, CI_hi), decimals = 3) |>
  cols_label(Method = "Estimator", Specification = "FE structure",
             Estimand = "Estimand", ATT = "ATT", SE = "SE",
             CI_lo = "95% CI lo", CI_hi = "95% CI hi") |>
  tab_style(style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
            locations = cells_column_labels()) |>
  tab_style(style = cell_fill(color = "#FFF3CD"),
            locations = cells_body(rows = Method == "Long-run DiD")) |>
  tab_source_note(md(
    "Outcome: log winsorised total income. CIs: 1.96 × cluster SE (4 clusters).
    gsynth SE via 200 nonparametric bootstrap draws.
    Long-run DiD captures 2002 + 2015 treatment effects jointly.
    See permutation tests (§4.4, §5.3) for finite-sample p-values."
  ))
```

---

## 8 Spatial RDD: Data Availability Assessment

The most powerful identification strategy would be a spatial RDD using
distance from the Ankarafantsika PA boundary as the running variable. The
2025 survey collected GPS coordinates for all survey sites.

```{r tbl_gps_assessment}
#| tbl-cap: "GPS data availability and spatial RDD feasibility"

tribble(
  ~`Data layer`, ~`Available?`, ~`Source`, ~`Implication`,
  "Fokontany centroid GPS", "Yes",
    "`01_description.qmd` (hardcoded tribble)",
    "Site-level distance to PA computable; too coarse for RDD",
  "Hamlet (hameau) GPS", "Yes",
    "`Coordonnées_GPS_OR_Alaotra_Marovoay_2025.ods`",
    "4–8 hamlets per site; still too coarse for within-site RDD",
  "Individual household GPS", "Unknown",
    "Would need field GPS from `res_deb.dta` or separate file",
    "Required for boundary-discontinuity RDD",
  "PA boundary polygon", "Yes",
    "`data/ankarafantsika_protected_area.geojson`",
    "Running variable = distance to boundary (metres)"
) |>
  gt() |>
  tab_style(style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
            locations = cells_column_labels()) |>
  tab_style(style = cell_fill(color = "#D4EDDA"),
            locations = cells_body(rows = `Available?` == "Yes")) |>
  tab_style(style = cell_fill(color = "#FFF3CD"),
            locations = cells_body(rows = `Available?` == "Unknown")) |>
  tab_source_note(md(
    "The hamlet-level GPS centroids give **4–8 points per site** — not enough for
    a smooth running-variable RDD. Household GPS would give ~120 observations per
    site in 2025, enabling a local linear RDD in the 0–15 km bandwidth around the
    PA boundary. If individual GPS was recorded in `res_deb.dta` (or a companion
    GPS file), it should be checked before concluding RDD is infeasible."
  ))
```

---

## 9 Interpretation

The rotating-panel structure of the Madagascar ROS data constrains but does
not preclude a credible impact evaluation. The key implications are:

1. **The site is the effective unit of observation** for treatment assignment.
   Any specification — whether household-level repeated cross-sections or
   site-level aggregation — ultimately clusters inference at the 4-site level.
   This hard ceiling on statistical power is intrinsic to the design.

2. **The RC DiD gains efficiency** relative to the site-level aggregate by
   using the full within-site variance in outcomes, not just site means. The
   point estimate should be identical to the site-level TWFE; the SE should be
   narrower if within-site variance is informative. With only $G = 4$ sites,
   conventional clustered SEs and the wild cluster bootstrap both undercover;
   the exact permutation test (§4.4) is the appropriate inference tool.

3. **The long-run DiD is the most policy-relevant estimate**: it captures
   the 23-year compound effect of the PA regime on household incomes, using
   a clean pre-treatment baseline (1999–2002) and a fresh 2025 cross-section.
   It is immune to the panel-gap problem and the `j5` instability issue. The
   extended specification including Alaotra controls (7 sites) reduces the
   minimum permutation p-value from 0.333 to 0.095, making statistically
   detectable effects possible for the first time in this design.

4. **Spatial RDD remains the priority extension**: if household-level GPS is
   embedded in `res_deb.dta` or a companion file, a boundary-discontinuity
   design in the 2025 cross-section would provide the sharpest identification
   and is entirely independent of the panel attrition problem.

## References

- Chabé-Ferret, S. *et al.* (2024). The Madagascar Rural Observatory Survey
  (1995–2015). *Scientific Data* 11: 1085.
  <https://doi.org/10.1038/s41597-024-03879-9>
- MacKinnon, J.G. & Webb, M.D. (2018). The wild bootstrap for few (treated)
  clusters. *Econometrics Journal* 21(2): 114–135.
- Callaway, B. & Sant'Anna, P. (2020). Difference-in-differences with
  multiple time periods. *Journal of Econometrics* 225(2): 200–230.
