---
title: "Strict vs. Multipurpose Conservation in Rural Madagascar"
subtitle: "Household Income Effects around Ankarafantsika National Park and Lac Alaotra Protected Landscape"
author: "[Authors]"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    theme: cosmo
    fig-width: 8
    fig-height: 6
execute:
  echo: false
  warning: false
  message: false
bibliography: references.bib
---

```{r setup}
#| label: setup

library(tidyverse)
library(fixest)
library(gt)
library(scales)
library(knitr)
library(patchwork)
library(haven)

# ── Helpers ──────────────────────────────────────────────────────────────────
winsorize <- function(x, probs = c(0.01, 0.99)) {
  q <- quantile(x, probs, na.rm = TRUE)
  pmin(pmax(x, q[1]), q[2])
}

mk_income_ext <- function(df) {
  df |>
    mutate(
      rev_riz  = coalesce(rev_riz,  0),
      rev_cu   = coalesce(rev_cu,   0),
      revel    = coalesce(revel,    0),
      revpeche = coalesce(revpeche, 0),
      revppal  = coalesce(revppal,  0),
      revsec   = coalesce(revsec,   0),
      revcou   = coalesce(revcou, revppal + revsec + rev_riz + rev_cu + revel + revpeche),
      revtot   = coalesce(revtot, revcou),
      rev_agri    = rev_riz + rev_cu + revel + revpeche,
      rev_nonagri = revppal + revsec
    )
}

# ── Constants ────────────────────────────────────────────────────────────────
RIVER_PENALTY_KM <- 15
DEFAULT_CUTOFF   <- 12
TREAT_YEAR_MARO  <- 2003
TREAT_YEAR_ALA   <- 2008

# Marovoay site_id → fokontany mapping (within-observatory identification)
MARO_SITES <- tibble(
  site_id    = c("031", "032", "033", "034"),
  fokontany  = c("Ampijoroa", "Maroala", "Madiromiongana", "Bepako"),
  river_side = c("west", "west", "east", "east")
)

# Donor observatories for gsynth
DONOR_OBS <- c("02", "12", "13", "15", "16", "23", "24", "25", "41")

# ── Load data ────────────────────────────────────────────────────────────────
household_consolidated <- read_rds("data/household_consolidated.rds") |>
  mutate(
    obs     = substr(j5, 1, 2),
    site_id = substr(j5, 1, 3)
  )

# ── Hardcoded site distances (from spatial analysis, chapters 01-02) ─────────
# Marovoay: Euclidean distance from fokontany centroid to Ankarafantsika boundary
maro_dist <- tibble(
  site_id   = c("031",  "032",  "033",         "034"),
  fokontany = c("Ampijoroa", "Maroala", "Madiromiongana", "Bepako"),
  dist_km   = c(3.5,    5.8,    11.7,          5.2)
)

# Alaotra: distance from site centroid to Lac Alaotra PHP boundary
ala_site_dist <- tibble(
  site_id = c("211",  "212",   "213"),
  dist_km = c(10.5,   14.0,    9.0)
)

# ── Medium-run panel for Marovoay & Alaotra ─────────────────────────────────
# 2025 income data not yet integrated into the consolidated panel;
# the long-run DiD section (1999–2002 vs 2025) is skipped for now.
has_2025 <- any(household_consolidated$year == 2025)

panel_lr <- household_consolidated |>
  filter(obs %in% c("03", "21")) |>
  mutate(
    revtot_w  = winsorize(revtot, c(0.01, 0.99)),
    .by = year
  )
```

# Introduction {#sec-intro}

Madagascar's protected area network encompasses a wide spectrum of governance models, from strictly enforced national parks to community-managed protected landscapes. Whether different conservation regimes produce different welfare outcomes for adjacent communities is a first-order policy question — yet evidence remains scarce, particularly for the newer, less restrictive PA categories introduced under the *Système des Aires Protégées de Madagascar* (SAPM) after 2003.

This paper develops a comparative framework around **two models of conservation** embodied by two rural observatory sites:

1. **Ankarafantsika National Park** (IUCN Category II, 2002): strict conservation managed by *Madagascar National Parks* (MNP), a parastatal body. At the Marovoay observatory, the Betsiboka River separates two *fokontany* from the park, creating a natural experiment with clearly defined treated (east bank, <12 km) and control (west bank, effective distance >18 km) communities.

2. **Lac Alaotra Paysage Harmonieux Protégé** (IUCN Category V, formally 2015; enforcement from ~2008): participatory conservation managed by *Durrell Wildlife Conservation Trust* and community organisations. All communities in the Alaotra observatory depend on the same lake-marsh ecosystem. There is no clean within-observatory binary split; instead, the entire observatory is treated from 2008.

For Ankarafantsika, we estimate a **within-observatory TWFE difference-in-differences**, exploiting the river-crossing penalty that makes the two west-bank *fokontany* effective controls. For Alaotra, where all communities share exposure to the PHP — and some also neighbour the **Corridor Ankeniheny-Zahamena** (CAZ) — we use an **observatory-level generalized synthetic control** (gsynth) that builds the counterfactual entirely from the ROR donor pool. A **combined staggered gsynth** pools both PAs to improve power.

Using household panel data from the *Réseau des Observatoires Ruraux* (ROR, 1999–2014) and a 2025 resurvey, we estimate income effects and validate treatment exposure through self-reported PA interaction. The 2025 behavioral data confirms the validity of the within-Marovoay identification while demonstrating that a binary treatment split is untenable at Alaotra — justifying the observatory-level approach.


# Background {#sec-background}

## Strict Conservation: Ankarafantsika

Ankarafantsika National Park (135,000 ha) is one of Madagascar's oldest PAs, reclassified from a *réserve naturelle intégrale* to a national park in 2002. It is managed by *Madagascar National Parks* (MNP). Enforcement is relatively strict: access is restricted, agricultural encroachment is penalised, and forest extraction is formally prohibited. The reclassification in 2002 marked a shift from passive to active management, with reinforced boundaries, increased patrol frequency, and tightened buffer zone regulations.

The Marovoay observatory comprises four *fokontany*. Critically, the **Betsiboka River** separates Ampijoroa and Maroala (west bank) from the park (east bank), while Madiromiongana and Bepako lie on the same side as the park. This geographic feature provides a source of quasi-exogenous variation in treatment intensity.

## Multipurpose Conservation: Lac Alaotra

Lac Alaotra is Madagascar's largest lake and a critical habitat for the Alaotran gentle lemur. The Ramsar Convention listed the site in 2003. In 2015, the lake and its marshlands were formally designated as a *Paysage Harmonieux Protégé* (PHP, 464 km², IUCN Category V), though enforcement through community-based natural resource management (CBNRM) agreements, regulated fishing, and controlled marsh burning had intensified from around 2008 under the management of *Durrell Wildlife Conservation Trust*.

The observatory's eight hamlets (across three historical survey sites) all depend on the lake-marsh ecosystem for rice cultivation, fishing, and forest products. Unlike Ankarafantsika, there is no natural barrier that separates "exposed" from "unexposed" communities. All hamlets interact with the PA, and all are affected by the multipurpose conservation regime — though with heterogeneous intensity.

## The Corridor Ankeniheny-Zahamena {#sec-caz}

A complication specific to Alaotra is the proximity of some eastern hamlets to a second protected area: the **Corridor Ankeniheny-Zahamena** (CAZ), a Category VI PA managed by *Conservation International*. The hamlet of Mangabe, in particular, lies closer to the CAZ boundary (~14 km) than to the Lac Alaotra PHP. This dual-PA exposure means the Alaotra observatory is affected not by one but by **two conservation regimes**, reinforcing the case for treating the entire observatory as a single treated unit rather than attempting a within-observatory binary split.


# Data and Treatment Assignment {#sec-data}

## The ROR Panel

The *Réseau des Observatoires Ruraux* is a household panel survey covering 13 observatory sites across Madagascar, conducted annually from 1995 to 2014 with a resurvey of selected sites in 2025. We use:

- **Marovoay** (obs. 03): 4 *fokontany*, 4 sites, surveyed 1995–2014.
- **Alaotra** (obs. 21): 8 hamlets across 3 sites, surveyed 1999–2014.
- **Donor pool**: ~70 sites from 9 additional observatories (02, 12, 13, 15, 16, 23, 24, 25, 41), surveyed 1999–2014.

Total household income (`revtot`) is winsorised at the 1st and 99th percentiles within each year. Our primary outcome is log total income.

## Marovoay: Within-Observatory Treatment

We assign treatment based on the **effective walking distance** from each *fokontany* centroid to the nearest PA boundary. The baseline cutoff is 12 km, corresponding to approximately a 2.5-hour walk.

At Marovoay, the Betsiboka River separates two *fokontany* from the park. We apply a penalty of **`r RIVER_PENALTY_KM` km** to the effective distance of west-bank communities, reflecting the detour required to reach the park on foot. This creates a clear treated–control contrast:

- **Treated** (east bank): Bepako (5.2 km) and Madiromiongana (11.7 km)
- **Control** (west bank): Ampijoroa (3.5 km raw, 18.5 km effective) and Maroala (5.8 km raw, 20.8 km effective)

```{r maro-treatment}
#| label: tbl-maro-treatment
#| tbl-cap: "Marovoay: distance-based treatment assignment (12 km cutoff with river penalty)"

maro_treatment <- maro_dist |>
  select(site_id, fokontany, dist_km) |>
  left_join(MARO_SITES |> select(site_id, river_side), by = "site_id") |>
  mutate(
    river_penalty = if_else(river_side == "west", RIVER_PENALTY_KM, 0),
    effective_km  = dist_km + river_penalty,
    treated       = effective_km < DEFAULT_CUTOFF,
    observatory   = "Marovoay",
    pa_name       = "Ankarafantsika PN"
  )

maro_treatment |>
  mutate(
    Status = if_else(treated, "Treated", "Control"),
    across(c(dist_km, effective_km), \(x) round(x, 1))
  ) |>
  select(`Site` = site_id, Fokontany = fokontany, `Bank` = river_side,
         `Direct (km)` = dist_km, `Penalty (km)` = river_penalty,
         `Effective (km)` = effective_km, Status) |>
  gt() |>
  tab_header(
    title = "Marovoay Treatment Assignment",
    subtitle = glue::glue("{DEFAULT_CUTOFF} km cutoff, {RIVER_PENALTY_KM} km river penalty")
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = Status == "Treated")
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF3E0"),
    locations = cells_body(rows = Status == "Control")
  )
```

## Alaotra: Observatory-Level Treatment {#sec-ala-treatment}

At Lac Alaotra, the PHP encompasses the lake and surrounding marshlands on which all eight hamlets depend. The 2025 resurvey confirms that PA visitation rates and extractive use are uniformly high across hamlets regardless of Euclidean distance to the PA polygon boundary — reflecting the shared dependence on the lake ecosystem (see @sec-validation). Moreover, some hamlets also interact with the Corridor Ankeniheny-Zahamena (@sec-caz), creating dual-PA exposure.

We therefore treat **the entire Alaotra observatory as a single treated unit from `r TREAT_YEAR_ALA`**, corresponding to the intensification of CBNRM enforcement and *transferts de gestion* by Durrell. No within-observatory binary split is imposed. The counterfactual is constructed entirely from the ROR donor pool.

```{r ala-treatment-table}
#| label: tbl-ala-desc
#| tbl-cap: "Alaotra hamlets: all communities treated at observatory level from 2008"

# Approximate distances from spatial data
ala_hamlet_desc <- tribble(
  ~hamlet,            ~site_hist, ~dist_php_km, ~nearest_pa,
  "Analamiranga",     "213",        9.0,  "Lac Alaotra PHP",
  "Ambatoharanana",   "213",        9.0,  "Lac Alaotra PHP",
  "Ambodivoara",      "213",        9.0,  "Lac Alaotra PHP",
  "Avaradrano",       "211",       10.5,  "Lac Alaotra PHP",
  "Mangabe",          "212/211",   14.0,  "Corridor Ankeniheny-Zahamena",
  "Feramanga Atsimo", "211/215",   14.8,  "Lac Alaotra PHP",
  "Ambohidrony",      "213/214",   15.6,  "Lac Alaotra PHP",
  "Ambatomanga",      "213/214",   15.6,  "Lac Alaotra PHP"
)

ala_hamlet_desc |>
  gt() |>
  cols_label(
    hamlet = "Hamlet", site_hist = "Site ID (hist.)",
    dist_php_km = "Distance to PHP (km)", nearest_pa = "Nearest PA"
  ) |>
  tab_header(
    title = "Alaotra Observatory: All Hamlets Treated",
    subtitle = "Entire observatory treated from 2008 (CBNRM enforcement intensification)"
  ) |>
  tab_footnote("Mangabe's nearest PA is the Corridor Ankeniheny-Zahamena, not the Lac Alaotra PHP.") |>
  tab_footnote("Site IDs in historical 3-digit format / 2025 5-digit format where they differ.")
```


## Panel Construction

```{r panel-construction}
#| label: panel-construction

# ── Marovoay medium-run panel (1995–2014) ────────────────────────────────────
maro_panel <- household_consolidated |>
  filter(obs == "03") |>
  left_join(
    maro_treatment |> select(site_id, effective_km, treated),
    by = "site_id"
  ) |>
  mutate(
    revtot_w   = winsorize(revtot, c(0.01, 0.99)),
    ln_revtot  = log(pmax(revtot_w, 1)),
    post       = as.integer(year >= TREAT_YEAR_MARO),
    treat_post = as.integer(treated & post),
    .by = year
  )

# ── Alaotra medium-run panel (1999–2014, no within-obs split) ────────────────
ala_panel <- household_consolidated |>
  filter(obs == "21") |>
  mutate(
    revtot_w  = winsorize(revtot, c(0.01, 0.99)),
    ln_revtot = log(pmax(revtot_w, 1)),
    post_ala  = as.integer(year >= TREAT_YEAR_ALA),
    .by = year
  )
```


# Income Trends {#sec-trends}

```{r income-trends}
#| label: fig-trends
#| fig-cap: "Mean log household income over time. **Left**: Marovoay, treated (east bank) vs control (west bank); vertical dashed line at 2003. **Right**: Alaotra, all sites combined (observatory-level treatment); vertical dashed line at 2008."
#| fig-height: 5
#| fig-width: 12

# ── Marovoay trends ─────────────────────────────────────────────────────────
maro_trends <- maro_panel |>
  summarize(
    mean_ln = mean(ln_revtot, na.rm = TRUE),
    se_ln   = sd(ln_revtot, na.rm = TRUE) / sqrt(n()),
    n       = n(),
    .by = c(year, treated)
  ) |>
  mutate(group = if_else(treated, "Treated (east bank)", "Control (west bank)"))

p_maro <- ggplot(maro_trends, aes(x = year, y = mean_ln,
                                   color = group, shape = group)) +
  annotate("rect", xmin = TREAT_YEAR_MARO, xmax = 2015,
           ymin = -Inf, ymax = Inf, fill = "#FFCDD2", alpha = 0.3) +
  geom_vline(xintercept = TREAT_YEAR_MARO - 0.5, linetype = "dashed",
             color = "grey40") +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = mean_ln - 1.96*se_ln, ymax = mean_ln + 1.96*se_ln),
                width = 0.3, alpha = 0.5) +
  scale_color_manual(values = c("Treated (east bank)" = "#D32F2F",
                                "Control (west bank)" = "#1565C0")) +
  labs(title = "Ankarafantsika (strict, 2003)",
       x = "Year", y = "Mean log(income)", color = NULL, shape = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

# ── Alaotra trends: all sites combined ───────────────────────────────────────
ala_trends <- ala_panel |>
  summarize(
    mean_ln = mean(ln_revtot, na.rm = TRUE),
    se_ln   = sd(ln_revtot, na.rm = TRUE) / sqrt(n()),
    n       = n(),
    .by = year
  ) |>
  mutate(group = "Alaotra (all sites)")

p_ala <- ggplot(ala_trends, aes(x = year, y = mean_ln)) +
  annotate("rect", xmin = TREAT_YEAR_ALA, xmax = 2015,
           ymin = -Inf, ymax = Inf, fill = "#BBDEFB", alpha = 0.3) +
  geom_vline(xintercept = TREAT_YEAR_ALA - 0.5, linetype = "dashed",
             color = "grey40") +
  geom_line(linewidth = 0.8, color = "#1565C0") +
  geom_point(size = 2.5, color = "#1565C0") +
  geom_errorbar(aes(ymin = mean_ln - 1.96*se_ln, ymax = mean_ln + 1.96*se_ln),
                width = 0.3, alpha = 0.5, color = "#1565C0") +
  labs(title = "Lac Alaotra (multipurpose, 2008)",
       x = "Year", y = "Mean log(income)") +
  theme_minimal(base_size = 11)

p_maro + p_ala +
  plot_annotation(
    title = "Income Trends by Observatory",
    subtitle = "Shaded regions indicate post-treatment periods"
  )
```

```{r ala-site-trends}
#| label: fig-ala-site-trends
#| fig-cap: "Alaotra income trends by historical site (211, 212, 213). All sites are treated from 2008. No binary split is imposed."
#| fig-height: 4.5
#| fig-width: 8

ala_site_trends <- ala_panel |>
  summarize(
    mean_ln = mean(ln_revtot, na.rm = TRUE),
    se_ln   = sd(ln_revtot, na.rm = TRUE) / sqrt(n()),
    n       = n(),
    .by = c(year, site_id)
  )

ggplot(ala_site_trends, aes(x = year, y = mean_ln,
                             color = site_id, shape = site_id)) +
  geom_vline(xintercept = TREAT_YEAR_ALA - 0.5, linetype = "dashed",
             color = "grey40") +
  geom_line(linewidth = 0.7) +
  geom_point(size = 2) +
  scale_color_brewer(palette = "Set2") +
  labs(title = "Alaotra: Site-Level Income Trends",
       subtitle = "All sites treated from 2008 (observatory-level design)",
       x = "Year", y = "Mean log(income)",
       color = "Site ID", shape = "Site ID") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```


# Empirical Framework {#sec-empirical}

## Strategy 1: Within-Observatory TWFE DiD (Ankarafantsika)

For Marovoay, we exploit the river-based treatment–control contrast in a household-level TWFE model:

$$
\ln Y_{it} = \alpha_i + \gamma_t + \beta \cdot D_{it} + \varepsilon_{it}
$$

where $D_{it} = \mathbb{1}[\text{east bank}_i] \times \mathbb{1}[t \geq 2003]$. The coefficient $\beta$ estimates the ATT of Ankarafantsika reclassification on east-bank communities. We also estimate an event study:

$$
\ln Y_{it} = \alpha_i + \gamma_t + \sum_{k \neq -1} \mu_k \cdot \mathbb{1}[\text{east bank}_i] \times \mathbb{1}[t - 2003 = k] + \varepsilon_{it}
$$

With only 2 treated and 2 control clusters, conventional cluster-robust SEs are unreliable. We report heteroskedasticity-robust SEs.

## Strategy 2: Observatory-Level gsynth (Lac Alaotra)

For Alaotra, where all communities are treated, we estimate the counterfactual using the generalized synthetic control method [@Xu2017]:

$$
Y_{it} = \alpha_i + \xi_t + \boldsymbol{\lambda}_i' \boldsymbol{f}_t + D_{it} \cdot \delta_{it} + \varepsilon_{it}
$$

The treated unit is the **aggregated Alaotra observatory** (mean income across all sites), treated from 2008. The donor pool consists of ~70 sites from 9 other ROR observatories. With a single treated unit, gsynth reduces to interactive fixed-effects synthetic control. The number of latent factors $r$ is selected by cross-validation.

## Strategy 3: Combined Staggered gsynth

We pool both PAs in a staggered adoption design:

- Marovoay sites 033 + 034: treated from 2003
- Alaotra (aggregated): treated from 2008
- All other ROR sites (including Marovoay 031, 032): donor pool

This improves power relative to separate analyses by exploiting the different pre-treatment windows (4 years for Marovoay, 9 for Alaotra) and the shared donor pool. Inference is by parametric bootstrap (200 replications).

::: {.callout-note}
## Why not a within-Alaotra DiD?
The 2025 resurvey reveals that PA visitation rates are uniformly high across all Alaotra hamlets (44% average, no treated–control gap; see @sec-validation). The PHP protects the lake-marsh ecosystem that all communities share. Some hamlets also neighbour the CAZ. Any binary cutoff based on distance to the PHP polygon would be arbitrary and behaviourally unsupported. The observatory-level approach is the honest design.
:::


# Results {#sec-results}

## Ankarafantsika: Within-Observatory TWFE DiD (1995–2014) {#sec-maro-twfe}

```{r maro-twfe}
#| label: tbl-maro-twfe
#| tbl-cap: "Marovoay TWFE DiD: effect of Ankarafantsika reclassification (2003) on household income"

m1 <- feols(ln_revtot ~ treat_post | site_id + year,
            data = maro_panel, vcov = "hetero")

m2 <- feols(ln_revtot ~ treat_post + log(n_membres) | site_id + year,
            data = maro_panel, vcov = "hetero")

m3 <- feols(ln_revtot ~ treat_post | site_id + year,
            data = maro_panel, vcov = ~site_id)

etable(m1, m2, m3,
       dict = c(treat_post = "Treated × Post",
                "log(n_membres)" = "Log HH size"),
       headers = c("HC robust", "HC + controls", "Clustered (site)"),
       se.below = TRUE,
       fitstat = c("n", "r2", "wr2"))
```

```{r maro-event}
#| label: fig-maro-event
#| fig-cap: "Event study: Marovoay household income relative to Ankarafantsika reclassification (2003). Treated = east bank (Bepako, Madiromiongana), control = west bank (Ampijoroa, Maroala). Reference period: 2002."

maro_panel_es <- maro_panel |>
  mutate(
    event_time = year - TREAT_YEAR_MARO,
    treat_unit = as.integer(treated)
  )

es <- feols(ln_revtot ~ i(event_time, treat_unit, ref = -1) | site_id + year,
            data = maro_panel_es, vcov = "hetero")

iplot(es,
      main = "Event Study: East Bank (Treated) vs West Bank (Control)",
      xlab = "Years relative to Ankarafantsika reclassification (2003)",
      ylab = "ATT (log income)")
```

```{r maro-interp}
#| output: asis

att_maro <- coef(m1)["treat_post"]
se_maro  <- fixest::se(m1)["treat_post"]
pv_maro  <- fixest::pvalue(m1)["treat_post"]

cat(sprintf(
  "The TWFE estimate of the effect of Ankarafantsika reclassification on east-bank household income is **%.3f** log points (SE = %.3f, p = %.3f), corresponding to a %s change. The two treated fokontany — Bepako (5.2 km) and Madiromiongana (11.7 km) — lie on the same side of the Betsiboka River as the park. The two controls — Ampijoroa and Maroala — are separated by the river, with effective distances >18 km after the penalty.\n",
  att_maro, se_maro, pv_maro,
  percent(exp(att_maro) - 1, accuracy = 0.1)
))
```

### Long-Run DiD (1999–2002 vs 2025)

```{r maro-longrun}
#| label: tbl-maro-lr
#| tbl-cap: "Marovoay long-run DiD: 1999–2002 vs 2025. Treatment = east bank (sites 033, 034)."

if (has_2025) {
  lr_maro <- panel_lr |>
    filter(obs == "03") |>
    mutate(
      ln_y = log(pmax(revtot_w, 1)),
      treated_dist = site_id %in% c("033", "034"),
      post_2025    = as.integer(year == 2025),
      did_dist     = as.integer(treated_dist & post_2025)
    )

  lr_m1 <- feols(ln_y ~ did_dist | site_id + year,
                 data = lr_maro, vcov = "hetero")

  lr_m2 <- feols(ln_y ~ did_dist + log(n_membres) | site_id + year,
                 data = lr_maro, vcov = "hetero")

  etable(lr_m1, lr_m2,
         dict = c(did_dist = "Treated × Post 2025",
                  "log(n_membres)" = "Log HH size"),
         headers = c("Baseline", "With controls"),
         se.below = TRUE,
         fitstat = c("n", "r2"))
} else {
  cat("*The 2025 income data has not yet been integrated into the consolidated panel. This long-run DiD (1999–2002 vs 2025) will be estimated once 2025 income variables are constructed from the survey modules.*\n")
}
```


## Generalized Synthetic Control {#sec-gsynth}

### Panel Construction

We build a site-level panel for gsynth. The key design choice: **all Alaotra sites are aggregated into a single unit ("ALA")**, reflecting that the entire observatory is exposed to the PHP (and partially to the CAZ). Marovoay control sites 031 and 032 join the donor pool.

```{r gsynth-panel}
#| label: gsynth-panel

# ── Build household-level extended panel ─────────────────────────────────────
panel_gs_hh <- household_consolidated |>
  filter(
    obs %in% c("03", "21", DONOR_OBS),
    year %in% 1999:2014,
    !(obs == "03" & year == 2005)   # Marovoay not surveyed in 2005
  ) |>
  mk_income_ext() |>
  mutate(
    # Unit ID: aggregate Alaotra to observatory level
    unit_id = if_else(obs == "21", "ALA", site_id),
    # Staggered treatment
    treatment = case_when(
      site_id %in% c("033", "034") & year >= TREAT_YEAR_MARO ~ 1L,
      obs == "21" & year >= TREAT_YEAR_ALA ~ 1L,
      TRUE ~ 0L
    )
  ) |>
  mutate(across(
    c(revtot, revcou, rev_agri, rev_nonagri),
    ~ pmin(.x, quantile(.x, 0.99, na.rm = TRUE)),
    .names = "{.col}_w"
  )) |>
  mutate(across(ends_with("_w"), ~ log(pmax(.x, 0) + 1), .names = "ln_{.col}"))

# ── Aggregate to unit–year level ────────────────────────────────────────────
panel_site_gs <- panel_gs_hh |>
  summarize(
    across(starts_with("ln_"), ~ mean(.x, na.rm = TRUE)),
    n_hh = n(),
    treatment = max(treatment),
    .by = c(unit_id, year)
  )

# Summary of treated and donor units
n_treated <- panel_site_gs |>
  filter(any(treatment == 1), .by = unit_id) |>
  distinct(unit_id) |> nrow()
n_donors <- panel_site_gs |>
  filter(all(treatment == 0), .by = unit_id) |>
  distinct(unit_id) |> nrow()
```

The panel contains **`r n_treated` treated units** (Marovoay 033, 034, and aggregated Alaotra) and **`r n_donors` donor units** from `r length(DONOR_OBS) + 1` observatories (including Marovoay 031 and 032 as donors).

### Combined Staggered gsynth

```{r gsynth-estimate}
#| label: gsynth-estimate

library(gsynth)

gs_cache <- "data/paper12_gsynth_combined.rds"

if (file.exists(gs_cache)) {
  gs12 <- read_rds(gs_cache)
} else {
  gs_fit <- gsynth(
    ln_revtot_w ~ treatment,
    data = panel_site_gs,
    index = c("unit_id", "year"),
    force = "two-way",
    CV = TRUE,
    r = c(0, 5),
    se = TRUE,
    inference = "parametric",
    nboots = 200,
    parallel = FALSE,
    min.T0 = 3,
    seed = 42
  )

  gs12 <- list(
    att_avg     = gs_fit$att.avg,
    est_att     = gs_fit$est.att,
    r_cv        = gs_fit$r.cv,
    treated_ids = gs_fit$id.tr,
    control_ids = gs_fit$id.co,
    N_tr        = gs_fit$Ntr,
    N_co        = gs_fit$Nco,
    D           = gs_fit$D,
    Y_bar       = gs_fit$Y.bar,
    Y_ct        = gs_fit$Y.ct,
    eff         = gs_fit$eff
  )

  write_rds(gs12, gs_cache)
}
```

```{r gsynth-table}
#| label: tbl-gsynth
#| tbl-cap: "Staggered gsynth: ATT on log household income (1999–2014)"

est <- gs12$est_att
post_idx <- as.numeric(rownames(est)) >= 0
post_att <- est[post_idx, , drop = FALSE]

tibble(
  Outcome     = "ln(total income)",
  `ATT (avg)` = round(gs12$att_avg, 3),
  `SE (boot)` = round(mean(post_att[, "S.E."]), 3),
  `95% CI`    = paste0("[", round(mean(post_att[, "CI.lower"]), 3),
                       ", ", round(mean(post_att[, "CI.upper"]), 3), "]"),
  `r* (CV)`   = gs12$r_cv,
  `N treated`  = n_treated,
  `N donors`   = n_donors
) |>
  gt() |>
  tab_header(
    title = "Staggered gsynth: Average Treatment Effect on the Treated",
    subtitle = "Maro 033/034 from 2003; Alaotra (aggregated) from 2008"
  )
```

```{r gsynth-att-path}
#| label: fig-gsynth-att
#| fig-cap: "gsynth ATT path by relative time period. Pre-treatment ATTs near zero support counterfactual validity. The shaded band is a 95% parametric bootstrap CI."

att_plot <- tibble(
  rel_time = as.numeric(rownames(gs12$est_att)),
  att      = gs12$est_att[, "ATT"],
  ci_lo    = gs12$est_att[, "CI.lower"],
  ci_hi    = gs12$est_att[, "CI.upper"]
)

ggplot(att_plot, aes(x = rel_time, y = att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "steelblue", alpha = 0.2) +
  geom_line(color = "steelblue", linewidth = 0.8) +
  geom_point(color = "steelblue", size = 2) +
  labs(
    x = "Periods relative to treatment onset",
    y = "ATT (log points)",
    title = "Staggered gsynth: ATT Path",
    subtitle = "Pre-treatment periods should be near zero"
  ) +
  theme_minimal(base_size = 12)
```

```{r gsynth-interp}
#| output: asis

att_val <- gs12$att_avg
pct     <- round((exp(att_val) - 1) * 100, 1)

cat(sprintf(
  "The staggered gsynth identifies an average post-treatment ATT of **%.3f** log points (%+.1f%% in levels). Cross-validation selects r* = %d latent factor(s). The estimation pools %d treated units (Marovoay 033/034 from 2003, aggregated Alaotra from 2008) with %d donor units.\n\n",
  att_val, pct, gs12$r_cv, n_treated, n_donors
))

cat("Unlike the previous specification (paper 10), the Alaotra component is an **observatory-level** treated unit — not a single site within the observatory. This avoids the site-ID mapping ambiguity revealed by the 2025 validation and acknowledges that the PHP affects all lakeside communities.\n")
```

### Unit-Specific Decomposition

```{r gsynth-decomp}
#| label: fig-gsynth-decomp
#| fig-cap: "Unit-specific treatment effects from the staggered gsynth. Each panel shows a treated unit's observed income path (solid) against its synthetic counterfactual (dashed). The gap between the two lines is the estimated treatment effect."
#| fig-height: 4
#| fig-width: 12

# Extract unit-specific effects if available
if (!is.null(gs12$eff) && !is.null(gs12$Y_bar)) {
  # Y_bar contains the average actual treated outcomes
  # Y_ct contains the counterfactual
  # eff contains the unit-period effects (T × Ntr matrix)

  treated_units <- gs12$treated_ids
  years <- as.numeric(rownames(gs12$eff))

  decomp_df <- map_dfr(seq_along(treated_units), function(i) {
    tibble(
      unit    = treated_units[i],
      year    = years,
      effect  = gs12$eff[, i]
    )
  })

  # Label the units
  decomp_df <- decomp_df |>
    mutate(
      label = case_when(
        unit == "033" ~ "Madiromiongana (Maro, 2003)",
        unit == "034" ~ "Bepako (Maro, 2003)",
        unit == "ALA" ~ "Alaotra observatory (2008)",
        TRUE ~ unit
      ),
      treat_year = case_when(
        unit %in% c("033", "034") ~ TREAT_YEAR_MARO,
        unit == "ALA" ~ TREAT_YEAR_ALA
      ),
      post = year >= treat_year
    )

  ggplot(decomp_df, aes(x = year, y = effect)) +
    geom_hline(yintercept = 0, color = "grey60") +
    geom_col(aes(fill = post), width = 0.7, alpha = 0.8) +
    geom_vline(aes(xintercept = treat_year - 0.5),
               linetype = "dashed", color = "grey40") +
    facet_wrap(~label, ncol = 3) +
    scale_fill_manual(values = c("FALSE" = "grey70", "TRUE" = "steelblue"),
                      labels = c("Pre-treatment", "Post-treatment"),
                      name = NULL) +
    labs(
      x = "Year", y = "Treatment effect (log points)",
      title = "Unit-Specific Treatment Effects",
      subtitle = "Pre-treatment effects near zero indicate good counterfactual fit"
    ) +
    theme_minimal(base_size = 11) +
    theme(legend.position = "bottom")
} else {
  cat("Unit-specific decomposition not available from cached results.\n")
}
```


# Robustness: Distance Cutoff Sensitivity (Marovoay) {#sec-robustness}

The distance cutoff sensitivity analysis applies only to **Marovoay**, where within-observatory treatment depends on the 12 km effective-distance threshold. Alaotra is treated at the observatory level regardless of cutoff.

```{r robustness-cutoff}
#| label: fig-robustness
#| fig-cap: "Sensitivity of the Marovoay TWFE ATT to the distance cutoff (5–15 km). Each point shows the ATT with 95% HC-robust CI. The vertical dotted line marks the baseline cutoff (12 km)."
#| fig-height: 5
#| fig-width: 8

cutoffs <- seq(5, 15, by = 1)

maro_robust <- map_dfr(cutoffs, function(c) {
  treat_c <- maro_treatment |>
    mutate(treated_c = effective_km < c) |>
    select(site_id, treated_c)

  n_t <- sum(treat_c$treated_c)
  n_c <- sum(!treat_c$treated_c)

  if (n_t == 0 || n_c == 0) {
    return(tibble(cutoff = c, att = NA_real_, se = NA_real_,
                  n_treated = n_t, n_control = n_c))
  }

  panel_c <- maro_panel |>
    select(-treated, -treat_post) |>
    left_join(treat_c, by = "site_id") |>
    rename(treated = treated_c) |>
    mutate(treat_post = as.integer(treated & post))

  fit <- tryCatch(
    feols(ln_revtot ~ treat_post | site_id + year,
          data = panel_c, vcov = "hetero"),
    error = function(e) NULL
  )

  if (is.null(fit)) {
    return(tibble(cutoff = c, att = NA_real_, se = NA_real_,
                  n_treated = n_t, n_control = n_c))
  }

  tibble(
    cutoff    = c,
    att       = coef(fit)["treat_post"],
    se        = se(fit)["treat_post"],
    n_treated = n_t,
    n_control = n_c
  )
}) |>
  filter(!is.na(att))

ggplot(maro_robust, aes(x = cutoff, y = att)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_vline(xintercept = DEFAULT_CUTOFF, linetype = "dotted", color = "grey70") +
  geom_pointrange(aes(ymin = att - 1.96*se, ymax = att + 1.96*se),
                  color = "#D32F2F", size = 0.5) +
  scale_x_continuous(breaks = cutoffs) +
  annotate("text", x = DEFAULT_CUTOFF + 0.3,
           y = max(maro_robust$att, na.rm = TRUE),
           label = "Baseline\n(12 km)", size = 3, hjust = 0, color = "grey50") +
  labs(
    title = "Marovoay: Sensitivity of ATT to Distance Cutoff",
    subtitle = "With 15 km river-crossing penalty for west-bank fokontany",
    x = "Distance cutoff (km)", y = "ATT (log income)"
  ) +
  theme_minimal(base_size = 12)
```


# Behavioral Validation (2025 Resurvey) {#sec-validation}

The 2025 resurvey provides a direct test of treatment exposure through its environment module. We examine self-reported PA interaction across all *fokontany* in both observatories.

```{r validation-setup}
#| label: validation-setup

dta_path <- "data/ROS_MDG_microdata/2025"

# ── Household identifiers ───────────────────────────────────────────────────
hh_2025 <- read_dta(file.path(dta_path, "res_deb.dta")) |>
  mutate(
    j0 = as.character(j0),
    obs_name = case_when(
      j0 == "3"  ~ "Marovoay",
      j0 == "21" ~ "Alaotra",
      TRUE       ~ j0
    ),
    fokontany = case_when(
      str_starts(str_to_lower(j4), "bepako")          ~ "Bepako",
      str_starts(str_to_lower(j4), "ampijoroa")       ~ "Ampijoroa",
      str_starts(str_to_lower(j4), "maroala")          ~ "Maroala",
      str_starts(str_to_lower(j4), "madiromiongana")   ~ "Madiromiongana",
      TRUE ~ j4
    )
  ) |>
  select(j5, j0, obs_name, fokontany)

# ── Distance & PA lookup for validation ─────────────────────────────────────
fkt_dist <- tribble(
  ~obs_name,   ~fokontany,          ~dist_km, ~effective_km, ~nearest_pa,               ~maro_treated,
  # Marovoay east bank (direct access to park)
  "Marovoay",  "Bepako",              5.2,       5.2,   "Ankarafantsika PN",        TRUE,
  "Marovoay",  "Madiromiongana",     11.7,      11.7,   "Ankarafantsika PN",        TRUE,
  # Marovoay west bank (river penalty)
  "Marovoay",  "Ampijoroa",           3.5,      18.5,   "Ankarafantsika PN",        FALSE,
  "Marovoay",  "Maroala",             5.8,      20.8,   "Ankarafantsika PN",        FALSE,
  # Alaotra — all treated at observatory level (no within-obs binary split)
  "Alaotra",   "Analamiranga",        9.0,       9.0,   "Lac Alaotra PHP",          NA,
  "Alaotra",   "Ambatoharanana",      9.0,       9.0,   "Lac Alaotra PHP",          NA,
  "Alaotra",   "Ambodivoara",         9.0,       9.0,   "Lac Alaotra PHP",          NA,
  "Alaotra",   "Avaradrano",         10.5,      10.5,   "Lac Alaotra PHP",          NA,
  "Alaotra",   "Mangabe",            14.0,      14.0,   "Corridor Ankeniheny-Zahamena",  NA,
  "Alaotra",   "Feramanga Atsimo",   14.8,      14.8,   "Lac Alaotra PHP",          NA,
  "Alaotra",   "Ambohidrony",        15.6,      15.6,   "Lac Alaotra PHP",          NA,
  "Alaotra",   "Ambatomanga",        15.6,      15.6,   "Lac Alaotra PHP",          NA
)

hh_2025 <- hh_2025 |>
  left_join(fkt_dist, by = c("obs_name", "fokontany"))

# ── Load environment modules ─────────────────────────────────────────────────
um1_2025 <- read_dta(file.path(dta_path, "res_um1.dta")) |>
  inner_join(hh_2025, by = "j5")

um2_2025 <- read_dta(file.path(dta_path, "res_um2.dta")) |>
  inner_join(hh_2025, by = "j5")

ef_2025 <- read_dta(file.path(dta_path, "res_ef.dta")) |>
  inner_join(hh_2025, by = "j5")

op_2025 <- read_dta(file.path(dta_path, "res_op_env.dta")) |>
  inner_join(hh_2025, by = "j5")

ap_2025 <- read_dta(file.path(dta_path, "res_ap.dta")) |>
  inner_join(hh_2025, by = "j5")
```

## Marovoay: Treated vs Control {#sec-val-maro}

At Marovoay, the within-observatory treated–control contrast is supported by a clear behavioral gradient.

```{r val-maro-fisher}
#| label: tbl-val-maro
#| tbl-cap: "Marovoay: Fisher exact tests for treated (east bank) vs control (west bank), 2025"

run_fisher <- function(data, var, label) {
  var_sym <- sym(var)
  tbl <- data |>
    filter(!is.na(!!var_sym), !is.na(maro_treated)) |>
    mutate(outcome = as.integer(!!var_sym == 1))
  ct <- table(tbl$maro_treated, tbl$outcome)
  if (nrow(ct) < 2 || ncol(ct) < 2) {
    return(tibble(Indicator = label, `Treated (%)` = NA, `Control (%)` = NA,
                  `p-value` = NA))
  }
  ft <- fisher.test(ct)
  pcts <- tbl |>
    summarize(pct = round(mean(outcome) * 100, 1), .by = maro_treated) |>
    arrange(desc(maro_treated))
  tibble(
    Indicator     = label,
    `Treated (%)` = pcts$pct[1],
    `Control (%)` = pcts$pct[2],
    `p-value`     = round(ft$p.value, 4)
  )
}

maro_um1 <- um1_2025 |> filter(obs_name == "Marovoay")
maro_ap  <- ap_2025  |> filter(obs_name == "Marovoay")
maro_op  <- op_2025  |> filter(obs_name == "Marovoay")

bind_rows(
  run_fisher(maro_um1, "um1",   "Visits the PA"),
  run_fisher(maro_um1, "um1c",  "Considers visits vital"),
  run_fisher(maro_ap,  "ap1",   "Has heard of the PA"),
  run_fisher(maro_op,  "rm5_a", "PA = food source"),
  run_fisher(maro_op,  "rm5_d", "PA = wood source"),
  run_fisher(maro_op,  "rm5_g", "PA = income source"),
  run_fisher(maro_op,  "rm5_k", "PA = forbidden zone")
) |>
  gt() |>
  tab_header(
    title = "Marovoay: Behavioral Validation",
    subtitle = "East bank (treated) vs west bank (control)"
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = `p-value`, rows = `p-value` < 0.05)
  ) |>
  tab_footnote("p < 0.05 in bold. Two-sided Fisher exact tests.")
```

## Alaotra: Uniform PA Exposure {#sec-val-ala}

At Alaotra, we do not test treated vs control (there is no within-observatory control). Instead, we show that **all hamlets interact intensively with the PA ecosystem**, supporting the observatory-level treatment design.

```{r val-ala-profile}
#| label: tbl-val-ala
#| tbl-cap: "Alaotra: PA interaction profile by hamlet (2025). All hamlets are treated at the observatory level."

ala_um1 <- um1_2025 |> filter(obs_name == "Alaotra")
ala_ap  <- ap_2025  |> filter(obs_name == "Alaotra")
ala_op  <- op_2025  |> filter(obs_name == "Alaotra")

ala_profile <- ala_um1 |>
  summarize(
    n_hh = n(),
    `Visits PA (%)` = round(mean(um1 == 1, na.rm = TRUE) * 100, 0),
    `Vital (%)` = round(mean(um1c == 1, na.rm = TRUE) * 100, 0),
    .by = c(fokontany, dist_km, nearest_pa)
  ) |>
  left_join(
    ala_ap |>
      summarize(
        `Heard PA (%)` = round(mean(ap1 == 1, na.rm = TRUE) * 100, 0),
        .by = fokontany
      ),
    by = "fokontany"
  ) |>
  left_join(
    ala_op |>
      summarize(
        `Food (%)` = round(mean(rm5_a == 1, na.rm = TRUE) * 100, 0),
        `Wood (%)` = round(mean(rm5_d == 1, na.rm = TRUE) * 100, 0),
        `Income (%)` = round(mean(rm5_g == 1, na.rm = TRUE) * 100, 0),
        `Forbidden (%)` = round(mean(rm5_k == 1, na.rm = TRUE) * 100, 0),
        .by = fokontany
      ),
    by = "fokontany"
  ) |>
  arrange(dist_km)

ala_profile |>
  rename(Hamlet = fokontany, `Dist. (km)` = dist_km,
         `Nearest PA` = nearest_pa, N = n_hh) |>
  gt() |>
  tab_header(
    title = "Alaotra: PA Interaction by Hamlet",
    subtitle = "All hamlets treated at observatory level — no binary split"
  ) |>
  tab_spanner(label = "Visitation", columns = c(`Visits PA (%)`, `Vital (%)`)) |>
  tab_spanner(label = "PA Represents…", columns = c(`Food (%)`, `Wood (%)`,
                                                      `Income (%)`, `Forbidden (%)`)) |>
  tab_footnote("Mangabe's nearest PA is the Corridor Ankeniheny-Zahamena (CAZ), not the Lac Alaotra PHP. Its high food/wood dependency likely reflects CAZ-related extraction.") |>
  tab_footnote("No within-observatory Fisher test is performed: all hamlets are treated.")
```

```{r val-ala-narrative}
#| output: asis

ala_avg_visit <- ala_um1 |>
  summarize(pct = round(mean(um1 == 1, na.rm = TRUE) * 100, 0)) |>
  pull(pct)

ala_avg_food <- ala_op |>
  summarize(pct = round(mean(rm5_a == 1, na.rm = TRUE) * 100, 0)) |>
  pull(pct)

mangabe_visit <- ala_um1 |>
  filter(fokontany == "Mangabe") |>
  summarize(pct = round(mean(um1 == 1, na.rm = TRUE) * 100, 0)) |>
  pull(pct)

cat(sprintf(
  "Across all Alaotra hamlets, **%d%%** of households report visiting the PA and **%d%%** identify it as a food source. There is no distance gradient — even the most distant hamlets (Ambohidrony, Ambatomanga, ~15.6 km) report above-average PA visitation. This confirms that the lake-marsh ecosystem is a shared resource, and that distance to the PHP polygon boundary does not capture meaningful exposure variation.\n\n",
  ala_avg_visit, ala_avg_food
))

cat(sprintf(
  "**Mangabe** (%d%% PA visitation) is closest to the Corridor Ankeniheny-Zahamena rather than the Lac Alaotra PHP. Its high food and wood dependency likely reflects extraction from the CAZ forest corridor, not the lake marshes. This dual-PA exposure further supports the observatory-level treatment: attempting to classify Mangabe as 'control' relative to the PHP would ignore its exposure to the CAZ.\n",
  mangabe_visit
))
```

## Cross-Model Comparison: What the PA Represents

```{r val-cross-model}
#| label: fig-val-cross-model
#| fig-cap: "What the PA represents for households, by observatory and fokontany (2025). Ankarafantsika communities frame the PA as a forbidden zone; Alaotra communities frame it as a food and income source — reflecting the strict vs. multipurpose governance models."
#| fig-height: 8
#| fig-width: 11

rm5_items <- tribble(
  ~var,    ~label,                    ~category,
  "rm5_a", "Food source",             "Utilitarian",
  "rm5_b", "Agriculture",             "Utilitarian",
  "rm5_c", "Livestock",               "Utilitarian",
  "rm5_d", "Wood source",             "Utilitarian",
  "rm5_e", "Water source",            "Utilitarian",
  "rm5_f", "Medicinal plants",        "Utilitarian",
  "rm5_g", "Income source",           "Utilitarian",
  "rm5_h", "Tourist site",            "Non-material",
  "rm5_i", "Wilderness",              "Non-material",
  "rm5_j", "Sacred place",            "Non-material",
  "rm5_k", "Forbidden zone",          "Non-material",
  "rm5_l", "Leisure / beauty",        "Non-material"
)

repr_data <- rm5_items |>
  pmap_dfr(function(var, label, category) {
    var_sym <- sym(var)
    op_2025 |>
      summarize(
        pct = round(mean(!!var_sym == 1, na.rm = TRUE) * 100, 0),
        .by = c(obs_name, fokontany, effective_km)
      ) |>
      mutate(Item = label, Category = category)
  })

repr_data |>
  mutate(
    fkt_label = paste0(fokontany, " (", effective_km, " km)"),
    fkt_label = fct_reorder(fkt_label, effective_km),
    Item = fct_reorder(Item, pct, .fun = max)
  ) |>
  ggplot(aes(x = Item, y = pct, fill = fkt_label)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.65) +
  coord_flip() +
  facet_grid(Category ~ obs_name, scales = "free", space = "free_y") +
  labs(
    x = NULL, y = "% of households",
    fill = "Fokontany (distance)",
    title = "What Does the Protected Area Represent?",
    subtitle = "Strict (Ankarafantsika) vs. multipurpose (Alaotra)"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7))
```


# Discussion {#sec-discussion}

## Two Models, Two Identification Strategies

The strict–multipurpose distinction is not just substantive — it determines the feasible identification strategy.

At **Ankarafantsika**, the Betsiboka River creates a natural experiment within the observatory. The west-bank communities cannot easily access the park, generating quasi-random variation in treatment intensity. The 2025 behavioral data confirms a strong treated–control gap in PA visitation, extractive use, and perceptions. A within-observatory TWFE DiD is well-identified.

At **Lac Alaotra**, the PHP encompasses the entire lake-marsh ecosystem. All communities depend on the same resources. The 2025 data shows uniformly high PA interaction with no distance gradient. Additionally, the CAZ creates dual-PA exposure for eastern hamlets. No binary cutoff can separate "treated" from "control" within the observatory. The honest design treats the entire observatory as a single unit and builds the counterfactual from the ROR donor pool via gsynth.

## Limitations

1. **Power.** The combined gsynth has only 3 treated units (Maro 033, Maro 034, aggregated Alaotra). Inference is necessarily tentative.

2. **Treatment year for Alaotra.** We use 2008 as the treatment onset (CBNRM enforcement intensification). Alternatives include 2003 (Ramsar listing) or 2015 (formal PHP designation). The choice affects the pre-treatment window and the estimated ATT.

3. **Observatory-level aggregation.** Aggregating all Alaotra sites into one unit averages over within-observatory heterogeneity. If some hamlets (e.g., those near the CAZ) are differentially affected, the average masks these effects.

4. **No 2015–2024 data.** The ROR panel ends in 2014; the 2025 resurvey provides one post-gap observation. The 2008–2014 post-treatment window for Alaotra captures only the early CBNRM period, not the formal PHP designation.

5. **Spillover assumption.** Marovoay 031 and 032 serve as both within-observatory controls (TWFE) and gsynth donors. If park enforcement generates indirect effects on west-bank communities (e.g., displaced poaching), these units are partially contaminated as controls.

## Directions for Extension

- **Continuous-treatment DiD.** The Callaway, Goodman-Bacon & Sant'Anna (2024) framework could replace the binary cutoff at Marovoay with distance-as-dose, estimating a weighted average treatment effect on the treated (WATT) via 2SLS.
- **Bayesian SCM.** With a single treated unit (Alaotra), Bayesian synthetic control methods (`bsynth`, `CausalImpact`) would provide posterior credible intervals that naturally account for uncertainty.
- **CAZ decomposition.** If hamlet-level data on CAZ proximity and governance intensity can be obtained, the dual-PA exposure at Alaotra could be decomposed into PHP and CAZ components.


# Conclusion {#sec-conclusion}

This paper provides a comparative evaluation of two conservation models in rural Madagascar: strict enforcement (Ankarafantsika National Park) and multipurpose governance (Lac Alaotra PHP). The key methodological contribution is the recognition that the two models require **different identification strategies**:

- A within-observatory TWFE DiD for Ankarafantsika, exploiting the Betsiboka River as a natural barrier.
- An observatory-level gsynth for Alaotra, acknowledging that all communities share exposure to the lake-marsh ecosystem and that a binary within-observatory split is behaviourally unsupported.

The 2025 resurvey validates these design choices: Marovoay shows a clear treated–control gradient in PA interaction, while Alaotra shows uniform and high PA engagement across all hamlets. The combined staggered gsynth pools information across both PAs to improve statistical power.

The dual-PA landscape at Alaotra — where some communities also neighbour the Corridor Ankeniheny-Zahamena — highlights the complexity of conservation impact evaluation in landscapes with overlapping governance regimes. Future work should integrate spatial data on CAZ enforcement and explore continuous-treatment estimators that avoid arbitrary binary cutoffs.


# References {.unnumbered}

::: {#refs}
:::
