---
title: "Conservation Governance and Household Income: Five Observatories"
subtitle: "Extending the Strict vs. Multipurpose Analysis to Farafangana, Toliara North, and Fénérive East"
author: "[Authors]"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    theme: cosmo
    fig-width: 8
    fig-height: 6
execute:
  echo: false
  warning: false
  message: false
bibliography: references.bib
---

```{r setup}
#| label: setup

library(tidyverse)
library(sf)
library(fixest)
library(gt)
library(scales)
library(knitr)
library(patchwork)
library(gsynth)
library(ggspatial)
library(ggrepel)
library(osmdata)

# -- Helpers -------------------------------------------------------------------
winsorize <- function(x, probs = c(0.01, 0.99)) {
  q <- quantile(x, probs, na.rm = TRUE)
  pmin(pmax(x, q[1]), q[2])
}

mk_income_ext <- function(df) {
  df |>
    mutate(
      rev_riz  = coalesce(rev_riz,  0),
      rev_cu   = coalesce(rev_cu,   0),
      revel    = coalesce(revel,    0),
      revpeche = coalesce(revpeche, 0),
      revppal  = coalesce(revppal,  0),
      revsec   = coalesce(revsec,   0),
      revcou   = coalesce(revcou, revppal + revsec + rev_riz + rev_cu + revel + revpeche),
      revtot   = coalesce(revtot, revcou),
      rev_agri    = rev_riz + rev_cu + revel + revpeche,
      rev_nonagri = revppal + revsec
    )
}

save_gsynth <- function(fit) {
  list(
    att_avg = fit$att.avg,
    est_att = fit$est.att,
    est_avg = fit$est.avg,
    r_cv    = fit$r.cv,
    N_tr    = fit$Ntr,
    N_co    = fit$Nco,
    eff     = fit$eff
  )
}

# -- Constants -----------------------------------------------------------------
# Treatment years (first full calendar year after the operative decree)
TREAT_YEAR_MARO <- 2003  # Ankarafantsika decree Aug 2002
TREAT_YEAR_ALA  <- 2008  # Alaotra CBNRM enforcement from 2008
TREAT_YEAR_FARA <- 2006  # CFAV temp decree 15 Sep 2006 (within survey year)
TREAT_YEAR_TOL  <- 2007  # Amoron'i Onilahy decree 17 Jan 2007
TREAT_YEAR_FEN  <- 2007  # Réserve de Tampolo decree 20 Aug 2007 (within survey year)

# Donor pool: CLEAN donors only (see Chapter 13 for contamination audit)
# Excludes Antsohihy (12), Farafangana (15), Toliara North (23), Fénérive East (24)
CLEAN_DONOR_OBS <- c("02", "13", "16", "25", "41")

# All treated observatories
TREATED_OBS <- c("03", "21", "15", "23", "24")

# All observatories used in this chapter
ALL_OBS_14 <- c(TREATED_OBS, CLEAN_DONOR_OBS)

# PA colours (one per treatment arm)
PA_COLS <- c(
  "Marovoay (strict)"       = "#D32F2F",
  "Alaotra (multipurpose)"  = "#1565C0",
  "Farafangana (multipurpose)" = "#2E7D32",
  "Toliara North (mixed)"   = "#E65100",
  "Fénérive East (low int.)"= "#6A1B9A"
)

# -- Load household data -------------------------------------------------------
household_consolidated <- read_rds("data/household_consolidated.rds") |>
  mutate(
    obs     = substr(j5, 1, 2),
    site_id = substr(j5, 1, 3)
  )

oecd_scales <- read_rds("data/oecd_equiv_scale.rds")
household_consolidated <- household_consolidated |>
  left_join(oecd_scales, by = c("j5", "year")) |>
  mutate(oecd_equiv = replace_na(oecd_equiv, 1))

# -- Load spatial data ---------------------------------------------------------
# Vahatra PAs for Marovoay / Alaotra maps
vahatra <- st_read("data/Vahatra/Shapefiles/AP_Vahatra.shp", quiet = TRUE) |>
  st_make_valid()

ankarafantsika_pa <- vahatra |> filter(grepl("Ankarafantsika", nom))
alaotra_pa        <- vahatra |> filter(grepl("Lac Alaotra", nom))
caz_pa            <- vahatra |> filter(grepl("Ankeniheny Zahamena", nom))
zahamena_pa       <- vahatra |> filter(grepl("Zahamena", nom) & !grepl("Ankeniheny", nom))

# Dynamic WDPA for the 3 new observatory maps (temporally accurate)
STUDY_START <- as.Date("1999-01-01")
STUDY_END   <- as.Date("2014-12-31")

dyn_wdpa <- read_rds("data/dynamic_wdpa.rds") |>
  st_make_valid()

dyn_study <- dyn_wdpa |>
  filter(zone_type == "external_boundary",
         valid_from <= STUDY_END,
         is.na(valid_to) | valid_to >= STUDY_START) |>
  group_by(WDPAID, NAME) |>
  summarise(geometry = st_union(geometry),
            valid_from = min(valid_from, na.rm = TRUE),
            .groups = "drop") |>
  st_make_valid()

# OSM water (cached, for Marovoay map)
osm_cache_maro <- "data/osm_water_maro.rds"
if (file.exists(osm_cache_maro)) {
  osm_water_maro <- read_rds(osm_cache_maro)
} else {
  q_riv <- opq(bbox = c(46.20, -16.60, 47.20, -15.80)) |>
    add_osm_feature(key = "waterway", value = "river") |> osmdata_sf()
  q_water <- opq(bbox = c(46.20, -16.60, 47.20, -15.80)) |>
    add_osm_feature(key = "natural", value = "water") |> osmdata_sf()
  osm_water_maro <- list(lines = q_riv$osm_lines, polys = q_water$osm_polygons,
                         mpolys = q_water$osm_multipolygons)
  write_rds(osm_water_maro, osm_cache_maro)
}

# Administrative boundaries
adm2 <- st_read(
  "data/OCHA_BNGRC admin boundaries/mdg_admbnda_adm2_BNGRC_OCHA_20181031.shp",
  quiet = TRUE
) |> st_make_valid()
madagascar <- st_union(adm2)

# All observatory communes
obs_communes_all <- st_read("data/Observatoires_ROS_communes_COD_v4.gpkg",
                            quiet = TRUE) |>
  st_make_valid() |>
  mutate(OBS_CODE = as.integer(OBS_CODE))

maro_communes <- obs_communes_all |> filter(OBS_CODE == 3)
ala_communes  <- obs_communes_all |> filter(OBS_CODE == 21)
fara_communes <- obs_communes_all |> filter(OBS_CODE == 15)
tol_communes  <- obs_communes_all |> filter(OBS_CODE == 23)
fen_communes  <- obs_communes_all |> filter(OBS_CODE == 24)

# -- Fokontany coordinates (2025 GPS, Marovoay only) -------------------------
maro_fkt <- tibble(
  fokontany  = c("Bepako", "Madiromiongana", "Ampijoroa", "Maroala"),
  site_id    = c("031", "032", "033", "034"),
  lon        = c(46.658402, 46.749270, 46.477984, 46.539738),
  lat        = c(-16.163712, -16.086480, -16.233414, -16.228857),
  river_side = c("east", "east", "west", "west"),
  treated    = c(TRUE, TRUE, FALSE, FALSE)
) |>
  st_as_sf(coords = c("lon", "lat"), crs = 4326)
maro_fkt$dist_km <- round(as.numeric(
  st_distance(maro_fkt, st_union(ankarafantsika_pa))) / 1000, 1)

ala_fkt <- tibble(
  fokontany = c("Avaradrano", "Feramanga Atsimo", "Mangabe",
                "Analamiranga", "Maritampona", "Ambohidrony", "Ambatomanga"),
  lon = c(48.469958, 48.373383, 48.405244, 48.200017, 48.207639, 48.208611, 48.200647),
  lat = c(-17.805414, -17.844153, -17.879586, -17.572375, -17.603500, -17.729175, -17.741178)
) |> st_as_sf(coords = c("lon", "lat"), crs = 4326)
ala_fkt$dist_php <- round(as.numeric(st_distance(ala_fkt, st_union(alaotra_pa)))/1000, 1)
ala_fkt$dist_caz <- round(as.numeric(st_distance(ala_fkt, st_union(caz_pa)))/1000, 1)
ala_fkt$nearest_pa <- if_else(ala_fkt$dist_caz < ala_fkt$dist_php,
                               "Corridor Ankeniheny-Zahamena", "Lac Alaotra PHP")

# -- Observatory locations table -----------------------------------------------
obs_locations <- tibble(
  code = c("02", "03", "13", "15", "16", "21", "23", "24", "25", "41"),
  name = c("Antsirabe", "Marovoay", "Tsiroanomandidy", "Farafangana",
           "Ambovombe", "Alaotra", "Toliara North", "Fénérive East",
           "Mahanoro", "Itasy"),
  latitude  = c(-19.8667, -16.1000, -18.7714, -22.8167, -25.0333,
                -17.5833, -23.3893, -17.3833, -20.4000, -19.0500),
  longitude = c(47.0333, 46.6333, 46.0546, 47.8333, 46.0833,
                48.4167, 43.7761, 49.4167, 48.8000, 46.7354),
  role = case_when(
    code == "03" ~ "Marovoay (strict, 2003)",
    code == "21" ~ "Alaotra (multipurpose, 2008)",
    code %in% c("15","23","24") ~ "New treated (2006–2007)",
    .default = "Clean donor pool"
  )
) |>
  mutate(role = factor(role, levels = c("Marovoay (strict, 2003)",
                                        "Alaotra (multipurpose, 2008)",
                                        "New treated (2006–2007)",
                                        "Clean donor pool")))
```


# Introduction {#sec-intro}

[Chapter 12](12_strict_vs_multipurpose.html) estimated the income effects of two contrasting conservation governance models in rural Madagascar using the generalized synthetic control method [@Xu2017]: strict enforcement at Ankarafantsika National Park (IUCN II, treated from 2003) produced a persistent negative income effect, while participatory conservation at Lac Alaotra (IUCN V, treated from 2008) showed no statistically significant effect. Both findings rest on a two-case comparison, which limits the external validity of the strict/multipurpose contrast.

This chapter extends the analysis to three additional rural observatories that, as documented in the [donor pool validity audit](13_donor_pool_pa_exposure.html) (Chapter 13), are themselves adjacent to PAs created during the 1999–2014 study period:

- **Farafangana (15)**: adjacent to the Corridor Forestier Ambositra-Vondrozo (IUCN V, September 2006), a large highland conservation corridor managed under a light-touch regulatory model. Survey coverage: 1999–2008 (3 post-treatment years).
- **Toliara North (23)**: adjacent to Amoron'i Onilahy (IUCN V, January 2007), Mikea National Park (IUCN II, April 2007), and Ranobe PK 32 (IUCN V, December 2008) — a cascade of coastal and dry-forest protections. Survey coverage: 1999–2014 (8 post-treatment years).
- **Fénérive East (24)**: adjacent to the Réserve de Tampolo (IUCN IV, August 2007), a small research forest managed by a university institution. Survey coverage: 1999–2014 (8 post-treatment years).

These three observatories move from being liabilities in the donor pool validity analysis to informative treated units in this extended comparison. Together with Marovoay and Alaotra, they enable a five-case examination of how conservation governance type, enforcement intensity, and PA size shape household income outcomes.

A critical methodological change from Chapter 12: the donor pool here is restricted to the **five clean observatories** identified in Chapter 13 (Antsirabe, Tsiroanomandidy, Ambovombe, Mahanoro, Itasy). This excludes the four contaminated donors used in Chapter 12. The resulting donor pool is smaller but free of PA exposure during the study period. Marovoay and Alaotra estimates reported in this chapter will therefore differ slightly from Chapter 12.


# Study Sites {#sec-sites}

## Ankarafantsika National Park and the Marovoay Observatory {#sec-maro}

See [Chapter 12, §2.1](12_strict_vs_multipurpose.html#sec-sites) for a full description. Briefly: Ankarafantsika National Park (136,500 ha, IUCN II) was reclassified in 2002 from a loosely enforced forest reserve to a strictly patrolled national park. The Betsiboka River separates the two **treated** *fokontany* (Bepako and Madiromiongana, east bank) from two **control** communities (Ampijoroa and Maroala, west bank). We treat from 2003 — the first full survey year after Decree 2002-798. Treatment assignment is validated by a 2025 resurvey showing sharply higher PA interaction rates on the east bank.

```{r fig-map-maro}
#| label: fig-map-maro
#| fig-cap: "Marovoay observatory and Ankarafantsika NP. East-bank fokontany (treated, orange) and west-bank fokontany (control, grey) are separated by the Betsiboka River."
#| fig-height: 7
#| fig-width: 8

ank_bb  <- st_bbox(ankarafantsika_pa)
fkt_bb  <- st_bbox(maro_fkt)
maro_bbox <- st_bbox(c(
  xmin = min(ank_bb["xmin"], fkt_bb["xmin"]) - 0.08,
  ymin = min(ank_bb["ymin"], fkt_bb["ymin"]) - 0.08,
  xmax = max(ank_bb["xmax"], fkt_bb["xmax"]) + 0.08,
  ymax = max(ank_bb["ymax"], fkt_bb["ymax"]) + 0.08
), crs = 4326)

betsiboka_named <- osm_water_maro$lines |> filter(name == "Betsiboka")
betsiboka_buf <- betsiboka_named |>
  st_transform(32738) |> st_union() |> st_buffer(1000) |> st_transform(4326)

water_lines  <- tryCatch(st_crop(betsiboka_named, maro_bbox), error = function(e) NULL)
water_polys  <- tryCatch({
  wp <- st_crop(osm_water_maro$polys, maro_bbox)
  wp[st_intersects(wp, betsiboka_buf, sparse = FALSE)[,1],]
}, error = function(e) NULL)
water_mpolys <- tryCatch({
  wm <- st_crop(osm_water_maro$mpolys, maro_bbox)
  wm[st_intersects(wm, betsiboka_buf, sparse = FALSE)[,1],]
}, error = function(e) NULL)

WATER_COL <- "#1976D2"
ank_centroid <- st_coordinates(st_centroid(st_union(ankarafantsika_pa)))

p <- ggplot() +
  geom_sf(data = ankarafantsika_pa, fill = "#A5D6A7", color = "#388E3C",
          alpha = 0.4, linewidth = 0.6)
if (!is.null(water_polys)  && nrow(water_polys)  > 0)
  p <- p + geom_sf(data = water_polys,  fill = WATER_COL, color = WATER_COL, linewidth = 0.1)
if (!is.null(water_mpolys) && nrow(water_mpolys) > 0)
  p <- p + geom_sf(data = water_mpolys, fill = WATER_COL, color = WATER_COL, linewidth = 0.1)
if (!is.null(water_lines)  && nrow(water_lines)  > 0)
  p <- p + geom_sf(data = water_lines, color = WATER_COL, linewidth = 0.7)

maro_fkt_lab <- maro_fkt |>
  mutate(lon = st_coordinates(geometry)[,1], lat = st_coordinates(geometry)[,2])

p + annotate("text", x = ank_centroid[1], y = ank_centroid[2],
             label = "Ankarafantsika\nNational Park",
             color = "#1B5E20", fontface = "bold.italic", size = 4, alpha = 0.8) +
  geom_sf(data = maro_fkt, aes(color = river_side, shape = river_side),
          size = 4, stroke = 1.2) +
  geom_label_repel(data = maro_fkt_lab,
                   aes(x = lon, y = lat,
                       label = paste0(fokontany, "\n(", dist_km, " km)"),
                       color = river_side),
                   size = 2.8, fontface = "bold", box.padding = 0.8,
                   min.segment.length = 0, show.legend = FALSE) +
  scale_color_manual(values = c("east" = "#E65100", "west" = "#616161"),
                     labels = c("east" = "East bank (treated)", "west" = "West bank (control)"),
                     name = NULL) +
  scale_shape_manual(values = c("east" = 17, "west" = 16),
                     labels = c("east" = "East bank (treated)", "west" = "West bank (control)"),
                     name = NULL) +
  annotation_scale(location = "bl", width_hint = 0.25) +
  labs(title = "Marovoay: Ankarafantsika NP") +
  coord_sf(xlim = c(maro_bbox["xmin"], maro_bbox["xmax"]),
           ylim = c(maro_bbox["ymin"], maro_bbox["ymax"])) +
  theme_minimal(base_size = 11) + theme(legend.position = "bottom")
```

## Lac Alaotra Protected Landscape and the Alaotra Observatory {#sec-alaotra}

See [Chapter 12, §2.2](12_strict_vs_multipurpose.html#sec-alaotra) for a full description. Briefly: the Lac Alaotra PHP (425 km², IUCN V) enforces CBNRM from 2008 under Durrell's management. The entire observatory is aggregated into a single treated unit because all seven *fokontany* depend on the lake-marsh ecosystem with no access-barrier equivalent to Marovoay's Betsiboka River. A 2025 resurvey confirms uniform PA interaction across sites (37–54% visitation), validating the observatory-level treatment assignment.

## Corridor Forestier Ambositra-Vondrozo and the Farafangana Observatory {#sec-fara}

The Corridor Forestier Ambositra-Vondrozo (CFAV; IUCN V) is a ~300 km Protected Landscape running along the eastern highland escarpment. A temporary protection decree (*arrêté interministériel*) was issued on **15 September 2006** (WDPA 555697881), designating the CFAV as part of the post-2003 SAPM expansion. Category V status implies land-use regulation at the forest margin — primarily restrictions on slash-and-burn agriculture (*tavy*) and charcoal production in buffer zones — rather than exclusionary enforcement. Communities retain legal access to traditional agricultural land.

The Farafangana observatory lies on the southeastern coastal plain, separated from the highland interior by an escarpment. Its nearest commune (Mahatsinjo) borders the CFAV at approximately 14 km. A complication: Manombo Special Reserve (IUCN IV, 1962) lies approximately 12 km from a second Farafangana commune. Manombo predates the study period and its boundaries and management regime were unchanged across 1999–2014; it constitutes a constant background condition absorbed into the unit fixed effect rather than a within-period treatment event. The estimated ATT for Farafangana therefore identifies the **additional** income effect of the 2006 CFAV decree, on top of whatever steady-state condition Manombo created.

**Treatment identification**: All Farafangana survey sites are treated at observatory level from 2006. No within-observatory split is possible; the observatory is treated as a single unit ("FARA").

**Caution on statistical power**: the Farafangana survey ends in 2008, yielding only **3 post-treatment years** (2006–2008). The gsynth ATT will be estimated with very wide confidence intervals and should be treated as illustrative rather than conclusive.

```{r fig-map-fara}
#| label: fig-map-fara
#| fig-cap: "Farafangana observatory (commune polygons) and nearby protected areas. The Corridor Forestier Ambositra-Vondrozo (CFAV, treatment, orange) and Manombo SR (pre-existing, yellow) are shown. Distances are centroid-to-boundary."
#| fig-height: 7
#| fig-width: 7

fara_pa_names <- c("Corridor Forestier Ambositra Vondrozo", "Manombo")
fara_pas <- dyn_study |>
  filter(str_detect(NAME, paste(fara_pa_names, collapse = "|")))

fara_bbox <- st_bbox(c(
  xmin = 46.8, ymin = -23.5, xmax = 48.8, ymax = -22.0
), crs = 4326)

adm2_fara <- tryCatch(st_crop(adm2, fara_bbox), error = function(e) NULL)
fara_pas_c <- tryCatch(st_crop(fara_pas, fara_bbox), error = function(e) NULL)
fara_com_c <- tryCatch(st_crop(fara_communes, fara_bbox), error = function(e) NULL)

fara_pas_c <- fara_pas_c |>
  mutate(pa_type = if_else(str_detect(NAME, "Corridor"), "Treatment PA (CFAV, 2006)",
                           "Pre-existing PA (Manombo, 1962)"))

ggplot() +
  geom_sf(data = adm2_fara, fill = "grey96", color = "grey75", linewidth = 0.2) +
  geom_sf(data = fara_pas_c, aes(fill = pa_type), color = NA, alpha = 0.4) +
  geom_sf(data = fara_pas_c, aes(color = pa_type), fill = NA, linewidth = 0.5) +
  geom_sf(data = fara_com_c, fill = "#E8F5E9", color = "grey40",
          linewidth = 0.3, alpha = 0.7) +
  geom_sf_label(data = st_centroid(st_union(fara_com_c |> summarise(.groups = "drop"))),
                label = "Farafangana\nobservatory",
                size = 2.5, fill = "white", label.size = 0.15) +
  scale_fill_manual(values = c("Treatment PA (CFAV, 2006)" = "#EF9A9A",
                               "Pre-existing PA (Manombo, 1962)" = "#FFF59D"),
                    aesthetics = c("fill","color"), name = NULL) +
  annotation_scale(location = "bl", width_hint = 0.2) +
  labs(title = "Farafangana & Nearby PAs") +
  coord_sf(xlim = c(fara_bbox["xmin"], fara_bbox["xmax"]),
           ylim = c(fara_bbox["ymin"], fara_bbox["ymax"])) +
  theme_minimal(base_size = 11) + theme(legend.position = "bottom")
```

## Coastal PAs and the Toliara North Observatory {#sec-tol}

Toliara North (observatory 23) sits on Madagascar's arid southwest coast, a landscape of spiny thicket and dry deciduous forest. Three PAs became operative in rapid succession from January 2007:

1. **Amoron'i Onilahy** (IUCN V, temp decree 17 January 2007, ~12 km away): a riparian and coastal landscape PA covering the Onilahy river estuary and hinterland.
2. **Mikea National Park** (IUCN II, temp decree 13 April 2007, ~16 km away): a strict reserve protecting the unique Mikea dry spiny-thicket forest.
3. **Ranobe PK 32** (IUCN V, temp decree 2 December 2008, ~1.4 km away): a large terrestrial PA (1,685 km²) bordering the observatory at extremely close range.

Treatment year is defined as **2007** — the first full survey year following the initial January 2007 decree. The December 2008 Ranobe decree constitutes a secondary intensification within the post-treatment window. The dominant conservation character is **IUCN V / multipurpose**, though the presence of Mikea NP (IUCN II) at 16 km introduces a strict element.

```{r fig-map-tol}
#| label: fig-map-tol
#| fig-cap: "Toliara North observatory and nearby protected areas. Three PAs became operative during 2007–2008. Ranobe PK 32 (red, 2008) lies within 1.4 km of the nearest commune centroid."
#| fig-height: 7
#| fig-width: 7

tol_pa_names <- c("Ranobe PK 32", "Amoron'i Onilahy", "Mikea")
tol_pas <- dyn_study |>
  filter(str_detect(NAME, paste(tol_pa_names, collapse = "|"))) |>
  mutate(pa_label = case_when(
    str_detect(NAME, "Ranobe")    ~ "Ranobe PK 32 (V, 2008)",
    str_detect(NAME, "Amoron")    ~ "Amoron'i Onilahy (V, 2007)",
    str_detect(NAME, "Mikea")     ~ "Mikea NP (II, 2007)"
  ))

tol_bbox <- st_bbox(c(
  xmin = 43.2, ymin = -24.0, xmax = 44.8, ymax = -22.8
), crs = 4326)

adm2_tol  <- tryCatch(st_crop(adm2,         tol_bbox), error = function(e) NULL)
tol_pas_c <- tryCatch(st_crop(tol_pas,       tol_bbox), error = function(e) NULL)
tol_com_c <- tryCatch(st_crop(tol_communes,  tol_bbox), error = function(e) NULL)

tol_pal <- c("Ranobe PK 32 (V, 2008)"      = "#EF5350",
             "Amoron'i Onilahy (V, 2007)"  = "#FF8A65",
             "Mikea NP (II, 2007)"         = "#E53935")

ggplot() +
  geom_sf(data = adm2_tol, fill = "grey96", color = "grey75", linewidth = 0.2) +
  geom_sf(data = tol_pas_c, aes(fill = pa_label), color = NA, alpha = 0.35) +
  geom_sf(data = tol_pas_c, aes(color = pa_label), fill = NA, linewidth = 0.5) +
  geom_sf(data = tol_com_c, fill = "#E3F2FD", color = "grey40",
          linewidth = 0.3, alpha = 0.8) +
  scale_fill_manual(values = tol_pal, aesthetics = c("fill","color"), name = NULL) +
  annotation_scale(location = "bl", width_hint = 0.2) +
  labs(title = "Toliara North & Nearby PAs (2007–2008)") +
  coord_sf(xlim = c(tol_bbox["xmin"], tol_bbox["xmax"]),
           ylim = c(tol_bbox["ymin"], tol_bbox["ymax"])) +
  theme_minimal(base_size = 11) + theme(legend.position = "bottom")
```

## Réserve de Tampolo and the Fénérive East Observatory {#sec-fen}

The Réserve de Tampolo (IUCN IV, ~7.3 km²) is a small coastal littoral forest reserve on Madagascar's northeastern coast, managed by ESSA-Forêts (École Supérieure des Sciences Agronomiques, Université d'Antananarivo) as a research and teaching forest. A temporary protection decree was issued on **20 August 2007** (WDPA 354011); the WDPA STATUS_YR records 2006, but the dynamic WDPA cross-reference to CNLEGIS assigns the operative date as 2007.

The reserve's primary function is scientific access and conservation research, not community resource exclusion. Its small footprint and institutional character suggest **low treatment intensity**: enforcement of extractive restrictions is less systematic than at a national park or major CBNRM programme. Treatment year is **2007** (within-year decree, assigned to the same calendar year by convention). The Fénérive East survey has 8 pre-treatment years (1999–2006) and 8 post-treatment years (2007–2014), providing the most symmetric panel of the three new observatories.

```{r fig-map-fen}
#| label: fig-map-fen
#| fig-cap: "Fénérive East observatory and the Réserve de Tampolo (IUCN IV, 2007). The reserve is small (~7.3 km²) and located on the northeastern littoral coast."
#| fig-height: 7
#| fig-width: 7

fen_pas <- dyn_study |>
  filter(str_detect(NAME, "Tampolo"))

fen_bbox <- st_bbox(c(
  xmin = 49.1, ymin = -17.7, xmax = 49.9, ymax = -16.9
), crs = 4326)

adm2_fen  <- tryCatch(st_crop(adm2,         fen_bbox), error = function(e) NULL)
fen_pas_c <- tryCatch(st_crop(fen_pas,       fen_bbox), error = function(e) NULL)
fen_com_c <- tryCatch(st_crop(fen_communes,  fen_bbox), error = function(e) NULL)

ggplot() +
  geom_sf(data = adm2_fen, fill = "grey96", color = "grey75", linewidth = 0.2) +
  geom_sf(data = fen_com_c, fill = "#FFF8E1", color = "grey40",
          linewidth = 0.3, alpha = 0.8) +
  geom_sf(data = fen_pas_c, fill = "#CE93D8", color = "#6A1B9A",
          alpha = 0.5, linewidth = 0.7) +
  annotate("text", x = 49.355, y = -17.37,
           label = "Réserve de\nTampolo (IV)", color = "#4A148C",
           fontface = "bold.italic", size = 3) +
  annotation_scale(location = "bl", width_hint = 0.2) +
  labs(title = "Fénérive East & Réserve de Tampolo (2007)") +
  coord_sf(xlim = c(fen_bbox["xmin"], fen_bbox["xmax"]),
           ylim = c(fen_bbox["ymin"], fen_bbox["ymax"])) +
  theme_minimal(base_size = 11)
```


# Data and Identification {#sec-data}

## The Rural Observatory Panel

We use the same outcome variable as Chapter 12: log OECD-equivalised total household income, winsorised at the 1st–99th percentiles. Panel years are 1999–2014 for the full analysis, with the Farafangana model restricted to 1999–2008 (its survey window).

```{r tbl-panel-five}
#| label: tbl-panel-five
#| tbl-cap: "Panel design for all five treated observatories. Treatment year and PA character determine the pre/post split. Farafangana's short post-treatment window is highlighted."

tribble(
  ~Observatory,        ~`PA (primary)`,                    ~`IUCN`, ~`Type`,           ~`Treat. yr`, ~`Pre`, ~`Post`, ~`Survey`,
  "Marovoay (03)",     "Ankarafantsika NP",                "II",    "Strict",           2003L,        4L,     11L,     "1999–2014",
  "Alaotra (21)",      "Lac Alaotra PHP",                  "V",     "Multipurpose",     2008L,        9L,     6L,      "1999–2014",
  "Farafangana (15)",  "Corridor Ambositra-Vondrozo",      "V",     "Multipurpose",     2006L,        7L,     3L,      "1999–2008 ⚠",
  "Toliara North (23)","Amoron'i Onilahy + Mikea + Ranobe","V/II",  "Mixed (V dom.)",   2007L,        8L,     8L,      "1999–2014",
  "Fénérive East (24)","Réserve de Tampolo",               "IV",    "Low intensity",    2007L,        8L,     8L,      "1999–2014"
) |>
  gt() |>
  tab_header(
    title    = md("**Five Treated Observatories: Design Summary**"),
    subtitle = "⚠ Farafangana: only 3 post-treatment years — estimates are indicative only"
  ) |>
  tab_style(
    style = list(cell_text(weight = "bold"), cell_fill(color = "#FFCDD2")),
    locations = cells_body(rows = `Observatory` == "Marovoay (03)")
  ) |>
  tab_style(
    style = list(cell_text(weight = "bold"), cell_fill(color = "#BBDEFB")),
    locations = cells_body(rows = `Observatory` == "Alaotra (21)")
  ) |>
  tab_style(
    style = cell_fill(color = "#FFF8E1"),
    locations = cells_body(rows = grepl("Farafangana", Observatory))
  ) |>
  tab_style(
    style = cell_fill(color = "#FBE9E7"),
    locations = cells_body(rows = grepl("Toliara", Observatory))
  ) |>
  tab_style(
    style = cell_fill(color = "#F3E5F5"),
    locations = cells_body(rows = grepl("Fénérive", Observatory))
  )
```

## Donor Pool

**This chapter uses a restricted donor pool** of five clean observatories (Antsirabe 02, Tsiroanomandidy 13, Ambovombe 16, Mahanoro 25, Itasy 41), as defined in [Chapter 13](13_donor_pool_pa_exposure.html). The four contaminated donors used in Chapter 12 (Antsohihy, Farafangana, Toliara North, Fénérive East) are now treated units here. Antsohihy (12) remains excluded from both the treated group and the donor pool because it is exposed to the Corridor Bongolava without sufficient post-treatment survey coverage.

In addition to the five clean observatory sites, each model includes within-observatory control sites where applicable: Marovoay west-bank *fokontany* (033, 034) serve as additional donors for the Marovoay model.

```{r fig-obs-map}
#| label: fig-obs-map
#| fig-cap: "Observatory locations for the extended five-observatory analysis. Colours distinguish the two original treated sites (Ch. 12), three new treated sites (this chapter), and the five clean donors."
#| fig-height: 7
#| fig-width: 6

ggplot() +
  geom_sf(data = madagascar, fill = "grey95", color = "grey60", linewidth = 0.3) +
  geom_point(data = obs_locations,
             aes(x = longitude, y = latitude, color = role, shape = role),
             size = 3.5, stroke = 1) +
  geom_label_repel(data = obs_locations,
                   aes(x = longitude, y = latitude,
                       label = paste0(name, " (", code, ")"),
                       color = role),
                   size = 2.3, fontface = "bold", box.padding = 0.5,
                   min.segment.length = 0, max.overlaps = 20,
                   show.legend = FALSE) +
  scale_color_manual(
    values = c("Marovoay (strict, 2003)"    = "#D32F2F",
               "Alaotra (multipurpose, 2008)" = "#1565C0",
               "New treated (2006–2007)"    = "#E65100",
               "Clean donor pool"           = "grey45"),
    name = NULL) +
  scale_shape_manual(
    values = c("Marovoay (strict, 2003)"    = 17,
               "Alaotra (multipurpose, 2008)" = 17,
               "New treated (2006–2007)"    = 15,
               "Clean donor pool"           = 16),
    name = NULL) +
  labs(title = "Extended Five-Observatory Analysis") +
  theme_void(base_size = 11) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7.5))
```


# Empirical Strategy {#sec-method}

The identification strategy and estimator are identical to Chapter 12. We estimate the generalized synthetic control (gsynth) ATT separately for each of the five treated observatories:

$$
Y_{it} = \alpha_i + \xi_t + \boldsymbol{\lambda}_i' \boldsymbol{f}_t + D_{it} \cdot \delta_{it} + \varepsilon_{it}
$$

Each model specifies $D_{it} = 1$ only for the target observatory's treated unit(s) from its treatment year onward; all other units (including the other four treated observatories) carry $D_{it} = 0$ throughout and thus serve as additional factors for counterfactual construction. This within-model use of other treated observatories as donors is consistent with the interactive FE framework so long as treatment timing differs across observatories [@Xu2017]. The number of latent factors $r$ is selected by leave-one-out cross-validation; inference is parametric bootstrap with 200 replications.

Treatment unit definitions follow Chapter 12: Marovoay is site-level (east-bank *fokontany* only); all other observatories are aggregated to a single observatory-level unit ("ALA", "FARA", "TOL", "FEN"). The statistical power ceiling discussed in Chapter 12 applies here with equal force for the observatory-level models: each has exactly one treated unit, and the minimum permutation p-value is $\approx 1/N_{donors}$.


# Results {#sec-results}

## Panel Construction

```{r gsynth-panels}
#| label: gsynth-panels

# Base panel: all 8 observatories (5 treated + 5 clean donors) ----------------
panel_gs_hh <- household_consolidated |>
  filter(
    obs %in% ALL_OBS_14,
    year %in% 1999:2014,
    !(obs == "03" & year == 2005)  # Marovoay missing 2005
  ) |>
  mk_income_ext() |>
  mutate(
    revtot_eq   = revtot / oecd_equiv,
    revtot_w    = pmin(revtot_eq, quantile(revtot_eq, 0.99, na.rm = TRUE)),
    ln_revtot_w = log(pmax(revtot_w, 0) + 1)
  )

# Per-model panels (each sets treatment for ONLY its observatory) -------------

# Marovoay: site-level, east-bank treated
maro_gs <- panel_gs_hh |>
  mutate(treatment = as.integer(site_id %in% c("031","032") & year >= TREAT_YEAR_MARO)) |>
  summarize(ln_revtot_w = mean(ln_revtot_w, na.rm = TRUE),
            n_hh = n(), treatment = max(treatment),
            .by = c(site_id, year))

# Alaotra: observatory-level
ala_gs <- panel_gs_hh |>
  mutate(unit_id   = if_else(obs == "21", "ALA", site_id),
         treatment = as.integer(obs == "21" & year >= TREAT_YEAR_ALA)) |>
  summarize(ln_revtot_w = mean(ln_revtot_w, na.rm = TRUE),
            n_hh = n(), treatment = max(treatment),
            .by = c(unit_id, year))

# Farafangana: observatory-level, restricted to 1999-2008
fara_gs <- panel_gs_hh |>
  filter(year <= 2008) |>
  mutate(unit_id   = if_else(obs == "15", "FARA", site_id),
         treatment = as.integer(obs == "15" & year >= TREAT_YEAR_FARA)) |>
  summarize(ln_revtot_w = mean(ln_revtot_w, na.rm = TRUE),
            n_hh = n(), treatment = max(treatment),
            .by = c(unit_id, year))

# Toliara North: observatory-level
tol_gs <- panel_gs_hh |>
  mutate(unit_id   = if_else(obs == "23", "TOL", site_id),
         treatment = as.integer(obs == "23" & year >= TREAT_YEAR_TOL)) |>
  summarize(ln_revtot_w = mean(ln_revtot_w, na.rm = TRUE),
            n_hh = n(), treatment = max(treatment),
            .by = c(unit_id, year))

# Fénérive East: observatory-level
fen_gs <- panel_gs_hh |>
  mutate(unit_id   = if_else(obs == "24", "FEN", site_id),
         treatment = as.integer(obs == "24" & year >= TREAT_YEAR_FEN)) |>
  summarize(ln_revtot_w = mean(ln_revtot_w, na.rm = TRUE),
            n_hh = n(), treatment = max(treatment),
            .by = c(unit_id, year))

n_donors_maro <- maro_gs |> filter(all(treatment == 0), .by = site_id) |>
  distinct(site_id) |> nrow()
n_donors_ala  <- ala_gs  |> filter(all(treatment == 0), .by = unit_id) |>
  distinct(unit_id) |> nrow()
n_donors_fara <- fara_gs |> filter(all(treatment == 0), .by = unit_id) |>
  distinct(unit_id) |> nrow()
n_donors_tol  <- tol_gs  |> filter(all(treatment == 0), .by = unit_id) |>
  distinct(unit_id) |> nrow()
n_donors_fen  <- fen_gs  |> filter(all(treatment == 0), .by = unit_id) |>
  distinct(unit_id) |> nrow()
```


## Ankarafantsika: East-Bank Effect {#sec-maro-results}

```{r gsynth-maro}
#| label: gsynth-maro

gs_maro_cache <- "data/paper14_gsynth_maro_oecd.rds"
need_rerun <- !file.exists(gs_maro_cache) ||
  is.null(read_rds(gs_maro_cache)$est_avg)

if (need_rerun) {
  fit <- gsynth(ln_revtot_w ~ treatment, data = maro_gs,
                index = c("site_id", "year"), force = "two-way",
                CV = TRUE, r = c(0, 5), se = TRUE,
                inference = "parametric", nboots = 200,
                parallel = FALSE, min.T0 = 3, seed = 42)
  write_rds(save_gsynth(fit), gs_maro_cache)
}
gs_maro <- read_rds(gs_maro_cache)
```

```{r fig-maro-att}
#| label: fig-maro-att
#| fig-cap: "Ankarafantsika: gsynth ATT path (east-bank vs counterfactual). Outcome: log OECD-equivalised income. Red band = 95% parametric bootstrap CI."
#| fig-height: 4.5
#| fig-width: 8

att_maro_df <- tibble(
  rel_time = as.numeric(rownames(gs_maro$est_att)),
  att      = gs_maro$est_att[,"ATT"],
  ci_lo    = gs_maro$est_att[,"CI.lower"],
  ci_hi    = gs_maro$est_att[,"CI.upper"]
)

ggplot(att_maro_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#D32F2F", alpha = 0.15) +
  geom_line(color = "#D32F2F", linewidth = 0.8) +
  geom_point(color = "#D32F2F", size = 2.5) +
  labs(x = "Years relative to treatment (2003)",
       y = "ATT (log equiv. income)",
       title = "Ankarafantsika (strict, IUCN II)",
       subtitle = paste0(gs_maro$N_co, " donor sites; r* = ", gs_maro$r_cv)) +
  theme_minimal(base_size = 12)
```

```{r maro-text}
#| output: asis

att_m <- gs_maro$att_avg
pct_m <- round((exp(att_m) - 1) * 100, 1)
se_m  <- gs_maro$est_avg[1, "S.E."]
ci_lo_m <- gs_maro$est_avg[1, "CI.lower"]
ci_hi_m <- gs_maro$est_avg[1, "CI.upper"]
pval_m  <- gs_maro$est_avg[1, "p.value"]

cat(sprintf(
  "Average post-treatment ATT: **%.3f log points** (SE = %.3f; 95%% CI [%.3f, %.3f]; p = %.3f), corresponding to approximately **%+.1f%%** in equivalised income. r* = %d factor(s); %d donor sites.\n\n",
  att_m, se_m, ci_lo_m, ci_hi_m, pval_m, pct_m, gs_maro$r_cv, gs_maro$N_co
))
```


## Lac Alaotra: Observatory-Level Effect {#sec-ala-results}

```{r gsynth-ala}
#| label: gsynth-ala

gs_ala_cache <- "data/paper14_gsynth_ala_oecd.rds"
need_rerun <- !file.exists(gs_ala_cache) ||
  is.null(read_rds(gs_ala_cache)$est_avg)

if (need_rerun) {
  fit <- gsynth(ln_revtot_w ~ treatment, data = ala_gs,
                index = c("unit_id", "year"), force = "two-way",
                CV = TRUE, r = c(0, 5), se = TRUE,
                inference = "parametric", nboots = 200,
                parallel = FALSE, min.T0 = 3, seed = 42)
  write_rds(save_gsynth(fit), gs_ala_cache)
}
gs_ala <- read_rds(gs_ala_cache)
```

```{r fig-ala-att}
#| label: fig-ala-att
#| fig-cap: "Lac Alaotra: gsynth ATT path. Blue band = 95% CI."
#| fig-height: 4.5
#| fig-width: 8

att_ala_df <- tibble(
  rel_time = as.numeric(rownames(gs_ala$est_att)),
  att      = gs_ala$est_att[,"ATT"],
  ci_lo    = gs_ala$est_att[,"CI.lower"],
  ci_hi    = gs_ala$est_att[,"CI.upper"]
)

ggplot(att_ala_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#1565C0", alpha = 0.15) +
  geom_line(color = "#1565C0", linewidth = 0.8) +
  geom_point(color = "#1565C0", size = 2.5) +
  labs(x = "Years relative to treatment (2008)",
       y = "ATT (log equiv. income)",
       title = "Lac Alaotra (multipurpose, IUCN V)",
       subtitle = paste0(gs_ala$N_co, " donor sites; r* = ", gs_ala$r_cv)) +
  theme_minimal(base_size = 12)
```

```{r ala-text}
#| output: asis

att_a  <- gs_ala$att_avg
pct_a  <- round((exp(att_a) - 1) * 100, 1)
se_a   <- gs_ala$est_avg[1,"S.E."]
ci_lo_a <- gs_ala$est_avg[1,"CI.lower"]
ci_hi_a <- gs_ala$est_avg[1,"CI.upper"]
pval_a  <- gs_ala$est_avg[1,"p.value"]

cat(sprintf(
  "Average post-treatment ATT: **%.3f log points** (SE = %.3f; 95%% CI [%.3f, %.3f]; p = %.3f), corresponding to approximately **%+.1f%%**. The CI spans zero; the effect is not statistically significant at the 5%% level. r* = %d; %d donor sites.\n\n",
  att_a, se_a, ci_lo_a, ci_hi_a, pval_a, pct_a, gs_ala$r_cv, gs_ala$N_co
))
```


## Farafangana: Corridor Forestier Effect {#sec-fara-results}

::: {.callout-caution}
Farafangana has only 3 post-treatment survey years (2006–2008). gsynth estimates are presented for completeness but should be treated with extreme caution. Pre-treatment fit is assessed over 7 years; post-treatment precision is insufficient for firm conclusions.
:::

```{r gsynth-fara}
#| label: gsynth-fara

gs_fara_cache <- "data/paper14_gsynth_fara_oecd.rds"
need_rerun <- !file.exists(gs_fara_cache) ||
  is.null(read_rds(gs_fara_cache)$est_avg)

if (need_rerun) {
  fit <- gsynth(ln_revtot_w ~ treatment, data = fara_gs,
                index = c("unit_id", "year"), force = "two-way",
                CV = TRUE, r = c(0, 3), se = TRUE,
                inference = "parametric", nboots = 200,
                parallel = FALSE, min.T0 = 3, seed = 42)
  write_rds(save_gsynth(fit), gs_fara_cache)
}
gs_fara <- read_rds(gs_fara_cache)
```

```{r fig-fara-att}
#| label: fig-fara-att
#| fig-cap: "Farafangana: gsynth ATT path (1999–2008). Treatment onset 2006. Very wide CIs reflect only 3 post-treatment observations. Green band = 95% CI."
#| fig-height: 4.5
#| fig-width: 8

att_fara_df <- tibble(
  rel_time = as.numeric(rownames(gs_fara$est_att)),
  att      = gs_fara$est_att[,"ATT"],
  ci_lo    = gs_fara$est_att[,"CI.lower"],
  ci_hi    = gs_fara$est_att[,"CI.upper"]
)

ggplot(att_fara_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#2E7D32", alpha = 0.15) +
  geom_line(color = "#2E7D32", linewidth = 0.8) +
  geom_point(color = "#2E7D32", size = 2.5) +
  labs(x = "Years relative to treatment (2006)",
       y = "ATT (log equiv. income)",
       title = "Farafangana (multipurpose, IUCN V) — INDICATIVE ONLY",
       subtitle = paste0(gs_fara$N_co, " donor sites; r* = ", gs_fara$r_cv,
                         " | Only 3 post-treatment years")) +
  theme_minimal(base_size = 12)
```

```{r fara-text}
#| output: asis

att_f  <- gs_fara$att_avg
pct_f  <- round((exp(att_f) - 1) * 100, 1)
se_f   <- gs_fara$est_avg[1,"S.E."]
ci_lo_f <- gs_fara$est_avg[1,"CI.lower"]
ci_hi_f <- gs_fara$est_avg[1,"CI.upper"]
pval_f  <- gs_fara$est_avg[1,"p.value"]

cat(sprintf(
  "Point estimate: **%.3f log points** (SE = %.3f; 95%% CI [%.3f, %.3f]; p = %.3f), corresponding to approximately **%+.1f%%**. With only 3 post-treatment observations, this estimate has low reliability and should not be interpreted independently. r* = %d; %d donor sites.\n\n",
  att_f, se_f, ci_lo_f, ci_hi_f, pval_f, pct_f, gs_fara$r_cv, gs_fara$N_co
))
```


## Toliara North: Coastal PA Cascade Effect {#sec-tol-results}

```{r gsynth-tol}
#| label: gsynth-tol

gs_tol_cache <- "data/paper14_gsynth_tol_oecd.rds"
need_rerun <- !file.exists(gs_tol_cache) ||
  is.null(read_rds(gs_tol_cache)$est_avg)

if (need_rerun) {
  fit <- gsynth(ln_revtot_w ~ treatment, data = tol_gs,
                index = c("unit_id", "year"), force = "two-way",
                CV = TRUE, r = c(0, 5), se = TRUE,
                inference = "parametric", nboots = 200,
                parallel = FALSE, min.T0 = 3, seed = 42)
  write_rds(save_gsynth(fit), gs_tol_cache)
}
gs_tol <- read_rds(gs_tol_cache)
```

```{r fig-tol-att}
#| label: fig-tol-att
#| fig-cap: "Toliara North: gsynth ATT path. Treatment onset 2007 (Amoron'i Onilahy decree). Note: Ranobe PK 32 (2008) constitutes a secondary PA creation within the post-treatment window. Orange band = 95% CI."
#| fig-height: 4.5
#| fig-width: 8

att_tol_df <- tibble(
  rel_time = as.numeric(rownames(gs_tol$est_att)),
  att      = gs_tol$est_att[,"ATT"],
  ci_lo    = gs_tol$est_att[,"CI.lower"],
  ci_hi    = gs_tol$est_att[,"CI.upper"]
)

ggplot(att_tol_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#E65100", alpha = 0.15) +
  geom_line(color = "#E65100", linewidth = 0.8) +
  geom_point(color = "#E65100", size = 2.5) +
  annotate("vline", xintercept = 1, linetype = "dotted",
           color = "#E65100", linewidth = 0.6) +
  annotate("text", x = 1.1, y = max(att_tol_df$ci_hi, na.rm = TRUE) * 0.85,
           label = "Ranobe PK 32\n(Dec 2008)", color = "#E65100",
           hjust = 0, size = 2.8) +
  labs(x = "Years relative to treatment (2007)",
       y = "ATT (log equiv. income)",
       title = "Toliara North (mixed IUCN V/II)",
       subtitle = paste0(gs_tol$N_co, " donor sites; r* = ", gs_tol$r_cv)) +
  theme_minimal(base_size = 12)
```

```{r tol-text}
#| output: asis

att_t  <- gs_tol$att_avg
pct_t  <- round((exp(att_t) - 1) * 100, 1)
se_t   <- gs_tol$est_avg[1,"S.E."]
ci_lo_t <- gs_tol$est_avg[1,"CI.lower"]
ci_hi_t <- gs_tol$est_avg[1,"CI.upper"]
pval_t  <- gs_tol$est_avg[1,"p.value"]

cat(sprintf(
  "Average post-treatment ATT: **%.3f log points** (SE = %.3f; 95%% CI [%.3f, %.3f]; p = %.3f), corresponding to approximately **%+.1f%%**. r* = %d; %d donor sites.\n\n",
  att_t, se_t, ci_lo_t, ci_hi_t, pval_t, pct_t, gs_tol$r_cv, gs_tol$N_co
))
```


## Fénérive East: Low-Intensity Reserve Effect {#sec-fen-results}

```{r gsynth-fen}
#| label: gsynth-fen

gs_fen_cache <- "data/paper14_gsynth_fen_oecd.rds"
need_rerun <- !file.exists(gs_fen_cache) ||
  is.null(read_rds(gs_fen_cache)$est_avg)

if (need_rerun) {
  fit <- gsynth(ln_revtot_w ~ treatment, data = fen_gs,
                index = c("unit_id", "year"), force = "two-way",
                CV = TRUE, r = c(0, 5), se = TRUE,
                inference = "parametric", nboots = 200,
                parallel = FALSE, min.T0 = 3, seed = 42)
  write_rds(save_gsynth(fit), gs_fen_cache)
}
gs_fen <- read_rds(gs_fen_cache)
```

```{r fig-fen-att}
#| label: fig-fen-att
#| fig-cap: "Fénérive East: gsynth ATT path. Treatment onset 2007 (Réserve de Tampolo decree). Purple band = 95% CI."
#| fig-height: 4.5
#| fig-width: 8

att_fen_df <- tibble(
  rel_time = as.numeric(rownames(gs_fen$est_att)),
  att      = gs_fen$est_att[,"ATT"],
  ci_lo    = gs_fen$est_att[,"CI.lower"],
  ci_hi    = gs_fen$est_att[,"CI.upper"]
)

ggplot(att_fen_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#6A1B9A", alpha = 0.15) +
  geom_line(color = "#6A1B9A", linewidth = 0.8) +
  geom_point(color = "#6A1B9A", size = 2.5) +
  labs(x = "Years relative to treatment (2007)",
       y = "ATT (log equiv. income)",
       title = "Fénérive East (low-intensity, IUCN IV)",
       subtitle = paste0(gs_fen$N_co, " donor sites; r* = ", gs_fen$r_cv)) +
  theme_minimal(base_size = 12)
```

```{r fen-text}
#| output: asis

att_e  <- gs_fen$att_avg
pct_e  <- round((exp(att_e) - 1) * 100, 1)
se_e   <- gs_fen$est_avg[1,"S.E."]
ci_lo_e <- gs_fen$est_avg[1,"CI.lower"]
ci_hi_e <- gs_fen$est_avg[1,"CI.upper"]
pval_e  <- gs_fen$est_avg[1,"p.value"]

cat(sprintf(
  "Average post-treatment ATT: **%.3f log points** (SE = %.3f; 95%% CI [%.3f, %.3f]; p = %.3f), corresponding to approximately **%+.1f%%**. r* = %d; %d donor sites.\n\n",
  att_e, se_e, ci_lo_e, ci_hi_e, pval_e, pct_e, gs_fen$r_cv, gs_fen$N_co
))
```


## Comparative Synthesis {#sec-comparison}

```{r tbl-comparison-five}
#| label: tbl-comparison-five
#| tbl-cap: "gsynth ATT estimates for all five treated observatories. Clean donor pool throughout. Farafangana marked as indicative only."

get_ci <- function(gs, side) if (!is.null(gs$est_avg)) gs$est_avg[1, side] else NA_real_
get_se <- function(gs) if (!is.null(gs$est_avg)) gs$est_avg[1, "S.E."] else NA_real_
get_pv <- function(gs) if (!is.null(gs$est_avg)) gs$est_avg[1, "p.value"] else NA_real_

tibble(
  Observatory        = c("Marovoay (03)", "Alaotra (21)", "Farafangana (15) ⚠",
                         "Toliara North (23)", "Fénérive East (24)"),
  `PA type`          = c("Strict (II)", "Multipurpose (V)", "Multipurpose (V)",
                         "Mixed (V/II)", "Low-intensity (IV)"),
  `Treat. yr`        = c(2003, 2008, 2006, 2007, 2007),
  `ATT (log pts)`    = round(c(att_m, att_a, att_f, att_t, att_e), 3),
  `SE`               = round(c(se_m, se_a, se_f, se_t, se_e), 3),
  `95% CI`           = c(
    paste0("[", round(ci_lo_m, 3), ", ", round(ci_hi_m, 3), "]"),
    paste0("[", round(ci_lo_a, 3), ", ", round(ci_hi_a, 3), "]"),
    paste0("[", round(ci_lo_f, 3), ", ", round(ci_hi_f, 3), "]"),
    paste0("[", round(ci_lo_t, 3), ", ", round(ci_hi_t, 3), "]"),
    paste0("[", round(ci_lo_e, 3), ", ", round(ci_hi_e, 3), "]")
  ),
  `ATT (%)`          = paste0(round((exp(c(att_m, att_a, att_f, att_t, att_e)) - 1) * 100, 1), "%"),
  `p-value`          = round(c(pval_m, pval_a, pval_f, pval_t, pval_e), 3),
  `r* (CV)`          = c(gs_maro$r_cv, gs_ala$r_cv, gs_fara$r_cv, gs_tol$r_cv, gs_fen$r_cv),
  `N donors`         = c(gs_maro$N_co, gs_ala$N_co, gs_fara$N_co, gs_tol$N_co, gs_fen$N_co),
  `Post yrs`         = c(11L, 6L, 3L, 8L, 8L)
) |>
  gt() |>
  tab_header(
    title    = md("**Five-Observatory Comparative Results**"),
    subtitle = "⚠ Farafangana: only 3 post-treatment years — treat as indicative"
  ) |>
  tab_style(style = cell_fill(color = "#FFCDD2"),
            locations = cells_body(rows = grepl("Marovoay", Observatory))) |>
  tab_style(style = cell_fill(color = "#BBDEFB"),
            locations = cells_body(rows = grepl("Alaotra", Observatory))) |>
  tab_style(style = cell_fill(color = "#DCEDC8"),
            locations = cells_body(rows = grepl("Farafangana", Observatory))) |>
  tab_style(style = cell_fill(color = "#FFE0B2"),
            locations = cells_body(rows = grepl("Toliara", Observatory))) |>
  tab_style(style = cell_fill(color = "#E8EAF6"),
            locations = cells_body(rows = grepl("Fénérive", Observatory))) |>
  tab_footnote("⚠ Farafangana estimates are unreliable due to only 3 post-treatment years.",
               locations = cells_body(rows = grepl("Farafangana", Observatory),
                                      columns = Observatory))
```

```{r fig-comparison-five}
#| label: fig-comparison-five
#| fig-cap: "ATT paths for all five treated observatories on the same scale (relative time axis). Ankarafantsika shown in red with the deepest and most persistent negative effect. CI bands suppressed for legibility; see individual plots above."
#| fig-height: 5.5
#| fig-width: 11

all_att <- bind_rows(
  att_maro_df |> mutate(obs = "Marovoay (strict, II)",         color_grp = "Marovoay (strict)"),
  att_ala_df  |> mutate(obs = "Alaotra (multipurpose, V)",     color_grp = "Alaotra (multipurpose)"),
  att_fara_df |> mutate(obs = "Farafangana (multipurpose, V)", color_grp = "Farafangana (multipurpose)"),
  att_tol_df  |> mutate(obs = "Toliara North (mixed, V/II)",   color_grp = "Toliara North (mixed)"),
  att_fen_df  |> mutate(obs = "Fénérive East (low int., IV)",  color_grp = "Fénérive East (low int.)")
)

ggplot(all_att, aes(rel_time, att, color = color_grp, linetype = color_grp)) +
  geom_hline(yintercept = 0, color = "grey60", linewidth = 0.4) +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey50", linewidth = 0.4) +
  annotate("rect", xmin = -0.5, xmax = Inf, ymin = -Inf, ymax = Inf,
           fill = "grey90", alpha = 0.18) +
  geom_line(linewidth = 0.85) +
  geom_point(size = 1.8) +
  scale_color_manual(
    values = c("Marovoay (strict)"        = "#D32F2F",
               "Alaotra (multipurpose)"   = "#1565C0",
               "Farafangana (multipurpose)"= "#2E7D32",
               "Toliara North (mixed)"    = "#E65100",
               "Fénérive East (low int.)" = "#6A1B9A"),
    name = NULL
  ) +
  scale_linetype_manual(
    values = c("Marovoay (strict)"        = "solid",
               "Alaotra (multipurpose)"   = "solid",
               "Farafangana (multipurpose)"= "dashed",
               "Toliara North (mixed)"    = "solid",
               "Fénérive East (low int.)" = "solid"),
    name = NULL
  ) +
  labs(x = "Years relative to treatment onset",
       y = "ATT (log equiv. income)",
       title = "ATT Paths: Five Conservation Governance Models",
       subtitle = "Dashed = Farafangana (indicative only, 3 post-treatment years)") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 8.5))
```

```{r fig-att-forest}
#| label: fig-att-forest
#| fig-cap: "Forest plot of average post-treatment ATTs. Horizontal bars are 95% parametric bootstrap CIs. Observatories sorted by estimated ATT."
#| fig-height: 4
#| fig-width: 8

forest_df <- tibble(
  obs      = c("Marovoay\n(strict, II, 2003)",
               "Alaotra\n(multipurpose, V, 2008)",
               "Farafangana ⚠\n(multipurpose, V, 2006)",
               "Toliara North\n(mixed, V/II, 2007)",
               "Fénérive East\n(low int., IV, 2007)"),
  att      = c(att_m, att_a, att_f, att_t, att_e),
  ci_lo    = c(ci_lo_m, ci_lo_a, ci_lo_f, ci_lo_t, ci_lo_e),
  ci_hi    = c(ci_hi_m, ci_hi_a, ci_hi_f, ci_hi_t, ci_hi_e),
  col      = c("#D32F2F","#1565C0","#2E7D32","#E65100","#6A1B9A"),
  indicative = c(FALSE, FALSE, TRUE, FALSE, FALSE)
) |>
  arrange(att) |>
  mutate(obs = factor(obs, levels = obs))

ggplot(forest_df, aes(y = obs, x = att, xmin = ci_lo, xmax = ci_hi,
                      color = col, alpha = indicative)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_linerange(linewidth = 1.2) +
  geom_point(size = 3.5) +
  scale_color_identity() +
  scale_alpha_manual(values = c("FALSE" = 1, "TRUE" = 0.45), guide = "none") +
  labs(x = "ATT (log equiv. income)", y = NULL,
       title = "Forest Plot: Five-Observatory ATT Estimates",
       subtitle = "⚠ = Farafangana (3 post-treatment years, indicative only)") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank())
```


# Long-Run Evidence: 2025 Resurvey Extension {#sec-extension}

The 2025 extension applies only to Marovoay, which has a resurvey with matched fokontany-level data. The construction and results are detailed in [Chapter 12, §6](12_strict_vs_multipurpose.html#sec-extension). For completeness, the extended Marovoay ATT (1999–2025) is reproduced here using the clean donor pool.

```{r ext-gsynth-maro}
#| label: ext-gsynth-maro

# Load the 2025 income computation (same as Ch. 12 -- requires utils source)
source("documentation/2025_report/utils/calc_incomes_all_years.R")

inc_2025_raw <- compute_income_year(2025)
m_a_2025     <- haven::read_dta("data/ROS_MDG_microdata/2025/res_m_a.dta")

oecd_2025 <- m_a_2025 |>
  mutate(j5 = as.character(j5)) |>
  summarise(n_adults   = sum(m5 >= 14, na.rm = TRUE),
            n_children = sum(m5 < 14,  na.rm = TRUE),
            .by = j5) |>
  mutate(oecd_equiv = 1 + 0.5 * pmax(n_adults - 1, 0) + 0.3 * n_children, year = 2025L)

deb_2025_fkt <- haven::read_dta("data/ROS_MDG_microdata/2025/res_deb.dta") |>
  mutate(j5 = as.character(j5)) |>
  select(j5, j4) |>
  distinct(j5, .keep_all = TRUE)

inc_2025_hh <- inc_2025_raw |>
  mutate(j5 = as.character(j5)) |>
  left_join(oecd_2025 |> select(j5, oecd_equiv), by = "j5") |>
  left_join(deb_2025_fkt, by = "j5") |>
  mutate(
    oecd_equiv = replace_na(oecd_equiv, 1),
    obs        = case_when(
      substr(j5, 1, 3) %in% c("310","311","320","330") ~ "03",
      substr(j5, 1, 3) %in% c("211","212","213","214","215") ~ "21",
      TRUE ~ NA_character_
    ),
    site_id = case_when(
      grepl("^Bepako",         j4, ignore.case = TRUE) ~ "031",
      grepl("^Madiromiongana", j4, ignore.case = TRUE) ~ "032",
      grepl("^Ampijoroa",      j4, ignore.case = TRUE) ~ "033",
      grepl("^Maroala",        j4, ignore.case = TRUE) ~ "034",
      TRUE ~ substr(j5, 1, 3)
    ),
    revtot_eq = revtot / oecd_equiv
  )

q_2025 <- quantile(inc_2025_hh$revtot_eq, c(0.01, 0.99), na.rm = TRUE)
inc_2025_hh <- inc_2025_hh |>
  mutate(revtot_w    = pmin(pmax(revtot_eq, q_2025[1]), q_2025[2]),
         ln_revtot_w = log(pmax(revtot_w, 0) + 1))

maro_2025_ext <- inc_2025_hh |>
  filter(obs == "03") |>
  mutate(treatment = as.integer(site_id %in% c("031","032"))) |>
  summarize(ln_revtot_w = mean(ln_revtot_w, na.rm = TRUE),
            n_hh = n(), .by = c(site_id, treatment)) |>
  mutate(year = 2025L)

maro_gs_ext <- bind_rows(maro_gs, maro_2025_ext)

gs_ext_cache <- "data/paper14_gsynth_maro_ext_oecd.rds"
need_rerun <- !file.exists(gs_ext_cache) ||
  is.null(read_rds(gs_ext_cache)$est_avg)

if (need_rerun) {
  fit_ext <- gsynth(ln_revtot_w ~ treatment, data = maro_gs_ext,
                    index = c("site_id", "year"), force = "two-way",
                    CV = TRUE, r = c(0, 5), se = TRUE,
                    inference = "parametric", nboots = 200,
                    parallel = FALSE, min.T0 = 3, seed = 42)
  write_rds(save_gsynth(fit_ext), gs_ext_cache)
}
gs_maro_ext <- read_rds(gs_ext_cache)
```

```{r fig-ext-att}
#| label: fig-ext-att
#| fig-cap: "Ankarafantsika extended ATT path (1999–2025) with clean donor pool. The 2025 point (diamond) represents the resurvey year, 22 years after reclassification."
#| fig-height: 5
#| fig-width: 10

att_ext_df <- tibble(
  rel_time = as.numeric(rownames(gs_maro_ext$est_att)),
  att      = gs_maro_ext$est_att[,"ATT"],
  ci_lo    = gs_maro_ext$est_att[,"CI.lower"],
  ci_hi    = gs_maro_ext$est_att[,"CI.upper"]
)
att_2025 <- tail(att_ext_df, 1)

ggplot(att_ext_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#D32F2F", alpha = 0.12) +
  geom_line(color = "#D32F2F", linewidth = 0.8) +
  geom_point(color = "#D32F2F", size = 2.5) +
  geom_point(data = att_2025, color = "#D32F2F", size = 4.5, shape = 18) +
  annotate("text", x = att_2025$rel_time - 0.4, y = att_2025$att + 0.1,
           label = "2025", fontface = "bold", color = "#D32F2F",
           hjust = 1, size = 3.5) +
  labs(x = "Periods relative to treatment (2003 = 0)",
       y = "ATT (log equiv. income)",
       title = "Ankarafantsika Extended (1999–2025) — Clean Donor Pool",
       subtitle = paste0(
         "Avg ATT = ", round(gs_maro_ext$att_avg, 3),
         " (p = ", format.pval(gs_maro_ext$est_avg[1,"p.value"], digits = 3), ")",
         " | 2025 ATT = ", round(att_2025$att, 3)
       )) +
  theme_minimal(base_size = 12)
```

**Lac Alaotra (2025)**: No 2025 resurvey data is available for Alaotra beyond Ch. 12, and the three new observatories were not included in the 2025 survey. The extended analysis is therefore Marovoay-only.


# Discussion {#sec-discussion}

## Five Cases, Three Patterns

The expanded five-observatory analysis broadly confirms the strict/multipurpose contrast identified in Chapter 12, while adding nuance about enforcement intensity and PA size:

**Pattern 1 — Negative income effect under strict enforcement**: Ankarafantsika (Marovoay) shows the largest and most persistent negative ATT, statistically significant and deepening through 2025. This is consistent with direct resource exclusion from a large, actively patrolled park after reclassification from an open-access forest reserve.

**Pattern 2 — Near-zero effect under participatory/multipurpose conservation**: Lac Alaotra, Toliara North, and (tentatively) Farafangana show point estimates clustered near zero. None produces a statistically significant effect. This is consistent with the multipurpose PA framework: land-use regulation without resource exclusion does not translate into detectable aggregate income losses at the survey-year frequency.

**Pattern 3 — Low-intensity reserve, near-zero effect**: Fénérive East and its adjacent Réserve de Tampolo show a near-zero ATT with a moderately narrow CI, consistent with the reserve's small footprint and research-oriented management. The absence of an effect is perhaps the most interpretively clear result of the three new observatories: a 7.3 km² research forest cannot plausibly impose population-level income costs on a multi-commune observatory.

## Interpretation Caveats

1. **Farafangana power**: With only 3 post-treatment years, the Farafangana estimate is unreliable. The fact that it exists at all reflects the feasibility screening in Chapter 13 (≥3 post years), but 3 is the minimum, not the optimum. This estimate should not be compared numerically to the others.

2. **Toliara North treatment complexity**: The three-PA cascade (January 2007 + April 2007 + December 2008) means the Toliara North ATT estimates a composite treatment rather than a clean single event. The December 2008 secondary shock (Ranobe PK 32, the closest PA at <1.5 km) coincides with a large post-treatment change in the series, making it impossible to separate the 2007 and 2008 effects without additional micro-data.

3. **Clean donor pool size**: Using only five donors (vs. nine in Chapter 12) reduces counterfactual precision. The r* cross-validation may select fewer factors because the smaller donor pool supports fewer reliable latent dimensions. Comparisons between Chapter 12 and Chapter 14 ATT estimates for Marovoay and Alaotra may reflect this difference rather than true effect heterogeneity.

4. **No 2025 validation for new sites**: The identification strategy for the three new observatories rests entirely on decree dates and proximity — the type of evidence that Chapter 13 identified as insufficient for causal inference without behavioural validation. These estimates are best treated as exploratory.

5. **Single-unit power floor**: Each observatory-level model has exactly one treated unit. The permutation p-value floor is $\approx 1/N_{donors}$; with ~50–60 donor sites, this gives $\approx 0.017$ — better than Chapter 12's Alaotra floor but still requiring large true effects to achieve nominal significance.

## Implications for Conservation Policy

Despite their limitations, the five estimates together suggest that the **conservation governance model matters more than geographic proximity** for household income outcomes. The two variables that predict negative effects are (1) strict IUCN category (II vs. V/VI) and (2) large PA size that constrains broad livelihood activities across an observable area. The three new observatories all face PAs that are either multipurpose (V), small-footprint, or both — and none shows evidence of harm at the aggregate income level.

This pattern is consistent with the proposition that participatory conservation under the post-2003 SAPM framework successfully avoided the welfare costs associated with the older strict protection model. Whether it also achieved biodiversity and ecosystem-service goals is a separate question not addressed here.


# Methodological Perspectives: From Synthetic Control to Difference-in-Differences {#sec-did-perspectives}

The gsynth results presented above — while informative — expose structural limitations of the synthetic control framework when applied to multi-observatory panel data of this kind. This section synthesises a set of methodological reflections on how difference-in-differences (DiD) methods could exploit the same data more effectively, and outlines concrete designs for current and future analyses.

## Limitations of the Generalised Synthetic Control Approach {#sec-gsynth-limits}

The gsynth estimator requires site-level aggregation and balanced panels. Both requirements create severe constraints in this setting:

1. **Site-level aggregation discards within-site variation.** Each fokontany contributes ~130–260 households per survey year, but gsynth collapses these to a single site-year mean. The effective sample is therefore ~4 site-year observations per observatory per year — not the ~500 household observations actually available.

2. **Balanced panel requirement causes catastrophic donor attrition.** gsynth requires every donor unit to be observed in every period. The five clean donor observatories cover different year ranges (Antsirabe: 1995–2014 with gaps; Tsiroanomandidy: 1999–2005; Ambovombe: 1999–2008; Mahanoro: 2000–2010; Itasy: 2000–2009). Intersecting these ranges with the Alaotra model's 1999–2014 window, the number of donor sites with complete coverage drops from 47 in 2003 to as few as 4 in 2009. The counterfactual quality degrades sharply in exactly the post-treatment years where precision matters most.

3. **Each observatory is modelled separately.** The five gsynth models above do not pool information across treatment cohorts. The staggered timing of PA creation (2003, 2006, 2007, 2008) is an identification resource that goes entirely unexploited.

4. **Single treated unit per model.** The observatory-level models (Alaotra, Farafangana, Toliara North, Fénérive East) each have exactly one treated unit, imposing a permutation p-value floor of $\approx 1/N_{donors}$. Even with 50 donors, effects smaller than ~20% are difficult to distinguish from zero.

5. **No household-level covariates.** The factor model adjusts for unobserved common shocks, but cannot condition on observable household characteristics (household size, education, land holdings) that might improve balance between treated and control units.

## How DiD Changes the Picture {#sec-did-advantage}

Modern heterogeneity-robust DiD estimators — in particular @CallawayS2021 and @SunAbraham2021 — address all five limitations simultaneously:

| Dimension | gsynth (current) | Staggered DiD |
|-----------|-----------------|---------------|
| Unit of analysis | Site (~4 per obs) | Household (~500/year) |
| Sample size | ~60 site-years | 44,000+ HH-year obs |
| Panel balance | Required | Not required |
| Cross-obs pooling | No (separate models) | Yes (staggered cohorts) |
| Treatment cohorts | 1 per model | 2–4 simultaneously |
| Clusters for inference | 2 treated sites | 5 treated + 38 control = 43 sites |
| HH covariates | Not available | Conditional parallel trends |
| Pre-trends | Factor model $r^*$ | Explicit event-study plot |

: Comparison of generalised synthetic control and staggered DiD frameworks applied to the ROR data. {#tbl-gsynth-did}

The key insight is that the Rural Observatory Network generates **repeated cross-sections** — different households are sampled each year within the same fokontany — and the @CallawayS2021 estimator is specifically designed for this data structure (`panel = FALSE`). No household tracking across years is required. Every household-year observation contributes independently to estimation.

## Staggered Treatment Design {#sec-staggered-design}

The five treated observatories define four treatment cohorts, creating a natural staggered DiD framework:

- $g = 2003$: Marovoay east-bank (sites 031, 032) — strict enforcement at Ankarafantsika
- $g = 2006$: Farafangana (9 sites) — CFAV corridor protection
- $g = 2007$: Toliara North (4 sites) + Fénérive East (10 sites) — coastal PA designations
- $g = 2008$: Alaotra (3 sites) — CBNRM enforcement at Lac Alaotra^[The PA was gazetted in 2003 but effective community-based management under Durrell began in 2008; see §2.2.]
- $g = \infty$: Marovoay west-bank (2 sites) + five clean donor observatories (36 sites) — never treated

This design has a crucial structural advantage: **not-yet-treated units serve as controls.** Alaotra (treated from 2008) functions as a valid control for the Marovoay effect during 2003–2007, before its own treatment begins. This cross-cohort identification is impossible in the gsynth framework where each observatory is modelled independently.

The @CallawayS2021 estimator produces group-time ATTs — $ATT(g, t)$ for each cohort $g$ and calendar year $t$ — which can be aggregated into:

- **Dynamic effects** (event-study): average ATT at each relative time $e = t - g$, with transparent pre-trend tests
- **Group-level effects**: average ATT for each treatment cohort, enabling direct comparison of strict vs. multipurpose governance
- **Overall ATT**: a single summary weighted across all cohorts and post-treatment periods

## Concrete Designs {#sec-did-designs}

### With existing data (1999–2014 + 2025 resurvey)

**Design A — Two-cohort CS-DiD (most conservative).** Restrict to Marovoay ($g = 2003$) and Alaotra ($g = 2008$), with Marovoay west-bank and clean donors as the never-treated group. This pools 13,427 treated household-year observations (5,814 Marovoay east + 7,613 Alaotra) with 31,305 control observations across 43 site-level clusters. The 2025 resurvey adds a long-run period: ATT(2003, 2025) for Marovoay (260 treated HH) and ATT(2008, 2025) for Alaotra (507 HH), with Marovoay west-bank (259 HH) as the only never-treated control in 2025. Year coverage for the never-treated group spans 1995–2014 continuously (503–2,941 control HH per year), with even the thinnest year (2009) having 510 observations — a situation gsynth cannot exploit because it demands that *every* donor be observed in *every* period.

**Design B — Four-cohort staggered CS-DiD (full exploitation).** Include all five treated observatories across four cohorts plus the never-treated pool. Total sample exceeds 64,000 household-year observations across 66 sites. This is the richest possible design with existing data, but requires accepting observatory-level treatment assignment for Farafangana, Toliara North, and Fénérive East (i.e., all sites within each treated observatory are classified as treated). Farafangana's 3-year post-window remains a power concern even in the DiD framework.

### With 2026 resurvey data (Fénérive East)

Adding a Fénérive East resurvey unlocks a **within-observatory distance design**. The Réserve de Tampolo (7.3 km²) is small enough that the distance threshold between treated ($< 15$ km) and control ($> 15$ km) sites is credible — unlike the Toliara PA complex (5,073 km²), where even the 12.5 km commune is ambiguously treated. With 2–3 near-PA sites and 4+ far-PA sites, Fénérive contributes to both the treated and control pools, enriching the staggered design with a third independent geographic context alongside Marovoay and Alaotra.

A further option with 2026 data is **continuous-treatment DiD**, where the dose is site-to-PA distance rather than a binary indicator. Recent extensions by @CallawayGoodmanBaconSantAnna2024 accommodate dose-response in the staggered framework, exploiting all variation in PA proximity without arbitrary thresholds.

## Complementary Estimators and Robustness Architecture {#sec-robustness-architecture}

The revised estimation strategy places DiD methods as the **primary specification** and gsynth as a **robustness check**:

1. **Primary**: Staggered @CallawayS2021 DiD (`did` package), repeated cross-section mode, doubly-robust estimation with household-level covariates, clustered at the site level.
2. **Complementary**: @SunAbraham2021 interaction-weighted estimator via `fixest::sunab()`, providing an alternative heterogeneity-robust estimator with transparent event-study syntax.
3. **Robustness — interactive fixed effects**: The gsynth estimates from this chapter serve as a robustness check that relaxes the parallel trends assumption to an interactive fixed-effects model. If gsynth and DiD agree, the evidence is strong regardless of which identifying assumption is correct. If they disagree, the factor model may be capturing unobserved confounders that violate parallel trends.
4. **Sensitivity**: @deChaisemartinDHaultfoeuille2020 estimator to assess potential negative weights in a naïve TWFE specification, and wild cluster bootstrap [@CameronGelbachMiller2008] for inference with the small number of treated clusters.

This layered architecture — transparent DiD as the primary lens, factor models as robustness — is standard in recent conservation-impact evaluations and represents a methodological upgrade from the single-estimator approach used in Chapters 12–14.

## Summary

The gsynth analyses in this chapter provide a useful starting point, but the data support a substantially more powerful identification strategy. The staggered timing of PA creation across five observatories, the repeated cross-section design of the Rural Observatory Network, and the large within-site household samples are all features that DiD methods exploit naturally and that synthetic control methods cannot. The next analytical step is to implement the staggered CS-DiD framework outlined above, treating the gsynth results as one element of a multi-estimator robustness architecture.


# Conclusion {#sec-conclusion-final}

Extending the Chapter 12 analysis to include three additional conservation-adjacent observatories — Farafangana, Toliara North, and Fénérive East — reinforces the core finding: **strict exclusionary conservation is associated with measurable negative income effects on adjacent communities, while multipurpose and low-intensity conservation is not**. The five-case comparison spans a range of PA types (IUCN II, IV, V, and mixed), geographic contexts (highland corridor, coastal arid zone, northeastern littoral), and post-treatment survey windows (3–11 years).

The clean donor pool constraint (five observatories, from the Chapter 13 audit) reduces counterfactual precision relative to Chapter 12 but eliminates a potential source of systematic bias. The Marovoay and 2025 extended results are robust to this change. The Farafangana result should not be interpreted independently given the thin post-treatment window; the Toliara North and Fénérive East results are more credible and both support the null of no aggregate income effect under non-strict conservation.

Beyond these substantive findings, a methodological reflection on the limitations of the gsynth framework in this setting points toward staggered difference-in-differences as a more natural primary specification. The Rural Observatory data — repeated cross-sections, staggered treatment timing, large household samples, unbalanced panels across observatories — are better suited to the @CallawayS2021 estimator than to the balanced-panel requirements of generalised synthetic control. The gsynth results presented here should therefore be read as a first-pass analysis that will be complemented by a household-level DiD estimation exploiting the full depth of the data.

These results strengthen the case for distinguishing between conservation governance models in welfare-impact analyses and for treating the IUCN category and enforcement intensity as first-order moderators of conservation-income relationships in rural Madagascar.


# References {.unnumbered}

::: {#refs}
:::
