---
title: "Appendix A: Data Pipeline"
subtitle: "Georeferencing, PA overlaps, and variable harmonisation"
---

```{r setup}
#| label: setup
source("R/load_data.R")
```

This appendix documents the full reproducible data pipeline from raw ROS survey files to the consolidated household-level panel dataset. The paper chapter (@sec-data) provides a condensed summary; this appendix gives all details needed for reproducibility.


# Georeferencing ROS localities {#sec-a-georef}

## The problem

The Rural Observatory System collected household panel data in rural Madagascar from 1995 to 2015. Critically, surveys were **not geolocated**: location fields were free-text, not linked to official administrative codes. To enable spatial analysis --- particularly overlay with protected-area boundaries --- every ROS observation must be aligned with a standardised commune and *fokontany* using P-codes from the OCHA/BNGRC Common Operational Datasets (CODs).

## Reference datasets

Two spatial reference datasets anchor the georeferencing:

1. **Madagascar Subnational Administrative Boundaries** (COD-AB): five administrative levels from country to *fokontany*, with official P-codes.
2. **Madagascar Populated Places** (COD-PS): 28,184 populated places with toponyms and P-codes (source: NGA/GIST, distributed by UN OCHA ROSA).

## String cleaning

All location strings are normalised before matching. The `clean_string()` function lowercases, strips non-alphanumeric characters, removes qualifiers ("centre", "haut", "bas"), translates French cardinal directions to Malagasy, standardises alternative city names (e.g. "Fort Dauphin" → "Tolanaro"), and converts Roman numerals I--VI to Arabic.

## Hierarchical fuzzy matching

Matching uses **Jaro-Winkler distance** (default threshold 0.25), applied observatory-wise to restrict candidate sets. A four-method cascade assigns communes:

1. **Method 1** --- Fuzzy-match the municipality field (`j42`) against known observatory commune names. Visually verified; false positives removed manually.
2. **Method 2** --- For records with no municipality, extract the first word of the village name and fuzzy-match it against commune names.
3. **Method 3** --- Fuzzy-match the full village name against COD populated places (threshold tightened to 0.22).
4. **Method 4** --- Fuzzy-match remaining village names against already-matched names from prior steps (threshold 0.28).

After commune assignment, a dedicated `fuzzy_match_village()` function matches village names to ADM4 (*fokontany*) names within the matched commune (Jaro-Winkler threshold 0.2). A final **Method 6** adds 13 hand-coded entries for observatory 52 (Menabe).

```{r}
#| label: fig-map-georef
#| fig-cap: "All matched ROS communes and fokontany across Madagascar."
#| eval: false

# Interactive tmap of matched communes/fokontany coloured by observatory
```

## Match rates

Communes were re-identified for **91.8%** of observations; villages for **61.5%** of name variations. The 8.2% commune failures are concentrated in observatories with highly abbreviated or non-standard location strings. 98% of unique commune name variations and 62% of village name variations were resolved.


# Protected-area overlap analysis {#sec-a-overlaps}

## Spatial overlaps

ROS commune and *fokontany* polygons are overlaid with the **Vahatra PA shapefile** (`AP_Vahatra.shp`), which provides boundaries for all Madagascar PAs including post-2003 SAPM additions. An interactive map visualises all observatory polygons against PA boundaries.

```{r}
#| label: fig-map-overlaps
#| fig-cap: "ROS survey communes (coloured by observatory) overlaid with PA boundaries (dark green)."
#| eval: false

# Interactive tmap: ROS polygons + Vahatra PA boundaries
```

## ROS survey inventory

Household-level datasets for 1995--2015 are loaded and tabulated by observatory and year. Survey continuity varies: some observatories have nearly continuous 20-year panels, others cover only 3--8 years.

```{r}
#| label: tbl-survey-inventory
#| tbl-cap: "Number of surveyed households per observatory per year (1995--2015)."
#| eval: false

# gt table of observatory × year household counts
```

## Temporal overlap logic

Three scenarios determine usability for impact evaluation:

1. **PA predates ROS entirely** → no pre-treatment baseline → unusable (e.g. Ranomafana, created 1991).
2. **PA created during ROS surveys** → impact analysis possible → *these are our treatment cases*.
3. **PA created after ROS ended** → baseline only; resurvey needed for evaluation.

## Decree date revision

Official creation years were cross-checked against the Madagascar legal database ([CNLEGIS](https://cnlegis.gov.mg)). Several temporary decree dates were revised:

| PA | Original year | Revised year | Source |
|:---|:---:|:---:|:---|
| Ankarafantsika NP | 1927 | **2002** | Decree 2002-798 (reclassification) |
| Lac Alaotra PHP | 2003 | **2007** | Temporary decree |
| Corridor Ankeniheny--Zahamena | 2002 | **2005** | Temporary decree |
| Menabe Antimena | 2003 | **2006** | Temporary decree |
| Corridor Ambositra-Vondrozo | 2003 | **2006** | WDPA / CNLEGIS |
| Mikea NP | 2003 | **2007** | Temporary decree |
| Ranobe PK 32 | 2003 | **2008** | Temporary decree |
| Amoron'i Onilahy | 2003 | **2007** | Temporary decree |

: Revised PA decree dates used in this study {#tbl-decree-dates}

```{r}
#| label: fig-timeline
#| fig-cap: "Timeline of ROS survey years (blue dots) vs. revised PA creation years (red triangles) for priority observatory--PA pairs."
#| eval: false

# Faceted dot-plot timeline
```

## Priority pairs

The spatial and temporal overlay identifies Marovoay--Ankarafantsika and Alaotra--Lac Alaotra as the most promising pairs for causal analysis (long panels, clear treatment dates). Farafangana, Toliara North, and Fénérive East are secondary candidates explored in @sec-extensions. A 10 km buffer (CRS EPSG:29701, Laborde Madagascar) around PAs is used for the proximity screening.

::: {.callout-note}
ROS observatories were selected for agricultural and socio-economic monitoring, not proximity to PAs. Overlaps are partly coincidental and require cautious interpretation.
:::


# Variable construction {#sec-a-variables}

## Georeferencing households

Household identifiers (`j5`) are joined to the locality correspondence table from @sec-a-georef. A critical correction: **1995 household IDs are re-aligned using the 1996 survey** (which records whether a household was surveyed in 1995 and provides its 1996 identifier).

## Household composition

Individual-level files (member records) provide sex, age, and headship. Households with multiple heads are resolved by keeping the first member ID. Key variables: `household_size` (member count) and head's sex and age.

## Education

Education variables present several data gaps:

- Literacy (`s1a`, `s1b`): missing in 1996 and 1998.
- Primary school attendance (`s2`): missing in 1996.
- Years of schooling (`s3a`): missing in 1996.

Household-level indicators are constructed: `any_literate_adult` (≥1 adult reads), `max_years_edu`, `mean_years_edu_adults`, `pct_with_schooling` (share of members ≥6 with any schooling), and head's education variables.

## Agricultural activities

Activity codes (`a1`) are harmonised across the 1996--2015 coding schemes. **1995 is a special case**: no individual `a1` variable; activities are reconstructed from separate main/secondary activity files (`res_apm.dta`, `res_asm.dta`). Two agricultural classifications are used:

- **Stricto sensu** (codes 95, 96): farming or livestock only.
- **Lato sensu** (19 codes): farming, livestock, fishing, gathering, apiculture, agriculture-related crafts.

## Food security

The core measure is **months of food self-sufficiency from own production**, harmonised across varying question formats:

- 1995: single question (`sa3a`).
- 1996--1997: single question (`sa2`).
- 1998--2015: three components (`sa2a` + `sa2b` + `sa2c` for rice 1st season + rice 2nd season + maize).

Summed and capped at 12 months.

## Income construction

Income is reconstructed via dedicated functions for each component:

- **Current income** = main activity + secondary activity + net rice + other crops + livestock + fishing.
- **Exceptional income** = land rents + miscellaneous + HIMO (public works) + asset sales + monetary and non-monetary transfers.
- **Total income** (`revtot`) = current + exceptional.

The OECD-modified equivalence scale assigns weight 1.0 to the first adult, 0.5 to each additional adult (≥14), and 0.3 to each child (<14). OECD-equivalised income = `revtot / oecd_scale`. Both total and equivalised income are log-transformed and winsorised at the 1st--99th percentiles for the main analysis.

```{r}
#| label: tbl-income-components
#| tbl-cap: "Income component construction and variable sources."
#| eval: false

# Table: Component | Survey variables | Coverage | Notes
```

## Consolidation

All component tables are deduplicated to unique (year, household ID) via mean aggregation, then left-joined into a single dataset: `household_consolidated.rds` ({{< var n_hh_obs >}} household-year observations across {{< var n_obs >}} observatories). A uniqueness check confirms one row per household per year.
