---
title: "Appendix B: Donor Pool Validity"
subtitle: "Are donor observatories truly unexposed to conservation interventions?"
---

```{r setup}
#| label: setup
source("R/load_data.R")
```

The generalised synthetic control estimator requires that **donor pool units are never exposed to the treatment of interest**. In this study, the nine donor observatories used in the primary analysis (@sec-results) should not themselves be adjacent to protected areas that became actively enforced during the 1999--2014 study period. This assumption is plausible by design --- ROS observatories were chosen for agricultural monitoring, not PA proximity --- but has never been systematically verified. This appendix provides a spatial audit.


# Why static PA boundaries are insufficient {#sec-b-static}

Standard PA databases --- including the current WDPA and the Vahatra dataset --- record the **most recent boundary and designation** of each protected area. They do not convey:

1. **Boundary changes**: many Malagasy PAs were significantly extended or reorganised between 1999 and 2015.
2. **Temporary protection decrees**: Madagascar's post-2003 SAPM expansion created PAs first via temporary decrees (*décrets provisoires*) before formal designation.
3. **Date accuracy**: numerous PAs in the static WDPA carry incorrect or missing `STATUS_YR` values.


# Dynamic WDPA construction {#sec-b-wdpa}

We use the **dynamic WDPA for Madagascar** (`dynamic_wdpa.rds`), a temporally explicit database that corrects and extends the standard WDPA with historical states derived from legal texts, multiple institutional datasets, and spatially referenced decree dates. Each record carries a valid-from / valid-to date range, allowing us to reconstruct which PA boundaries were legally operative in any given year.

```{r}
#| label: fig-wdpa-timeline
#| fig-cap: "Timeline of PA activation dates by observatory. Study period (1999--2014) shaded."
#| eval: false

# Dot-plot of PA activation dates, colour-coded by observatory exposure type
```


# Proximity analysis {#sec-b-proximity}

## Method

Observatory locations are represented by commune-level polygon centroids from `Observatories_ROS_communes_COD_v4.gpkg`. The dynamic WDPA is filtered to external boundaries active during 1999--2014, then unioned per WDPAID. Both layers are projected to UTM 38S for distance computation.

For each (commune centroid, PA boundary) pair, we compute centroid-to-boundary distance in km. All pairs within **{{< var threshold_km >}} km** are retained and classified by exposure timing:

- **Pre-existing**: PA active before 1999 with unchanged boundaries throughout.
- **New during study**: PA became operative between 1999 and 2014.
- **Post-study**: PA created after 2014.

```{r}
#| label: tbl-proximity
#| tbl-cap: "All observatory--PA pairs within 20 km, colour-coded by exposure type."
#| eval: false

# Long-form table: observatory, commune, PA name, IUCN, distance, activation date, exposure type
```


# Results {#sec-b-results}

## Exposure classification

The nine donor observatories fall into three categories:

**Clean** (no PA within 20 km during study period):

- Antsirabe (02)
- Tsiroanomandidy (13)
- Ambovombe (16)
- Mahanoro (25)
- Itasy (41)

**Potentially contaminated** (new PA within 20 km during study period):

- **Antsohihy (12)** --- within 0.3 km of Corridor forestier Bongolava (IUCN V, 15 September 2006). The commune of Tsarahasina borders the corridor directly. However, Antsohihy's survey ends in 2005, providing only 1 post-2006 year. It is excluded from both the donor pool and the treated sample.
- **Farafangana (15)** --- within 14 km of Corridor Forestier Ambositra-Vondrozo (IUCN V, 15 September 2006), and with pre-existing proximity to Manombo SR (IUCN IV, 1962). Pre-existing exposure is absorbed by the unit fixed effect; the 2006 CFAV decree represents new treatment.
- **Toliara North (23)** --- within 1.4 km of Ranobe PK 32 (December 2008), 12 km of Amoron'i Onilahy (January 2007), and 16 km of Mikea NP (April 2007). The most problematic case: a cascade of three PAs starting January 2007.
- **Fénérive East (24)** --- within 6 km of Réserve de Tampolo (IUCN IV, August 2007). Small reserve with low enforcement intensity, but proximity is clear.

```{r}
#| label: tbl-donor-status
#| tbl-cap: "Donor pool PA exposure status summary."
#| eval: false

# Table: Observatory | Status | Nearest new PA | Distance (km) | PA year | Recommendation
```

```{r}
#| label: fig-map-exposure
#| fig-cap: "Map of observatory communes coloured by exposure status. Nearby PA polygons shown."
#| eval: false

# Map: clean (green), pre-existing (yellow), contaminated (red) + PA polygons
```

## Detailed exposure profiles

```{r}
#| label: fig-contaminated-detail
#| fig-cap: "Distance-to-PA for each potentially contaminated donor observatory."
#| eval: false

# Lollipop chart: distance to each PA, faceted by observatory
```


# Implications {#sec-b-implications}

## For the primary analysis

The contamination audit has two consequences:

1. **Restricted donor pool**: removing the four contaminated observatories leaves five clean donors (Antsirabe, Tsiroanomandidy, Ambovombe, Mahanoro, Itasy). Gsynth estimates for Marovoay and Alaotra should be re-run with this restricted pool as a robustness check.

2. **Pre-treatment validity preserved for Marovoay**: the treatment onset (2003) predates all contamination events (earliest 2006). Pre-treatment balance tests (placebo ATTs before 2003) are unaffected by post-2006 donor contamination.

## For the extended analysis

The four contaminated donors become **treated units** in the extended analysis (@sec-extensions). Antsohihy is excluded (insufficient post-treatment coverage). Farafangana, Toliara North, and Fénérive East are viable treated cases with ≥3 post-treatment years and plausible identification stories. Their reclassification from donors to treated units simultaneously:

- Purifies the donor pool (removing contaminated units).
- Enriches the treated sample (five observatory--PA pairs instead of two).
- Enables a broader comparison of conservation governance types.

```{r}
#| label: fig-clean-donors
#| fig-cap: "Map of Madagascar with observatory locations coloured by status: treated (Ch. 4), new treated (Ch. 5), clean donor, excluded."
#| eval: false

# Point map of Madagascar with observatory classification
```
