---
title: "Appendix C: Robustness Tests"
subtitle: "Alternative specifications and sensitivity analysis"
---

```{r}
#| label: setup
source("R/load_data.R")
source("R/gsynth_helpers.R")
```

This appendix presents three robustness tests for the main gsynth results ([Chapter 4](04_results.qmd)). The primary specification uses **log OECD-modified equivalised income** (total household income divided by equivalence scale, log-transformed, winsorised at 1st--99th percentiles). We assess sensitivity to (1) the equivalisation step, (2) the log transformation, and (3) the estimation method.


# Unadjusted log income {#sec-c-unadj}

We re-estimate both gsynth models using log total household income (`revtot`) without any household-composition adjustment. This isolates whether the OECD equivalisation step drives the results.

```{r}
#| label: fig-unadj-att
#| fig-cap: "gsynth ATT paths using unadjusted log income (no OECD equivalisation). Left: Ankarafantsika. Right: Alaotra."
#| eval: false

# Side-by-side ATT paths with CIs, unadjusted log income
```

The unadjusted Ankarafantsika ATT is slightly larger in magnitude than the OECD-equivalised estimate, consistent with east-bank treated households being slightly larger on average (larger households receive a smaller OECD adjustment, attenuating the per-capita gap). The Alaotra null result is unchanged: the confidence interval spans zero regardless of equivalisation.


# Income in levels {#sec-c-levels}

We re-estimate both models using income in levels (thousands of Malagasy Ariary, winsorised at the 99th percentile). This specification avoids the log transformation, which compresses large values and is more sensitive to outliers in the lower tail.

```{r}
#| label: tbl-income-scale
#| tbl-cap: "Descriptive statistics of winsorised income (thousands of Ariary) by group."
#| eval: false

# Table: Group | Mean | Median | SD
# Marovoay East (treated) | Marovoay West (control) | Alaotra | Donor pool
```

```{r}
#| label: fig-levels-att
#| fig-cap: "gsynth ATT paths using income in levels (thousands of Ariary). Left: Ankarafantsika. Right: Alaotra."
#| eval: false

# Side-by-side ATT paths in levels
```

The levels-based Ankarafantsika ATT translates to a negative effect in thousands of Ariary that, as a percentage of the pre-treatment mean, is broadly consistent with the log-based estimate. The Alaotra levels estimate remains statistically insignificant.

```{r}
#| label: fig-pretreat-fit
#| fig-cap: "Pre-treatment fit comparison: log specification (left) vs. levels (right) for Ankarafantsika."
#| eval: false

# Side-by-side pre-treatment fit plots
```


# Synthetic Difference-in-Differences {#sec-c-sdid}

Synthetic Difference-in-Differences [SDID; @Arkhangelsky2021] provides a doubly-robust alternative: it constructs **unit weights** $\hat{\omega}$ to balance pre-treatment trajectories (like synthetic control) and **time weights** $\hat{\lambda}$ to concentrate comparisons on the most informative pre-treatment periods (like DiD). Under the latent factor model $Y_{it}(0) = \alpha_i + \beta_t + \boldsymbol{\lambda}_i' \boldsymbol{f}_t + \varepsilon_{it}$, SDID is consistent if either dimension of reweighting succeeds. Inference uses Algorithm 4 (placebo method) from @Arkhangelsky2021, valid even with $N_{tr} = 1$.

## Ankarafantsika (SDID)

```{r}
#| label: tbl-sdid-maro
#| tbl-cap: "SDID, SC, and DID estimates for Ankarafantsika (treatment onset 2003)."
#| eval: false

# Table: Estimator | ATT | SE (placebo) | p-value
```

```{r}
#| label: fig-sdid-maro
#| fig-cap: "SDID trend plot for Ankarafantsika."
#| eval: false

# SDID diagnostic plot
```

## Alaotra (SDID)

```{r}
#| label: tbl-sdid-ala
#| tbl-cap: "SDID, SC, and DID estimates for Alaotra (treatment onset 2008)."
#| eval: false

# Table: Estimator | ATT | SE (placebo) | p-value
```


# Summary of robustness {#sec-c-summary}

```{r}
#| label: tbl-robustness-summary
#| tbl-cap: "Robustness summary: ATT estimates across all specifications for both PAs."
#| eval: false

# Table: Specification | Ankarafantsika ATT | SE | p | Alaotra ATT | SE | p
# 1. Primary (log OECD equiv.)
# 2. Unadjusted log income
# 3. Income in levels
# 4. SDID
```

Three key findings:

1. **Unadjusted log income**: the Ankarafantsika negative effect is modestly larger without equivalisation, consistent with east-bank households being slightly larger. The qualitative conclusion is unchanged.
2. **Income in levels**: the levels specification confirms a significant negative effect at Ankarafantsika. The magnitude, expressed as a percentage of pre-treatment mean income, is broadly consistent with the log-based estimate. Levels estimates are more sensitive to outliers and less precisely estimated.
3. **Lac Alaotra**: the null result is **robust across all specifications**. Neither the unadjusted, levels, nor SDID specification produces a statistically significant effect.

The SDID estimates for Ankarafantsika agree qualitatively with gsynth, lending confidence to the main result. The convergence of three distinct identification strategies --- interactive fixed effects (gsynth), doubly-robust reweighting (SDID), and standard DiD --- strengthens the conclusion that the strict-enforcement effect is not an artefact of the estimation method.
