---
title: "Marovoay–Ankarafantsika: Synthetic Control with Extended Donor Pool"
author: BETSAKA project
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    self-contained: true
editor_options:
  chunk_output_type: console
---

```{r setup}
#| include: false
proj_root <- normalizePath(file.path(getwd(), ".."))
knitr::opts_knit$set(root.dir = proj_root)
setwd(proj_root)

library(tidyverse)
library(haven)
library(gt)
library(fixest)
library(gsynth)
library(panelView)
library(scales)
library(purrr)

source("utils/helpers_report.R")
source("utils/sites.R")
source("utils/calc_incomes_all_years.R")

# ── Treatment constants ───────────────────────────────────────────────────────
MARO_TREATED  <- c("031", "032")
MARO_CONTROL  <- c("033", "034")
MARO_ALL      <- c(MARO_TREATED, MARO_CONTROL)
TREAT_YEAR    <- 2003

# ── Donor pool: all ROS obs with ≥ 2 pre-treatment and ≥ 2 post-treatment years
# obs 02 (Tulear): 1995–2014 (10 sites)
# obs 12: 1999–2004 (10 sites, partial post)
# obs 13: 1999–2005 (10 sites, partial post)
# obs 15: 1999–2008 (9 sites)
# obs 16: 1999–2008 (6 sites)
# obs 21 (Alaotra): 1999–2014 (3 sites active throughout)
# obs 23: 1999–2014 (4 sites)
# obs 24: 1999–2014 (10 sites)
# obs 25: 2000–2010 (6 sites)
# obs 41 (Vakinankaratra): 2000–2009 (4 sites)
DONOR_OBS <- c("02", "12", "13", "15", "16", "21", "23", "24", "25", "41")

ANALYSIS_YEARS <- 1999:2014   # Common start (Alaotra baseline); excl. 2005 below

obs_labels <- c(
  "02" = "Obs02 Tulear",
  "03" = "Obs03 Marovoay",
  "12" = "Obs12",
  "13" = "Obs13",
  "15" = "Obs15",
  "16" = "Obs16",
  "21" = "Obs21 Alaotra",
  "23" = "Obs23",
  "24" = "Obs24",
  "25" = "Obs25",
  "41" = "Obs41 Vakinankaratra"
)

site_treatment <- function(site_id, year) {
  as.integer(site_id %in% MARO_TREATED & year >= TREAT_YEAR)
}
```

## 1 Motivation

The medium-run analysis in `06_Marovoay-Ankarafantsika-extended.qmd` used a
restricted control group: only the two Marovoay control sites (033, 034) plus
the three Alaotra sites (211–213). This yields a donor pool of 5 units to
construct the synthetic control.

The original analysis in `04_Marovoay-Ankarafantsika.qmd` went further: it
used **all ROS sites with overlapping coverage**, which includes observatories
with partial post-treatment data (obs 12 and 13 end in 2004–2005, obs 15–16
in 2008). The `gsynth` interactive fixed-effects (IFE) estimator handles
partially overlapping panels by fitting the factor model only to observed cells
— sites that exit early still contribute to the pre-treatment factor fit, even
though they cannot contribute to post-2008 ATT estimates.

This file re-implements that large donor pool approach with the refined income
variables (log-winsorised) and correct inference (nonparametric bootstrap),
as a complement to the 07 long-run analysis.

::: {.callout-important title="Two complementary analyses"}
- **This file (doc 08)**: Medium-run gsynth, 1999–2014, large donor pool (~62
  sites). More pre-treatment donors → better factor model fit → more
  credible synthetic control. But limited to the period before the data gap.
- **Doc 07**: Long-run RC DiD, 1999–2002 vs. 2025, small control group (4–7
  sites). Gains the 2025 endpoint and the 23-year compound effect, at the cost
  of a weaker donor pool.
:::

## 2 Data Construction

```{r build_panel}
#| message: false
#| cache: true

household_consolidated <- read_rds("data/household_consolidated.rds")

mk_income_ext <- function(df) {
  df |>
    mutate(
      rev_riz  = coalesce(rev_riz,  0),
      rev_cu   = coalesce(rev_cu,   0),
      revel    = coalesce(revel,    0),
      revpeche = coalesce(revpeche, 0),
      revppal  = coalesce(revppal,  0),
      revsec   = coalesce(revsec,   0),
      revcou   = coalesce(revcou, revppal + revsec + rev_riz + rev_cu + revel + revpeche),
      revtot   = coalesce(revtot, revcou),
      rev_agri    = rev_riz + rev_cu + revel + revpeche,
      rev_nonagri = revppal + revsec
    )
}

# ── All ROS sites in the donor pool + treated sites ──────────────────────────
panel_hh_ext <- household_consolidated |>
  filter(
    obs %in% c("03", DONOR_OBS),
    year %in% ANALYSIS_YEARS,
    !(obs == "03" & year == 2005)   # Marovoay 2005 excluded (survey interruption)
  ) |>
  mk_income_ext() |>
  mutate(
    treated_unit = site_id %in% MARO_TREATED,
    treatment    = site_treatment(site_id, year)
  ) |>
  # Winsorise at 99th percentile pooled across all site-years
  mutate(across(c(revtot, revcou, rev_agri, rev_nonagri),
                ~ pmin(.x, quantile(.x, 0.99, na.rm = TRUE)),
                .names = "{.col}_w")) |>
  mutate(across(ends_with("_w"), ~ log(.x + 1), .names = "ln_{.col}")) |>
  distinct(j5, year, .keep_all = TRUE)

# ── Site-year aggregation ─────────────────────────────────────────────────────
panel_site_ext <- panel_hh_ext |>
  group_by(obs, site_id, year, treated_unit, treatment) |>
  summarise(
    across(starts_with("ln_"), ~ mean(.x, na.rm = TRUE)),
    n_membres = mean(n_membres, na.rm = TRUE),
    n_hh      = n(),
    .groups   = "drop"
  )
```

### 2.1 Donor pool coverage

```{r tbl_donor_coverage}
#| tbl-cap: "Donor pool: sites by observatory and year coverage"

panel_site_ext |>
  group_by(obs) |>
  summarise(
    `Observatory` = obs_labels[first(obs)],
    `Sites`       = n_distinct(site_id),
    `First year`  = min(year),
    `Last year`   = max(year),
    `Pre-T waves` = n_distinct(year[year < TREAT_YEAR]),
    `Post-T waves`= n_distinct(year[year >= TREAT_YEAR]),
    `Treated`     = any(treated_unit),
    .groups = "drop"
  ) |>
  select(-obs) |>
  gt() |>
  tab_style(style = list(cell_fill(color = "#D4EDDA"), cell_text(weight = "bold")),
            locations = cells_body(rows = `Treated`)) |>
  tab_style(style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
            locations = cells_column_labels()) |>
  cols_hide(columns = `Treated`) |>
  tab_footnote(
    "Green rows: treated (Marovoay). gsynth uses all available site-year cells;
    sites with partial post-treatment coverage contribute to pre-treatment factor fit.",
    locations = cells_column_labels(`Observatory`)
  ) |>
  tab_source_note(md(
    paste0("Total sites in donor pool: ",
           n_distinct(panel_site_ext$site_id[!panel_site_ext$treated_unit]),
           "; treated sites: 2. 2005 Marovoay excluded.")
  ))
```

### 2.2 Panel coverage heatmap

```{r fig_panel_view}
#| fig-cap: "Panel coverage: site-years with observations (green = post-treatment)"
#| fig-height: 14
#| fig-width: 10
#| message: false
#| warning: false

panelview(
  ln_revtot_w ~ treatment,
  data  = as.data.frame(panel_site_ext),
  index = c("site_id", "year"),
  pre.post = TRUE,
  axis.lab.gap = 2,
  main = "Site-year coverage — extended donor pool"
)
```

---

## 3 gsynth: Interactive Fixed Effects with Extended Donor Pool

The gsynth IFE estimator fits:

$$Y_{it} = \mathbf{\lambda}_i' \mathbf{f}_t + D_{it} \delta + \varepsilon_{it}$$

where $\mathbf{\lambda}_i$ are unit-specific factor loadings, $\mathbf{f}_t$
are common time factors, and $\delta$ is the treatment effect (here allowed
to be heterogeneous by time since treatment via the event-study output). The
number of factors $r$ is selected by cross-validation.

Sites with partial post-treatment coverage contribute to estimating
$(\mathbf{\lambda}_i, \mathbf{f}_t)$ during the periods they are observed,
but their ATT contribution is limited to those periods. The synthetic control
for the treated sites is thus built on a much richer factor structure than
in the 4-donor specification.

```{r gsynth_ext_main}
#| message: false
#| warning: false
#| cache: true

run_gsynth_ext <- function(outcome, data, r_range = c(0, 3)) {
  d <- data |> filter(!is.na(.data[[outcome]]))
  suppressWarnings(
    gsynth(
      formula   = as.formula(paste0(outcome, " ~ treatment + n_membres")),
      data      = d,
      index     = c("site_id", "year"),
      se        = TRUE,
      inference = "nonparametric",
      nboots    = 500,
      r         = r_range,
      CV        = TRUE,
      force     = "two-way",
      parallel  = FALSE,
      min.T0    = 4,
      seed      = 2139
    )
  )
}

gs_ext_revtot    <- run_gsynth_ext("ln_revtot_w",    panel_site_ext)
gs_ext_revcou    <- run_gsynth_ext("ln_revcou_w",    panel_site_ext)
gs_ext_agri      <- run_gsynth_ext("ln_rev_agri_w",  panel_site_ext)
gs_ext_nonagri   <- run_gsynth_ext("ln_rev_nonagri_w", panel_site_ext)
```

### 3.1 Factors selected by cross-validation

```{r tbl_factors}
#| tbl-cap: "Number of interactive factors selected by CV (extended donor pool)"

tribble(
  ~Outcome, ~Factors,
  "Log total income",      gs_ext_revtot$r,
  "Log coupled income",    gs_ext_revcou$r,
  "Log agricultural rev.", gs_ext_agri$r,
  "Log non-agri rev.",     gs_ext_nonagri$r
) |>
  gt() |>
  tab_style(style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
            locations = cells_column_labels()) |>
  tab_source_note("CV selects r from {0,…,3}. r=0 corresponds to two-way FE (no interactive factors).")
```

### 3.2 Gap plots — all outcomes

```{r fig_gap_revtot}
#| fig-cap: "gsynth gap: log total income (extended donor pool, 500 bootstrap draws)"
#| fig-width: 9
#| fig-height: 4

plot(gs_ext_revtot, type = "gap",
     main  = "ATT over time: log total income",
     ylab  = "ATT (log points)",
     xlab  = "Year")
```

```{r fig_gap_income_components}
#| fig-cap: "gsynth gap: income components (extended donor pool)"
#| fig-width: 10
#| fig-height: 8
#| layout-ncol: 2

plot(gs_ext_revcou, type = "gap",
     main = "Log coupled income",    ylab = "ATT (log points)", xlab = "Year")
plot(gs_ext_agri,   type = "gap",
     main = "Log agricultural rev.", ylab = "ATT (log points)", xlab = "Year")
plot(gs_ext_nonagri, type = "gap",
     main = "Log non-agri rev.",     ylab = "ATT (log points)", xlab = "Year")
```

### 3.3 Counterfactual trajectories

```{r fig_cf_revtot}
#| fig-cap: "Observed vs. counterfactual — log total income (extended donor pool)"
#| fig-width: 9
#| fig-height: 4

plot(gs_ext_revtot, type = "counterfactual",
     main = "Observed vs. counterfactual: log total income",
     ylab = "Mean log income", xlab = "Year",
     id   = MARO_TREATED)
```

### 3.4 Pre-treatment fit diagnostics

The mean squared prediction error (MSPE) ratio compares post-treatment MSPE
to pre-treatment MSPE. A ratio >> 1 suggests a treatment effect; a ratio ≈ 1
or below suggests the post-treatment trajectory is no worse than the
pre-treatment fit.

```{r tbl_mspe}
#| tbl-cap: "Pre/post MSPE ratio (log total income, extended donor pool)"

compute_mspe_ratio <- function(gs, label) {
  rel_time <- gs$time                          # relative: negative = pre, non-negative = post
  att      <- gs$att
  att_t    <- att[rel_time >= 0]
  att_0    <- att[rel_time <  0]
  tibble(
    Outcome          = label,
    `Pre MSPE`       = round(mean(att_0^2, na.rm = TRUE), 5),
    `Post MSPE`      = round(mean(att_t^2, na.rm = TRUE), 5),
    `Post/Pre ratio` = round(mean(att_t^2, na.rm = TRUE) /
                               mean(att_0^2, na.rm = TRUE), 2)
  )
}

bind_rows(
  compute_mspe_ratio(gs_ext_revtot,  "Log total income"),
  compute_mspe_ratio(gs_ext_revcou,  "Log coupled income"),
  compute_mspe_ratio(gs_ext_agri,    "Log agricultural rev."),
  compute_mspe_ratio(gs_ext_nonagri, "Log non-agri rev.")
) |>
  gt() |>
  tab_style(style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
            locations = cells_column_labels()) |>
  tab_source_note(
    "Post/Pre MSPE ratio > 1 suggests a post-treatment gap larger than
    pre-treatment noise. Used as an informal effect-size measure alongside
    the bootstrap p-values."
  )
```

---

## 4 In-Space Placebo Tests

With the extended donor pool (~60 donor sites), we can run a much more
informative placebo test: assign treatment to each donor site in turn and
estimate a placebo ATT. The true treated-site p-value is the rank of its
MSPE ratio among all placebo MSPE ratios.

```{r in_space_placebo}
#| cache: true
#| message: false
#| warning: false
#| fig-cap: "In-space placebo test: distribution of post/pre MSPE ratios"
#| fig-width: 8
#| fig-height: 4

# Collect all non-treated sites as potential placebo units
placebo_sites <- panel_site_ext |>
  filter(!treated_unit) |>
  pull(site_id) |>
  unique()

# Get actual ATT for comparison
actual_att_site <- coef(
  feols(ln_revtot_w ~ treatment | site_id + year, data = panel_site_ext,
        warn = FALSE, notes = FALSE)
)[["treatment"]]

# Run TWFE placebo for all donor sites
placebo_results <- map_dfr(placebo_sites, function(s) {
  d <- panel_site_ext |>
    mutate(treat_pl = as.integer(site_id == s & year >= TREAT_YEAR))
  att_pl <- tryCatch(
    coef(feols(ln_revtot_w ~ treat_pl | site_id + year,
               data = d, warn = FALSE, notes = FALSE))[["treat_pl"]],
    error = function(e) NA_real_
  )
  if (length(att_pl) == 0) att_pl <- NA_real_
  tibble(site_id = s, att_pl = att_pl)
})

# Compute p-value rank
p_rank <- mean(abs(placebo_results$att_pl) >= abs(actual_att_site), na.rm = TRUE)

# Plot distribution
ggplot(placebo_results, aes(x = att_pl)) +
  geom_histogram(binwidth = 0.05, fill = "#74B3CE", color = "white", alpha = 0.8) +
  geom_vline(xintercept = actual_att_site, color = "#C0392B", linewidth = 1) +
  geom_vline(xintercept = -actual_att_site, color = "#C0392B",
             linewidth = 1, linetype = "dashed") +
  annotate("text", x = actual_att_site, y = Inf, hjust = -0.15, vjust = 1.4,
           label = paste0("Actual ATT\n= ", round(actual_att_site, 3)),
           color = "#C0392B", size = 3) +
  labs(
    title = "In-space placebo test: TWFE ATT distribution",
    subtitle = paste0(
      "Placebo p = ", round(p_rank, 3),
      "  (", sum(abs(placebo_results$att_pl) >= abs(actual_att_site), na.rm = TRUE),
      " of ", sum(!is.na(placebo_results$att_pl)), " placebos ≥ |actual ATT|)"
    ),
    x = "Placebo ATT (log points)",
    y = "Count"
  ) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))
```

---

## 5 Comparison: Restricted vs. Extended Donor Pool

```{r tbl_comparison}
#| tbl-cap: "gsynth ATT: restricted (4–5 donors) vs. extended (~62 donors) donor pool"

extract_gs <- function(gs, label) {
  att  <- tryCatch(gs$att.avg, error = function(e) NA_real_)
  # SE from att.avg.boot (vector of bootstrap ATT.avg draws)
  boot_avg <- tryCatch(gs$att.avg.boot, error = function(e) NULL)
  se_b <- if (!is.null(boot_avg)) sd(as.vector(boot_avg), na.rm = TRUE) else NA_real_
  tibble(
    Specification = label,
    ATT           = round(att, 4),
    `Bootstrap SE`= round(se_b, 4),
    `95% CI lo`   = round(att - 1.96 * se_b, 4),
    `95% CI hi`   = round(att + 1.96 * se_b, 4),
    Factors       = tryCatch(gs$r.cv, error = function(e) NA_integer_)
  )
}

# Run restricted spec for direct comparison
gs_restricted <- tryCatch(
  suppressWarnings(
    gsynth(
      ln_revtot_w ~ treatment + n_membres,
      data      = panel_site_ext |> filter(obs %in% c("03", "21")),
      index     = c("site_id", "year"),
      se        = TRUE, inference = "nonparametric", nboots = 500,
      r = 0, CV = FALSE, force = "two-way", parallel = FALSE,
      min.T0 = 4, seed = 2139
    )
  ),
  error = function(e) NULL
)

bind_rows(
  if (!is.null(gs_restricted))
    extract_gs(gs_restricted, "Restricted: Maro + Alaotra only (5 donors)")
  else
    tibble(Specification = "Restricted (failed)", ATT=NA, `Bootstrap SE`=NA,
           `95% CI lo`=NA, `95% CI hi`=NA, Factors=NA),
  extract_gs(gs_ext_revtot, "Extended: full donor pool (~62 donors)")
) |>
  gt() |>
  fmt_number(columns = c(ATT, `Bootstrap SE`, `95% CI lo`, `95% CI hi`), decimals = 3) |>
  tab_style(style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
            locations = cells_column_labels()) |>
  tab_style(style = cell_fill(color = "#FFF3CD"),
            locations = cells_body(rows = str_detect(Specification, "Extended"))) |>
  tab_source_note(md(
    "Outcome: log winsorised total income. Bootstrap SE from 500 draws.
    Period: 1999–2014 (2005 Marovoay excluded). Treated: sites 031 and 032."
  ))
```

---

## 6 Interpretation

```{r tbl_att_summary}
#| tbl-cap: "Annual ATT estimates — extended donor pool, log total income"

if (!is.null(gs_ext_revtot) && length(gs_ext_revtot$att) > 0) {
  rel_time <- gs_ext_revtot$time           # integer relative time: -n…m
  att_vec  <- gs_ext_revtot$att
  year_vec <- TREAT_YEAR + rel_time

  # Bootstrap SE from att.boot matrix [time × nboots]
  att_boot <- gs_ext_revtot$att.boot
  if (!is.null(att_boot) && nrow(att_boot) == length(att_vec)) {
    se_vec <- apply(att_boot, 1, sd, na.rm = TRUE)
  } else {
    se_vec <- rep(NA_real_, length(att_vec))
  }

  tibble(
    Year   = year_vec,
    ATT    = round(att_vec, 3),
    SE     = round(se_vec, 3),
    `95% CI lo` = round(att_vec - 1.96 * se_vec, 3),
    `95% CI hi` = round(att_vec + 1.96 * se_vec, 3)
  ) |>
    mutate(
      Period = if_else(Year < TREAT_YEAR, "Pre-treatment", "Post-treatment")
    ) |>
    gt(groupname_col = "Period") |>
    fmt_number(columns = c(ATT, SE, `95% CI lo`, `95% CI hi`), decimals = 3) |>
    tab_style(style = list(cell_text(weight = "bold"), cell_fill(color = "#DCDCDC")),
              locations = cells_column_labels()) |>
    tab_style(style = cell_fill(color = "#D4EDDA"),
              locations = cells_body(rows = Year >= TREAT_YEAR)) |>
    tab_source_note("SE from 500 nonparametric bootstrap draws. 2005 Marovoay excluded.")
}
```

### Summary

The extended donor pool analysis provides three improvements over the
restricted-donor gsynth in `06_Marovoay-Ankarafantsika-extended.qmd`:

1. **Factor model fit**: With ~60 donor sites, gsynth can identify latent
   time patterns (weather shocks, macro trends, rice price cycles) that are
   common across Madagascar's rural observatories. The treated sites' factor
   loadings are estimated more precisely against this richer background.

2. **In-space placebo p-value**: With ~60 donors, the minimum achievable
   p-value from the placebo distribution is ~1/62 ≈ 0.016, compared to 1/5 = 0.20
   with the restricted donor pool. A significant placebo p-value (< 0.05) is
   now achievable if the treatment effect is large.

3. **Pre-treatment MSPE**: A low post/pre MSPE ratio confirms that the
   synthetic control tracks the treated units closely before 2003, lending
   credibility to the parallel-trends assumption underlying the IFE model.

::: {.callout-note title="What this analysis cannot do"}
This analysis covers only 1999–2014. It cannot detect the long-run effect of
the PA regime as of 2025. That is covered by the long-run DiD in
`07_household_analysis.qmd`. The two analyses are **complementary**: the
medium-run gsynth provides a precise trajectory 2003–2014; the long-run DiD
provides the 23-year compound effect.
:::
