---
title: "Protected Areas and Rural Livelihoods: Long-Run Evidence from Madagascar"
subtitle: "Draft Manuscript"
author:
  - name: "[Author 1]"
    affiliation: "[Affiliation]"
  - name: "[Author 2]"
    affiliation: "[Affiliation]"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    fig-width: 8
    fig-height: 5
    theme: cosmo
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
bibliography: ../references.bib
---

```{r setup}
#| label: setup

suppressPackageStartupMessages({
  library(tidyverse)
  library(haven)
  library(fixest)
  library(gt)
  library(scales)
  library(knitr)
})

# ── Constants ──
MARO_TREATED  <- c("031", "032")
MARO_CONTROL  <- c("033", "034")
MARO_ALL      <- c(MARO_TREATED, MARO_CONTROL)
ALAOTRA_ALL   <- c("211", "212", "213")
ALL_SITES     <- c(MARO_ALL, ALAOTRA_ALL)
TREAT_YEAR    <- 2003
PRE_YEARS     <- 1999:2002
HIST_YEARS    <- 1999:2014
ANALYSIS_YEARS <- 1999:2014
DONOR_OBS     <- c("02", "12", "13", "15", "16", "21", "23", "24", "25", "41")

# ── Load ALL pre-computed results ──
# Heavy computations (gsynth bootstrap, wild bootstrap, permutation)
# were run in the replication QMDs (07, 08) and saved to data/.
# This manuscript loads results only — no expensive estimation at render.
res      <- read_rds("../data/manuscript_results.rds")
gs_res   <- read_rds("../data/gs_main_results.rds")
loo_data <- read_rds("../data/loo_gsynth_results.rds")
```

# Abstract {.unnumbered}

We estimate the causal effect of the 2002 reclassification of Ankarafantsika as a national park on household income in surrounding rural communities in Madagascar. The area had been protected since 1927, but the merger of pre-existing reserves into a single national park (≈130,000 ha) under Madagascar National Parks entailed a sharp tightening of enforcement. Combining a unique household panel (1999–2014) from the Réseau des Observatoires Ruraux — a network of up to 26 rural observatories across Madagascar — with a new 2025 resurvey, we employ two complementary identification strategies. First, a generalized synthetic control method (gsynth) exploiting 74 donor sites from 10 rural observatories identifies a medium-run average treatment effect on the treated of approximately −0.43 log points on total household income over 2003–2014, robust across a leave-one-out sensitivity analysis (SD = 0.023 across 74 exclusions). Second, a long-run repeated cross-section difference-in-differences comparing 1999–2002 baselines to 2025 outcomes estimates a decline of 0.43–0.75 log points, depending on the control group. Both estimates are imprecise given the small number of treated clusters (2 sites), and permutation inference and wild cluster bootstrap yield p-values of 0.19–0.53. The results suggest a persistent negative effect of the park on local incomes, with point estimates that are policy-relevant, but statistical power is inherently limited by the quasi-experimental setting.

# Introduction

Protected areas are a cornerstone of global biodiversity conservation policy. Under the 2022 Kunming-Montreal Global Biodiversity Framework, signatories have committed to protecting 30% of land and sea areas by 2030. Madagascar, one of the world's foremost biodiversity hotspots, has been at the forefront of this expansion: the country's protected area network tripled between 2003 and 2015, growing from approximately 1.7 million to over 7 million hectares [@Vieilledent2018].

Yet the welfare consequences for communities living near these protected areas remain contested. Conservation interventions restrict access to forests, fisheries, and agricultural land—resources on which rural households in developing countries depend for their livelihoods. Whether the benefits of conservation (ecosystem services, ecotourism revenues, transfers from conservation organizations) offset these costs is an empirical question with profound implications for both conservation policy and poverty reduction.

This paper asks: What is the long-run causal effect of the Ankarafantsika National Park reclassification on the income of surrounding rural households? The area, located in northwestern Madagascar, has been under protection since 1927 — first as a strict nature reserve (Réserve Naturelle Intégrale, 66,541 ha), then alongside adjacent forest reserves classified in 1929. In 2002, these pre-existing protected areas were merged into a single national park (≈130,000 ha) under Madagascar National Parks (formerly ANGAP), with a sharp tightening of enforcement. We exploit this policy change alongside a unique longitudinal dataset to estimate its impact.

Our setting offers several advantages for causal inference. The Réseau des Observatoires Ruraux (ROR), a network of rural observatories operated continuously from 1995 to 2014, provides standardized annual household surveys across multiple sites in Madagascar. Two observatory sites adjacent to Ankarafantsika (Marovoay) were directly affected by the reclassification; two nearby control sites from the same observatory were not. Moreover, the ROR covers 10 additional observatories across Madagascar that serve as a broad donor pool for synthetic control methods. A new 2025 resurvey of the original Marovoay and Lac Alaotra sites extends the time horizon to more than two decades after treatment.

We employ two complementary empirical strategies:

1. **Generalized synthetic control** (gsynth; @Xu2017) using the 1999–2014 panel, which leverages 74 donor sites from 10 observatories to construct counterfactual income trajectories for treated sites via an interactive fixed-effects model. This approach relaxes the parallel trends assumption by allowing for unobserved time-varying confounders captured by latent factors.

2. **Long-run repeated cross-section difference-in-differences** comparing pre-treatment (1999–2002) outcomes to those measured in the 2025 resurvey. With only 4 to 7 sites, we rely on permutation inference following @Conley2011 and wild cluster bootstrap following @Cameron2008.

Our main findings are as follows. The gsynth estimator identifies a medium-run average treatment effect on the treated (ATT) of approximately −0.43 log points on total household income over the 2003–2014 post-treatment period, driven primarily by declines in non-agricultural income. This estimate is highly stable across a leave-one-out donor exclusion exercise: excluding each of the 74 donor sites one at a time, the ATT ranges from −0.49 to −0.35 (SD = 0.023). The long-run DiD yields ATT estimates of −0.43 to −0.75 log points depending on whether control sites are limited to Marovoay or extended to include the Lac Alaotra observatory. Both approaches suggest a persistent negative income effect, though individual estimates are imprecise given the small number of treated clusters.

This paper contributes to three literatures. First, it adds to the growing body of quasi-experimental evidence on the welfare impacts of protected areas in developing countries [see @Ferraro2011; @Sims2010; @Andam2010], which has predominantly relied on matching methods or short-run evaluations. We provide among the longest-horizon causal estimates in this literature. Second, it contributes methodologically by combining generalized synthetic control methods [@Xu2017] with an extended donor pool approach, demonstrating their applicability in a developing-country panel setting with multiple study areas. Third, it provides evidence on a specific, policy-relevant case—Ankarafantsika National Park in Madagascar—where no rigorous impact evaluation has previously been conducted.

The remainder of this paper is organized as follows. @sec-background provides background on the Ankarafantsika National Park and the ROR data infrastructure. @sec-data describes the data and presents descriptive statistics. @sec-empirical lays out the two empirical strategies. @sec-results presents the results, including robustness checks and sensitivity analyses. @sec-conclusion concludes with policy implications and directions for future research.

# Background {#sec-background}

## Ankarafantsika National Park

Ankarafantsika is located in the Boeny region of northwestern Madagascar, approximately 450 km from the capital, Antananarivo. The area’s protection history extends back nearly a century. In 1927, a strict nature reserve (Réserve Naturelle Intégrale, RNI) of 66,541 hectares was established. In 1929, adjacent forest reserves covering approximately 75,000 hectares were classified. These early designations were reclassified in 1966 (Décret n° 66-242), and in 1999 the Station Forestière d’Ampijoroa (4,974 ha) was created by inter-ministerial order. Despite formal protection, enforcement remained largely ineffective throughout this period: subsistence agriculture, cattle grazing, and forest product harvesting continued within and around the reserves [@Goodman2018].

The decisive policy change came in 2002 with Décret n° 2002-798 (7 August 2002), which merged the RNI, the forest reserves, and the Ampijoroa station into a single national park of approximately 130,000 hectares under the management of Madagascar National Parks (formerly ANGAP). This reclassification entailed a significant tightening of land-use restrictions. While the earlier reserve and forest station statuses had been only loosely enforced — allowing continued subsistence agriculture, cattle grazing, and forest product harvesting — the national park regime imposed strict prohibitions on these activities. Slash-and-burn agriculture (*tavy*), a traditional farming practice in the region, was explicitly banned within and adjacent to the park. Enforcement was supported by international conservation funding, particularly from the World Bank and bilateral donors as part of Madagascar's third Environmental Action Plan.

The communities most directly affected by the stricter enforcement are located in the Marovoay rural observatory sites (sites 031 and 032 in the ROR classification), which border the park. Two additional observatory sites in the broader Marovoay area (sites 033 and 034) are geographically more distant and were not directly subject to the same access restrictions.

## The Réseau des Observatoires Ruraux (ROR)

The ROR was established in 1995 by a Franco-Malagasy consortium (IRD, INSTAT, and the Malagasy Ministry of Agriculture) as a longitudinal monitoring system for rural livelihoods in Madagascar. At its peak, the network encompassed 26 observatories covering diverse agro-ecological zones, from the humid eastern forests to the semi-arid south, although many were not surveyed consistently enough for panel analysis. Each observatory tracked approximately 500 households annually through a standardized questionnaire covering demographics, agriculture, income, consumption, and asset ownership.

Key features of the ROR for this study include:

- **Annual frequency** from 1995 to 2014 (with a gap in 2005 for the Marovoay sites due to funding interruptions), providing a rich pre-treatment baseline (1999–2002) and extended post-treatment series (2003–2014).
- **Standardized income measurement** across all observatories, allowing direct comparison of income trajectories between treated and control sites.
- **Broad geographic coverage**, providing donor sites for synthetic control methods that span different agricultural systems and regions of Madagascar.

In 2025, a resurvey was conducted in the original Marovoay and Lac Alaotra observatory areas, following the same questionnaire structure, enabling long-run (22-year) comparisons with the pre-treatment period.

# Data and Descriptive Statistics {#sec-data}

```{r load-data}
#| label: load-data

# Load consolidated household panel
household_consolidated <- read_rds("../data/household_consolidated.rds")

# ── Income construction ──
mk_income_ext <- function(df) {
  df |>
    mutate(
      rev_riz  = coalesce(rev_riz,  0),
      rev_cu   = coalesce(rev_cu,   0),
      revel    = coalesce(revel,    0),
      revpeche = coalesce(revpeche, 0),
      revppal  = coalesce(revppal,  0),
      revsec   = coalesce(revsec,   0),
      revcou   = coalesce(revcou, revppal + revsec + rev_riz + rev_cu + revel + revpeche),
      revtot   = coalesce(revtot, revcou),
      rev_agri    = rev_riz + rev_cu + revel + revpeche,
      rev_nonagri = revppal + revsec
    )
}

site_treatment <- function(site_id, year) {
  as.integer(site_id %in% MARO_TREATED & year >= TREAT_YEAR)
}

# ── Household-level panel for gsynth (all observatories) ──
panel_hh_ext <- household_consolidated |>
  filter(
    obs %in% c("03", DONOR_OBS),
    year %in% ANALYSIS_YEARS,
    !(obs == "03" & year == 2005)
  ) |>
  mk_income_ext() |>
  mutate(
    treated_unit = site_id %in% MARO_TREATED,
    treatment    = site_treatment(site_id, year)
  ) |>
  mutate(across(
    c(revtot, revcou, rev_agri, rev_nonagri),
    ~ pmin(.x, quantile(.x, 0.99, na.rm = TRUE)),
    .names = "{.col}_w"
  )) |>
  mutate(across(ends_with("_w"), ~ log(pmax(.x, 0) + 1), .names = "ln_{.col}")) |>
  distinct(j5, year, .keep_all = TRUE)

# ── Site-level aggregation ──
panel_site_ext <- panel_hh_ext |>
  group_by(obs, site_id, year, treated_unit, treatment) |>
  summarise(
    across(starts_with("ln_"), ~ mean(.x, na.rm = TRUE)),
    n_membres = mean(n_membres, na.rm = TRUE),
    n_hh = n(),
    .groups = "drop"
  )

n_treated <- panel_site_ext |> filter(treated_unit) |> distinct(site_id) |> nrow()
n_donors  <- panel_site_ext |> filter(!treated_unit) |> distinct(site_id) |> nrow()
n_obs_total <- panel_site_ext |> distinct(obs) |> nrow()
```

## Sample construction

We construct two analysis panels from the ROR microdata:

1. **Medium-run site-level panel (1999–2014):** We aggregate household-level income data to the site level, yielding `r n_treated` treated sites and `r n_donors` donor sites from `r n_obs_total` observatories. The panel is unbalanced due to the 2005 survey gap in Marovoay.

2. **Long-run household-level panel (1999–2002 vs. 2025):** We pool the pre-treatment waves with the 2025 resurvey for a repeated cross-section design, using 4 sites (Marovoay only) or 7 sites (Marovoay + Lac Alaotra).

Income variables are constructed following the standard ROR methodology: total income (*revtot*) comprises agricultural income (*rev_agri*: rice, cash crops, livestock, fishing) and non-agricultural income (*rev_nonagri*: primary and secondary employment). All income measures are winsorized at the 99th percentile and log-transformed.

```{r desc-stats}
#| label: tbl-desc-stats
#| tbl-cap: "Descriptive statistics: site-level panel, 1999–2014"

panel_site_ext |>
  mutate(group = if_else(treated_unit, "Treated (031, 032)", "Donor pool")) |>
  group_by(group) |>
  summarise(
    `Sites`         = n_distinct(site_id),
    `Site-years`    = n(),
    `Mean ln(total income)` = round(mean(ln_revtot_w, na.rm = TRUE), 3),
    `SD ln(total income)`   = round(sd(ln_revtot_w, na.rm = TRUE), 3),
    `Mean HH per site`      = round(mean(n_hh, na.rm = TRUE), 1),
    .groups = "drop"
  ) |>
  gt() |>
  tab_header(title = "Summary Statistics by Treatment Status") |>
  cols_label(group = "Group")
```

```{r desc-trends}
#| label: fig-trends
#| fig-cap: "Mean log total income by treatment status, 1999–2014"

panel_site_ext |>
  mutate(group = if_else(treated_unit, "Treated", "Donor pool")) |>
  group_by(group, year) |>
  summarise(mean_y = mean(ln_revtot_w, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = year, y = mean_y, color = group)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_vline(xintercept = TREAT_YEAR - 0.5, linetype = "dashed", color = "grey40") +
  annotate("text", x = TREAT_YEAR - 0.3, y = Inf, label = "Park reclassified",
           hjust = 0, vjust = 1.5, color = "grey40", size = 3.5) +
  scale_color_manual(values = c("Treated" = "#D55E00", "Donor pool" = "#0072B2")) +
  labs(x = "Year", y = "Mean ln(total income)", color = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

@fig-trends displays the mean log total income trajectories for treated and donor-pool sites. The pre-treatment trends (1999–2002) are broadly parallel, though the treated sites start from a somewhat higher income level. A visible divergence emerges after 2003, with treated sites' incomes declining relative to the donor pool.

# Empirical Framework {#sec-empirical}

## Strategy 1: Generalized Synthetic Control (Medium Run)

For the medium-run analysis (1999–2014), we employ the generalized synthetic control (gsynth) method of @Xu2017, which extends the synthetic control framework of @Abadie2010 by estimating an interactive fixed-effects (IFE) model:

$$
Y_{it} = \alpha_i + \xi_t + \boldsymbol{\lambda}_i' \boldsymbol{f}_t + D_{it} \cdot \delta_{it} + \varepsilon_{it}
$$

where $Y_{it}$ is log total income for site $i$ in year $t$; $\alpha_i$ and $\xi_t$ are unit and time fixed effects; $\boldsymbol{\lambda}_i$ is a vector of unit-specific factor loadings and $\boldsymbol{f}_t$ the corresponding common factors; $D_{it}$ is a binary treatment indicator equal to 1 for treated sites post-2003; and $\delta_{it}$ is the treatment effect.

The method proceeds in three steps:

1. Estimate the IFE model on the untreated units to obtain $\hat{\boldsymbol{f}}_t$;
2. Estimate treated units' factor loadings $\hat{\boldsymbol{\lambda}}_i$ using pre-treatment data;
3. Construct counterfactuals $\hat{Y}_{it}(0) = \hat{\alpha}_i + \hat{\xi}_t + \hat{\boldsymbol{\lambda}}_i' \hat{\boldsymbol{f}}_t$ and compute $\hat{\delta}_{it} = Y_{it} - \hat{Y}_{it}(0)$.

The number of factors $r$ is selected by cross-validation over $r \in \{0, 1, 2, 3\}$, minimizing mean squared prediction error (MSPE). Inference is conducted via parametric bootstrap with 200 replications.

**Donor pool.** We use all `r n_donors` sites from the `r n_obs_total` ROR observatories that are not treated by the park as potential donors. This extended donor pool provides substantially more variation for factor estimation than using only the local Marovoay controls.

**Minimum pre-treatment periods.** Because treated sites have only 4 pre-treatment years (1999–2002), we set `min.T0 = 3` to prevent their automatic exclusion by the algorithm.

## Strategy 2: Long-Run Repeated Cross-Section DiD

For the long-run analysis, we pool the 1999–2002 pre-treatment waves with the 2025 resurvey to estimate a standard two-period DiD:

$$
Y_{ist} = \alpha + \gamma_s + \mu \cdot \text{Post}_t + \delta \cdot (\text{Treated}_s \times \text{Post}_t) + \varepsilon_{ist}
$$

where $Y_{ist}$ is log total income for household $i$ in site $s$ at time $t$; $\gamma_s$ are site fixed effects; $\text{Post}_t$ indicates the 2025 wave; and $\delta$ is the long-run ATT.

We estimate this specification with two control groups:

- **Marovoay only** (4 sites: 031, 032 treated; 033, 034 control), yielding the most locally comparable counterfactual.
- **Marovoay + Lac Alaotra** (7 sites), adding three rice-growing observatory sites from a different region to improve statistical power.

## Inference with Few Clusters

Both strategies confront the challenge of inference with very few treated clusters (2 sites). Standard cluster-robust standard errors are unreliable in this setting [@Cameron2008]. We address this through three approaches:

1. **Permutation inference** [@Conley2011]: We enumerate all possible treatment allocations among the available sites (${4 \choose 2} = 6$ for Marovoay; ${7 \choose 2} = 21$ for the extended sample) and compare the actual ATT to the empirical distribution of placebo ATTs.

2. **Wild cluster bootstrap** [@Cameron2008]: We implement a Rademacher wild bootstrap (B = 999) at the site level on the site-aggregated regressions.

3. **Parametric bootstrap** (gsynth): The gsynth framework provides bootstrap-based confidence intervals that do not rely on cluster asymptotics.

# Results {#sec-results}

## Medium-Run gsynth Results (1999–2014)

```{r gsynth-main}
#| label: gsynth-main

# gsynth results loaded from pre-computed RDS in setup chunk.
# See 08_gsynth_extended_donor.qmd for the full estimation code.
# gs_res contains: att_avg, boot_se, boot_ci, r_cv, est_att, att_boot
```

```{r gsynth-results-table}
#| label: tbl-gsynth
#| tbl-cap: "Generalized synthetic control estimates (site-level, 1999–2014)"

tibble(
  Outcome        = "ln(total income)",
  `ATT (avg)`    = round(gs_res$att_avg, 3),
  `SE (boot)`    = round(gs_res$boot_se, 3),
  `95% CI lower` = round(gs_res$boot_ci[1], 3),
  `95% CI upper` = round(gs_res$boot_ci[2], 3),
  `r* (CV)`      = gs_res$r_cv,
  `N donors`     = n_donors
) |>
  gt() |>
  tab_header(title = "gsynth Average Treatment Effect on the Treated")
```

Cross-validation selects $r^* = `r gs_res$r_cv`$ latent factor(s). The average post-treatment ATT on log total income is `r round(gs_res$att_avg, 3)` (bootstrap SE = `r round(gs_res$boot_se, 3)`), suggesting that households near the park experienced approximately a `r round(abs(gs_res$att_avg) * 100, 0)`% decline in total income relative to the synthetic counterfactual.

```{r gsynth-att-path}
#| label: fig-att-path
#| fig-cap: "gsynth ATT by relative time period"

att_est <- gs_res$est_att
rel_t <- as.numeric(rownames(att_est))

att_plot <- tibble(
  rel_time = rel_t,
  att      = att_est[, "ATT"],
  ci_lo    = att_est[, "CI.lower"],
  ci_hi    = att_est[, "CI.upper"]
)

ggplot(att_plot, aes(x = rel_time, y = att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "steelblue", alpha = 0.2) +
  geom_line(color = "steelblue", linewidth = 0.8) +
  geom_point(color = "steelblue", size = 2) +
  labs(
    x = "Years relative to park reclassification",
    y = "ATT (log points)"
  ) +
  theme_minimal(base_size = 12)
```

@fig-att-path shows the ATT trajectory. Pre-treatment effects (periods −3 to −1) are close to zero, supporting the validity of the counterfactual. The effect becomes negative starting around period 1 (2004) and deepens over time, reaching approximately −0.5 to −0.6 log points by periods 9–12 (2012–2014).

## Long-Run DiD Results (1999–2002 vs. 2025)

<!-- Long-run panels and DiD fits pre-computed in 07_household_analysis.qmd -->

```{r lr-did-results}
#| label: tbl-lr-did
#| tbl-cap: "Long-run DiD estimates (1999–2002 vs. 2025)"

# Pre-computed in 07_household_analysis.qmd
res$lr_table |>
  gt() |>
  tab_header(title = "Long-Run DiD: Average Treatment Effect on the Treated") |>
  tab_footnote("Cluster-robust SEs at site level. Site and period FE included.")
```

The Marovoay-only specification yields an ATT of `r res$lr_table$ATT[1]` (SE = `r res$lr_table$SE[1]`) for total income. The extended specification with Alaotra controls gives `r res$lr_table$ATT[4]` (SE = `r res$lr_table$SE[4]`). Non-agricultural income effects are positive but imprecise; agricultural income effects vary in sign across specifications.

## Robustness and Sensitivity

### Permutation Inference

```{r permutation}
#| label: tbl-permutation
#| tbl-cap: "Permutation inference for long-run DiD (site-aggregated)"

# Pre-computed in 07_household_analysis.qmd
tibble(
  Specification = c("Marovoay (4 sites)", "Maro + Alaotra (7 sites)"),
  `Actual ATT` = round(c(res$actual_maro, res$actual_full), 3),
  `N permutations` = c(res$n_perm_maro, res$n_perm_full),
  `p-value (two-sided)` = round(c(res$p_maro, res$p_full), 3)
) |>
  gt() |>
  tab_header(title = "Permutation Inference (Site-Aggregated)")
```

With only ${4 \choose 2} = 6$ possible allocations in Marovoay and ${7 \choose 2} = 21$ in the extended sample, permutation p-values are inherently coarse. The two-sided p-values are `r round(res$p_maro, 3)` and `r round(res$p_full, 3)`, respectively. While neither achieves conventional significance, the Marovoay ATT ranks as one of the most extreme among all placebo allocations.

### Wild Cluster Bootstrap

```{r wild-boot}
#| label: tbl-wild-boot
#| tbl-cap: "Wild cluster bootstrap p-values (Rademacher, B = 999)"

# Pre-computed in 07_household_analysis.qmd
tibble(
  Specification = c("Marovoay (4 sites)", "Maro + Alaotra (7 sites)"),
  ATT = round(c(res$wb_maro_beta, res$wb_full_beta), 3),
  `Boot SE` = round(c(res$wb_maro_se, res$wb_full_se), 3),
  `p (wild boot)` = round(c(res$wb_maro_p, res$wb_full_p), 3)
) |>
  gt() |>
  tab_header(title = "Wild Cluster Bootstrap Inference")
```

The wild cluster bootstrap p-values (`r round(res$wb_maro_p, 3)` and `r round(res$wb_full_p, 3)`) confirm that the estimates do not achieve conventional statistical significance, consistent with the limited number of clusters.

### Leave-One-Out Donor Sensitivity (gsynth)

```{r loo-gsynth}
#| label: fig-loo
#| fig-cap: "Leave-one-out sensitivity: distribution of gsynth ATT across 74 donor exclusions"

loo_ok <- loo_data |> filter(conv)
baseline_att <- gs_res$att_avg

ggplot(loo_ok, aes(x = att_avg)) +
  geom_histogram(aes(y = after_stat(density)), bins = 20,
                 fill = "steelblue", alpha = 0.7, color = "white") +
  geom_density(linewidth = 0.8, color = "navy") +
  geom_vline(xintercept = baseline_att, linetype = "dashed",
             color = "#D55E00", linewidth = 0.8) +
  annotate("text", x = baseline_att + 0.005, y = 12,
           label = paste0("Baseline\nATT = ", round(baseline_att, 3)),
           color = "#D55E00", hjust = 0, size = 3.5) +
  labs(
    x = "ATT (average post-treatment)",
    y = "Density",
    subtitle = paste0("74 exclusions — Mean = ", round(mean(loo_ok$att_avg), 3),
                      ", SD = ", round(sd(loo_ok$att_avg), 3))
  ) +
  theme_minimal(base_size = 12)
```

To assess whether the gsynth result is driven by any single donor site, we re-estimate the model 74 times, each time excluding one donor site. @fig-loo displays the resulting distribution of ATT estimates. The mean across exclusions is `r round(mean(loo_ok$att_avg), 3)` with a standard deviation of only `r round(sd(loo_ok$att_avg), 3)`, and the range is [`r round(min(loo_ok$att_avg), 3)`, `r round(max(loo_ok$att_avg), 3)`]. Cross-validation selects $r^* = 1$ in `r sum(loo_ok$r_cv == 1)` of 74 exclusions. The gsynth estimate is not sensitive to the inclusion or exclusion of any individual donor.

### Summary of Results

```{r summary-table}
#| label: tbl-summary
#| tbl-cap: "Summary of ATT estimates across methods"

summary_tbl <- tibble(
  Method = c(
    "gsynth (site-level, 1999–2014)",
    "gsynth LOO mean",
    "Long-run DiD, Maro only (hh-level)",
    "Long-run DiD, extended (hh-level)"
  ),
  `ATT (log pts)` = c(
    round(gs_res$att_avg, 3),
    round(mean(loo_ok$att_avg), 3),
    res$lr_table$ATT[1],
    res$lr_table$ATT[4]
  ),
  SE = c(
    round(gs_res$boot_se, 3),
    paste0("[", round(min(loo_ok$att_avg), 3), ", ", round(max(loo_ok$att_avg), 3), "]"),
    res$lr_table$SE[1],
    res$lr_table$SE[4]
  ),
  `Inference` = c(
    "Parametric boot",
    "Range across 74 exclusions",
    paste0("Perm p = ", round(res$p_maro, 3)),
    paste0("Perm p = ", round(res$p_full, 3))
  ),
  Horizon = c("2003–2014", "2003–2014", "2025 vs 1999–2002", "2025 vs 1999–2002")
)

summary_tbl |>
  gt() |>
  tab_header(title = "Summary of Treatment Effect Estimates") |>
  tab_footnote("All outcomes are log total household income (winsorized).")
```

## Limitations

Several limitations merit discussion.

**Internal validity.** The primary threat to identification is that treated and control sites may have been on different trajectories for reasons unrelated to the park. The gsynth pre-treatment ATTs (close to zero for periods −3 to −1) provide some reassurance, but with only 3–4 pre-treatment periods for the treated units, this is a weak test. The treatment assignment is not random: sites 031 and 032 border the park because of their geographic location, which may also correlate with unobserved determinants of income trajectories.

**Statistical power.** With only 2 treated sites, conventional inference is severely limited. None of our estimates achieve statistical significance at conventional levels under permutation or wild cluster bootstrap. The point estimates are policy-relevant in magnitude, but we cannot rule out zero effects.

**External validity.** Our results apply specifically to the communities surrounding Ankarafantsika National Park. Generalization to other protected areas in Madagascar or elsewhere depends on the similarity of the institutional context, enforcement regime, and alternative livelihood opportunities.

**Control group contamination.** The three Lac Alaotra control sites used to extend the DiD sample are themselves 9–16 km from protected areas designated in 2005–2007 (Lac Alaotra Paysage Harmonieux Protégé and Corridor Ankeniheny-Zahamena). While these PAs post-date the Ankarafantsika reclassification and do not threaten the medium-run gsynth analysis (1999–2014), they may compromise the Alaotra sites’ status as pure controls for the long-run 2025 DiD. We address this by reporting Marovoay-only estimates alongside the extended sample and by planning to reactivate a third observatory unaffected by recent PA designations.

**Measurement.** Income is self-reported and subject to the usual measurement concerns in developing-country household surveys. The 2005 survey gap for the Marovoay sites introduces a missing year in the panel, though the gsynth method handles unbalanced panels.

# Conclusion {#sec-conclusion}

This paper provides quasi-experimental evidence on the long-run effect of Ankarafantsika National Park on the incomes of surrounding rural households in Madagascar. Combining generalized synthetic control methods with an extended donor pool from 10 rural observatories, we estimate a medium-run (2003–2014) ATT of approximately −0.43 log points, which is highly robust to leave-one-out donor exclusion. Long-run DiD estimates using a 2025 resurvey yield similar magnitudes (−0.43 to −0.75 log points). Both approaches point to a substantial and persistent negative income effect, though estimates are imprecise due to the small number of treated clusters.

These findings carry implications for conservation policy. As countries scale up protected areas toward the target of protecting 30% of land and sea by 2030, rigorous evidence on the costs borne by adjacent communities is essential. If the park reduces household income by the magnitudes we estimate, this represents a significant hidden cost of biodiversity conservation borne by some of Madagascar's poorest communities. Effective compensation mechanisms—direct payments, alternative livelihood programs, or benefit-sharing arrangements from ecotourism—appear warranted but are currently insufficient to offset income losses. A companion national-level analysis of 71 post-2008 protected areas across Madagascar (Razafimahenina et al.) finds no significant average effect on household wealth using DHS/MICS data, with a negative signal emerging only beyond ten years. Our case study complements this broad finding by documenting what happens at a specific PA where the enforcement shift was particularly marked, over a much longer horizon.

At the same time, the limitations of our study point to important directions for future research. We have identified intermediate survey rounds conducted in 2017 (Marovoay, Antsirabe) and 2019 (Marovoay, Alaotra, Androy) whose data we are working to recover; these would strengthen the time-series dimension between 2014 and 2025. Finer-grained data on the mechanisms through which the park affects incomes—restrictions on agricultural land, reduced access to forest products, or changes in labor allocation—would allow better-targeted policy responses. The reactivation of a third observatory in 2026 will increase the number of control clusters and improve statistical power. Finally, incorporating environmental outcomes (deforestation rates, biodiversity indicators) alongside welfare measures would enable a more complete cost-benefit assessment of conservation policy.

# References {.unnumbered}

::: {#refs}
:::

---

# Appendix {.appendix}

## A. Replication Package

The analysis in this paper is supported by the following replication files:

| Document | Description |
|----------|-------------|
| `06_Marovoay-Ankarafantsika-extended.qmd` | Extended econometric analysis: 5 estimators (TWFE, event study, C&S, SDID, augsynth, gsynth) on Marovoay-only panel |
| `07_household_analysis.qmd` | Long-run RC DiD: 1999–2002 vs. 2025, permutation and wild bootstrap inference |
| `08_gsynth_extended_donor.qmd` | gsynth with extended donor pool (74 sites, 10 observatories) |
| `09_paper_manuscript.qmd` | This manuscript (loads pre-computed results from replication files) |
| `utils/calc_incomes_all_years.R` | Income construction functions for all survey years |

## B. Additional Results

```{r appendix-loo-table}
#| label: tbl-loo-detail
#| tbl-cap: "Leave-one-out gsynth: most and least influential donors"

bind_rows(
  loo_ok |> arrange(att_avg) |> head(5) |> mutate(group = "5 most negative"),
  loo_ok |> arrange(desc(att_avg)) |> head(5) |> mutate(group = "5 least negative")
) |>
  select(group, site, att_avg, att_se, r_cv) |>
  gt() |>
  tab_header(title = "Most and Least Influential Donor Sites") |>
  fmt_number(columns = c(att_avg, att_se), decimals = 3) |>
  cols_label(
    group = "Group", site = "Excluded Site",
    att_avg = "ATT (avg)", att_se = "SE (boot)", r_cv = "r*"
  )
```
