---
title: "Extensions"
subtitle: "Three additional observatory–PA pairs"
---

```{r}
#| label: setup
source("R/load_data.R")
source("R/gsynth_helpers.R")
source("R/plot_themes.R")
library(tmap)
```

The two-case comparison in [Chapter 4](04_results.qmd) — strict enforcement at Ankarafantsika versus participatory conservation at Lac Alaotra — limits external validity. This chapter extends the analysis to three additional rural observatories that the donor-pool audit ([Appendix B](A2_donor_validity.qmd)) flagged as themselves adjacent to protected areas created during the study period:

- Farafangana (obs. 15): adjacent to the Corridor Forestier Ambositra-Vondrozo (IUCN V, September 2006), a large highland corridor under light-touch regulation. Survey window: 1999--2008 (3 post-treatment years only).
- Toliara North (obs. 23): adjacent to Amoron'i Onilahy (IUCN V, January 2007), Mikea National Park (IUCN II, April 2007), and Ranobe PK 32 (IUCN V, December 2008) — a rapid cascade of coastal and dry-forest protections. Survey window: 1999--2014.
- Fénérive East (obs. 24): adjacent to the Réserve de Tampolo (IUCN IV, August 2007), a small (~7.3 km²) research forest managed by ESSA-Forêts. Survey window: 1999--2014.

These three observatories move from liabilities in the donor pool to informative exposed units. Together with Marovoay and Alaotra, they enable a five-case examination of how conservation governance type, enforcement intensity, and PA size shape household-income outcomes.

A methodological note: the donor pool in this chapter is restricted to five clean observatories (Antsirabe, Tsiroanomandidy, Ambovombe, Mahanoro, Itasy) as identified in [Appendix B](A2_donor_validity.qmd). The four also-exposed donors used in [Chapter 4](04_results.qmd) are now treated as exposed units here. Marovoay and Alaotra ATT estimates will therefore differ slightly from their Chapter 4 counterparts.


# Study sites {#sec-additional-sites}

## Farafangana and the Corridor Forestier Ambositra-Vondrozo {#sec-fara-site}

The CFAV (\~300 km, IUCN V) runs along the eastern highland escarpment. A temporary protection decree was issued on 15 September 2006 (WDPA 555697881) as part of the post-2003 SAPM expansion. Category V status implies land-use regulation at the forest margin --- primarily restrictions on slash-and-burn (*tavy*) and charcoal production in buffer zones --- rather than exclusionary enforcement.

The Farafangana observatory lies on the southeastern coastal plain; its nearest commune (Mahatsinjo) borders the CFAV at \~14 km. A complication: Manombo Special Reserve (IUCN IV, 1962) lies \~12 km from a second commune. Because Manombo predates the study period with unchanged boundaries and management, it constitutes a constant background condition absorbed into the unit fixed effect. The ATT for Farafangana therefore identifies the additional income effect of the 2006 CFAV decree.

All survey sites are exposed at observatory level from 2006; no within-observatory split is possible. With only 3 post-treatment years (2006--2008), gsynth estimates should be treated as illustrative.

```{r}
#| label: fig-map-fara
#| fig-cap: "Farafangana observatory and nearby protected areas. The CFAV (exposure PA, red) and Manombo SR (pre-existing, yellow) are shown over surveyed fokontany."
#| fig-height: 7
#| fig-width: 7

fara_pas <- dyn_study |>
  filter(str_detect(NAME, "Corridor Forestier Ambositra Vondrozo|Manombo")) |>
  mutate(pa_type = if_else(str_detect(NAME, "Corridor"),
                           "Exposure PA (CFAV, 2006)",
                           "Pre-existing PA (Manombo, 1962)"))

fara_fkt_bb <- st_bbox(fara_fokontany)
fara_bbox <- st_bbox(c(
  xmin = unname(fara_fkt_bb["xmin"]) - 0.3,
  ymin = unname(fara_fkt_bb["ymin"]) - 0.3,
  xmax = unname(fara_fkt_bb["xmax"]) + 0.3,
  ymax = unname(fara_fkt_bb["ymax"]) + 0.3
), crs = 4326)

fara_mad_c <- st_crop(madagascar, fara_bbox)
fara_pas_c <- tryCatch(st_crop(fara_pas, fara_bbox), error = function(e) fara_pas)
fara_fkt_c <- fara_fokontany |>
  mutate(layer = "Surveyed fokontany")

tm_shape(fara_mad_c) +
  tm_borders(col = "grey75", lwd = 0.3) +
  tm_fill(fill = "grey96") +
tm_shape(fara_pas_c) +
  tm_fill(fill = "pa_type",
          fill.scale = tm_scale_categorical(
            values = c("Exposure PA (CFAV, 2006)" = "#EF9A9A",
                       "Pre-existing PA (Manombo, 1962)" = "#FFF59D")),
          fill.legend = tm_legend(title = ""),
          fill_alpha = 0.5) +
  tm_borders(col = "pa_type",
             col.scale = tm_scale_categorical(
               values = c("Exposure PA (CFAV, 2006)" = "#D32F2F",
                          "Pre-existing PA (Manombo, 1962)" = "#F9A825")),
             col.legend = tm_legend(show = FALSE),
             lwd = 0.8) +
tm_shape(fara_fkt_c) +
  tm_fill(fill = "layer",
          fill.scale = tm_scale_categorical(
            values = c("Surveyed fokontany" = "#E8F5E9")),
          fill.legend = tm_legend(title = ""),
          fill_alpha = 0.7) +
  tm_borders(col = "grey40", lwd = 0.4) +
tm_scalebar(position = c("left", "bottom")) +
tm_title("Farafangana & Nearby PAs") +
tm_layout(legend.position = c("right", "top"))
```

## Toliara North and the coastal PA cascade {#sec-tol-site}

Observatory 23 sits on Madagascar's arid southwest coast, a landscape of spiny thicket and dry deciduous forest. Three PAs became operative in rapid succession:

1. Amoron'i Onilahy (IUCN V, 17 January 2007, ~12 km): riparian and coastal landscape PA covering the Onilahy estuary.
2. Mikea National Park (IUCN II, 13 April 2007, ~16 km): strict reserve protecting Mikea dry spiny-thicket forest.
3. Ranobe PK 32 (IUCN V, 2 December 2008, ~1.4 km): large terrestrial PA (1,685 km²) bordering the observatory at extremely close range.

Exposure year is {{< var treat_year_tol >}} — the first full survey year following the initial January 2007 decree. The December 2008 Ranobe decree constitutes a secondary intensification within the post-exposure window. The dominant conservation character is IUCN V / multipurpose, though the presence of Mikea NP (IUCN II) at 16 km introduces a strict element.

```{r}
#| label: fig-map-tol
#| fig-cap: "Toliara North observatory and nearby protected areas. Three PAs became operative during 2007--2008."
#| fig-height: 7
#| fig-width: 7

tol_pas <- dyn_study |>
  filter(str_detect(NAME, "Ranobe PK 32|Amoron|Mikea")) |>
  mutate(pa_label = case_when(
    str_detect(NAME, "Ranobe")  ~ "Ranobe PK 32 (V, 2008)",
    str_detect(NAME, "Amoron")  ~ "Amoron'i Onilahy (V, 2007)",
    str_detect(NAME, "Mikea")   ~ "Mikea NP (II, 2007)"
  ))

tol_bbox <- st_bbox(c(
  xmin = 43.2, ymin = -24.0, xmax = 44.8, ymax = -22.8
), crs = 4326)

tol_mad_c  <- st_crop(madagascar, tol_bbox)
tol_pas_c  <- tryCatch(st_crop(tol_pas, tol_bbox), error = function(e) tol_pas)
tol_fkt_c  <- tryCatch(st_crop(tol_fokontany, tol_bbox), error = function(e) tol_fokontany)
tol_fkt_c  <- tol_fkt_c |> mutate(layer = "Surveyed fokontany")

tol_pal <- c("Ranobe PK 32 (V, 2008)"      = "#EF5350",
             "Amoron'i Onilahy (V, 2007)"   = "#FF8A65",
             "Mikea NP (II, 2007)"           = "#E53935")

tm_shape(tol_mad_c) +
  tm_borders(col = "grey75", lwd = 0.3) +
  tm_fill(fill = "grey96") +
tm_shape(tol_pas_c) +
  tm_fill(fill = "pa_label",
          fill.scale = tm_scale_categorical(values = tol_pal),
          fill.legend = tm_legend(title = ""),
          fill_alpha = 0.45) +
  tm_borders(col = "pa_label",
             col.scale = tm_scale_categorical(values = tol_pal),
             col.legend = tm_legend(show = FALSE),
             lwd = 0.8) +
tm_shape(tol_fkt_c) +
  tm_fill(fill = "layer",
          fill.scale = tm_scale_categorical(
            values = c("Surveyed fokontany" = "#E3F2FD")),
          fill.legend = tm_legend(title = ""),
          fill_alpha = 0.8) +
  tm_borders(col = "grey40", lwd = 0.4) +
tm_scalebar(position = c("left", "bottom")) +
tm_title("Toliara North & Nearby PAs (2007\u20132008)") +
tm_layout(legend.position = c("right", "top"))
```

## Fénérive East and the Réserve de Tampolo {#sec-fen-site}

The Réserve de Tampolo (IUCN IV, \~7.3 km²) is a small coastal littoral forest managed by ESSA-Forêts as a research and teaching forest. A temporary protection decree was issued on 20 August 2007 (WDPA 354011). The reserve's primary function is scientific access and conservation research, not community resource exclusion. Its small footprint and institutional character suggest low exposure intensity.

Exposure year is {{< var treat_year_fen >}}. The survey has 8 pre-treatment years (1999--2006) and 8 post-treatment years (2007--2014) — the most symmetric panel among the new sites.

```{r}
#| label: fig-map-fen
#| fig-cap: "Fénérive East observatory and the Réserve de Tampolo (IUCN IV, 2007). The reserve is small (~7.3 km²) on the northeastern littoral coast."
#| fig-height: 7
#| fig-width: 7

fen_pas <- dyn_study |> filter(str_detect(NAME, "Tampolo")) |>
  mutate(pa_label = "Réserve de Tampolo (IV, 2007)")

fen_bbox <- st_bbox(c(
  xmin = 49.1, ymin = -17.7, xmax = 49.9, ymax = -16.9
), crs = 4326)

fen_mad_c  <- st_crop(madagascar, fen_bbox)
fen_fkt_c  <- tryCatch(st_crop(fen_fokontany, fen_bbox), error = function(e) fen_fokontany)
fen_pas_c  <- tryCatch(st_crop(fen_pas, fen_bbox), error = function(e) fen_pas)
fen_fkt_c  <- fen_fkt_c |> mutate(layer = "Surveyed fokontany")

tm_shape(fen_mad_c) +
  tm_borders(col = "grey75", lwd = 0.3) +
  tm_fill(fill = "grey96") +
tm_shape(fen_pas_c) +
  tm_fill(fill = "pa_label",
          fill.scale = tm_scale_categorical(
            values = c("Réserve de Tampolo (IV, 2007)" = "#CE93D8")),
          fill.legend = tm_legend(title = ""),
          fill_alpha = 0.5) +
  tm_borders(col = "pa_label",
             col.scale = tm_scale_categorical(
               values = c("Réserve de Tampolo (IV, 2007)" = "#6A1B9A")),
             col.legend = tm_legend(show = FALSE),
             lwd = 0.8) +
tm_shape(fen_fkt_c) +
  tm_fill(fill = "layer",
          fill.scale = tm_scale_categorical(
            values = c("Surveyed fokontany" = "#FFF8E1")),
          fill.legend = tm_legend(title = ""),
          fill_alpha = 0.8) +
  tm_borders(col = "grey40", lwd = 0.4) +
tm_scalebar(position = c("left", "bottom")) +
tm_title("Fénérive East & Réserve de Tampolo (2007)") +
tm_layout(legend.position = c("right", "top"))
```

## Design summary {#sec-ext-design}

```{r}
#| label: tbl-ext-design
#| tbl-cap: "Five exposed observatories: design summary"

tribble(
  ~Observatory,        ~`PA (primary)`,                     ~IUCN, ~Type,            ~`Treat yr`, ~Pre, ~Post, ~Survey,
  "Marovoay (03)",     "Ankarafantsika NP",                 "II",  "Strict",          2003L,       4L,   11L,   "1999\u20132014",
  "Alaotra (21)",      "Lac Alaotra PHP",                   "V",   "Multipurpose",    2008L,       9L,    6L,   "1999\u20132014",
  "Farafangana (15)",  "Corridor Ambositra-Vondrozo",       "V",   "Multipurpose",    2006L,       7L,    3L,   "1999\u20132008 \u26a0",
  "Toliara N. (23)",   "Amoron'i Onilahy + Mikea + Ranobe", "V/II","Mixed (V dom.)",  2007L,       8L,    8L,   "1999\u20132014",
  "F\u00e9n\u00e9rive E. (24)", "R\u00e9serve de Tampolo",   "IV",  "Low intensity",   2007L,       8L,    8L,   "1999\u20132014"
) |>
  gt() |>
  tab_header(
    title = md("**Five Exposed Observatories: Design Summary**"),
    subtitle = "\u26a0 Farafangana: only 3 post-treatment years \u2014 estimates are indicative only"
  ) |>
  tab_style(style = cell_fill(color = "#FFCDD2"),
            locations = cells_body(rows = grepl("Marovoay", Observatory))) |>
  tab_style(style = cell_fill(color = "#BBDEFB"),
            locations = cells_body(rows = grepl("Alaotra", Observatory))) |>
  tab_style(style = cell_fill(color = "#DCEDC8"),
            locations = cells_body(rows = grepl("Farafangana", Observatory))) |>
  tab_style(style = cell_fill(color = "#FFE0B2"),
            locations = cells_body(rows = grepl("Toliara", Observatory))) |>
  tab_style(style = cell_fill(color = "#E8EAF6"),
            locations = cells_body(rows = grepl("n\u00e9rive", Observatory)))
```


# Results by observatory {#sec-ext-results}

The identification strategy and estimator are identical to [Chapter 4](04_results.qmd). We estimate the gsynth ATT separately for each exposed observatory. Each model specifies $D_{it} = 1$ only for the target observatory's exposed unit(s) from its exposure year onward; all other units carry $D_{it} = 0$ and serve as additional factors for counterfactual construction. This within-model use of other exposed observatories as donors is consistent with the interactive FE framework so long as exposure timing differs across observatories [@Xu2017]. The number of latent factors $r$ is selected by leave-one-out cross-validation; inference is parametric bootstrap with 200 replications.

## Ankarafantsika (clean donors) {#sec-ext-maro}

Re-estimating the Marovoay east-bank model with the restricted donor pool yields an average post-treatment ATT of {{< var att_maro_clean >}} log points ($\approx$ {{< var att_maro_clean_pct >}}, $p$ = {{< var att_maro_clean_p >}}), consistent with the full-donor estimate in [Chapter 4](04_results.qmd). The point estimate is slightly attenuated because the smaller donor pool supports fewer latent factors, but the qualitative conclusion is unchanged: strict enforcement is associated with a persistent, statistically significant income decline.

```{r}
#| label: fig-att-maro-clean
#| fig-cap: "Ankarafantsika gsynth ATT path (clean donor pool). Exposure onset 2003."
#| fig-height: 4.5
#| fig-width: 8

gs_maro <- read_rds("data/paper14_gsynth_maro_oecd.rds")

att_maro_df <- tibble(
  rel_time = as.numeric(rownames(gs_maro$est_att)),
  att      = gs_maro$est_att[, "ATT"],
  ci_lo    = gs_maro$est_att[, "CI.lower"],
  ci_hi    = gs_maro$est_att[, "CI.upper"]
)

ggplot(att_maro_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#D32F2F", alpha = 0.15) +
  geom_line(color = "#D32F2F", linewidth = 0.8) +
  geom_point(color = "#D32F2F", size = 2.5) +
  labs(x = "Years relative to exposure (2003)",
       y = "ATT (log equiv. income)",
       title = "Ankarafantsika (strict, IUCN II)",
       subtitle = paste0(gs_maro$N_co, " donor sites; r* = ", gs_maro$r_cv)) +
  theme_minimal(base_size = 12)
```

## Lac Alaotra (clean donors) {#sec-ext-ala}

The Alaotra observatory-level ATT with clean donors remains near zero and statistically insignificant: {{< var att_ala_clean >}} log points ($p$ = {{< var att_ala_clean_p >}}). The null result is robust to the donor-pool change, reinforcing the interpretation that participatory conservation under CBNRM did not impose detectable aggregate income costs.

```{r}
#| label: fig-att-ala-clean
#| fig-cap: "Lac Alaotra gsynth ATT path (clean donor pool). Exposure onset 2008."
#| fig-height: 4.5
#| fig-width: 8

gs_ala <- read_rds("data/paper14_gsynth_ala_oecd.rds")

att_ala_df <- tibble(
  rel_time = as.numeric(rownames(gs_ala$est_att)),
  att      = gs_ala$est_att[, "ATT"],
  ci_lo    = gs_ala$est_att[, "CI.lower"],
  ci_hi    = gs_ala$est_att[, "CI.upper"]
)

ggplot(att_ala_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#1565C0", alpha = 0.15) +
  geom_line(color = "#1565C0", linewidth = 0.8) +
  geom_point(color = "#1565C0", size = 2.5) +
  labs(x = "Years relative to exposure (2008)",
       y = "ATT (log equiv. income)",
       title = "Lac Alaotra (multipurpose, IUCN V)",
       subtitle = paste0(gs_ala$N_co, " donor sites; r* = ", gs_ala$r_cv)) +
  theme_minimal(base_size = 12)
```

## Farafangana: Corridor Forestier effect {#sec-ext-fara}

::: {.callout-warning}
Farafangana has only 3 post-treatment survey years (2006--2008). The gsynth estimates are presented for completeness but should be treated with extreme caution. Pre-treatment fit is assessed over 7 years; post-treatment precision is insufficient for firm conclusions.
:::

The point estimate is {{< var att_fara >}} log points ($\approx$ {{< var att_fara_pct >}}), but with very wide confidence intervals ($p$ = {{< var att_fara_p >}}) that span both negative and positive values. With only 3 post-treatment observations, this estimate has low reliability and should not be interpreted independently.

```{r}
#| label: fig-att-fara
#| fig-cap: "Farafangana gsynth ATT path (1999--2008). Exposure onset 2006. Wide CIs reflect only 3 post-treatment years."
#| fig-height: 4.5
#| fig-width: 8

gs_fara <- read_rds("data/paper14_gsynth_fara_oecd.rds")

att_fara_df <- tibble(
  rel_time = as.numeric(rownames(gs_fara$est_att)),
  att      = gs_fara$est_att[, "ATT"],
  ci_lo    = gs_fara$est_att[, "CI.lower"],
  ci_hi    = gs_fara$est_att[, "CI.upper"]
)

ggplot(att_fara_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#2E7D32", alpha = 0.15) +
  geom_line(color = "#2E7D32", linewidth = 0.8) +
  geom_point(color = "#2E7D32", size = 2.5) +
  labs(x = "Years relative to exposure (2006)",
       y = "ATT (log equiv. income)",
       title = "Farafangana (multipurpose, IUCN V) \u2014 INDICATIVE ONLY",
       subtitle = paste0(gs_fara$N_co, " donor sites; r* = ", gs_fara$r_cv,
                         " | Only 3 post-treatment years")) +
  theme_minimal(base_size = 12)
```

## Toliara North: coastal PA cascade {#sec-ext-tol}

The Toliara North ATT estimates a composite treatment effect from the 2007--2008 PA cascade. The point estimate is {{< var att_tol >}} log points ($p$ = {{< var att_tol_p >}}), not statistically significant. The December 2008 secondary shock (Ranobe PK 32, the closest PA at <1.5 km) coincides with a visible shift in the ATT path, but it is impossible to separate the 2007 and 2008 effects without additional micro-data. The dominant conservation character of the three PAs is multipurpose (IUCN V), consistent with the near-zero aggregate income effect.

```{r}
#| label: fig-att-tol
#| fig-cap: "Toliara North gsynth ATT path. Exposure onset 2007 (Amoron'i Onilahy decree). An annotation marks the December 2008 Ranobe PK 32 secondary decree."
#| fig-height: 4.5
#| fig-width: 8

gs_tol <- read_rds("data/paper14_gsynth_tol_oecd.rds")

att_tol_df <- tibble(
  rel_time = as.numeric(rownames(gs_tol$est_att)),
  att      = gs_tol$est_att[, "ATT"],
  ci_lo    = gs_tol$est_att[, "CI.lower"],
  ci_hi    = gs_tol$est_att[, "CI.upper"]
)

ggplot(att_tol_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#E65100", alpha = 0.15) +
  geom_line(color = "#E65100", linewidth = 0.8) +
  geom_point(color = "#E65100", size = 2.5) +
  geom_vline(xintercept = 1, linetype = "dotted",
             color = "#E65100", linewidth = 0.6) +
  annotate("text", x = 1.2, y = max(att_tol_df$ci_hi, na.rm = TRUE) * 0.85,
           label = "Ranobe PK 32\n(Dec 2008)", color = "#E65100",
           hjust = 0, size = 2.8) +
  labs(x = "Years relative to exposure (2007)",
       y = "ATT (log equiv. income)",
       title = "Toliara North (mixed IUCN V/II)",
       subtitle = paste0(gs_tol$N_co, " donor sites; r* = ", gs_tol$r_cv)) +
  theme_minimal(base_size = 12)
```

## Fénérive East: low-intensity reserve {#sec-ext-fen}

The Fénérive East ATT shows a small positive but nonsignificant effect ({{< var att_fen >}} log points, $p$ = {{< var att_fen_p >}}) --- perhaps the most interpretively clear result among the three new sites. A 7.3 km² research forest cannot plausibly impose population-level income costs on a multi-commune observatory. The 8/8 pre-/post-treatment symmetry provides the most credible pre-trend assessment of the new sites.

```{r}
#| label: fig-att-fen
#| fig-cap: "Fénérive East gsynth ATT path. Exposure onset 2007 (Réserve de Tampolo decree)."
#| fig-height: 4.5
#| fig-width: 8

gs_fen <- read_rds("data/paper14_gsynth_fen_oecd.rds")

att_fen_df <- tibble(
  rel_time = as.numeric(rownames(gs_fen$est_att)),
  att      = gs_fen$est_att[, "ATT"],
  ci_lo    = gs_fen$est_att[, "CI.lower"],
  ci_hi    = gs_fen$est_att[, "CI.upper"]
)

ggplot(att_fen_df, aes(rel_time, att)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#6A1B9A", alpha = 0.15) +
  geom_line(color = "#6A1B9A", linewidth = 0.8) +
  geom_point(color = "#6A1B9A", size = 2.5) +
  labs(x = "Years relative to exposure (2007)",
       y = "ATT (log equiv. income)",
       title = "F\u00e9n\u00e9rive East (low-intensity, IUCN IV)",
       subtitle = paste0(gs_fen$N_co, " donor sites; r* = ", gs_fen$r_cv)) +
  theme_minimal(base_size = 12)
```


# Five-case synthesis {#sec-synthesis}

## Comparative results

```{r}
#| label: tbl-five-att
#| tbl-cap: "Five-observatory comparative results. Farafangana estimates (\u26a0) are unreliable due to only 3 post-treatment years."

get_ci <- function(gs, side) if (!is.null(gs$est_avg)) gs$est_avg[1, side] else NA_real_
get_se <- function(gs) if (!is.null(gs$est_avg)) gs$est_avg[1, "S.E."] else NA_real_
get_pv <- function(gs) if (!is.null(gs$est_avg)) gs$est_avg[1, "p.value"] else NA_real_

att_vals <- c(gs_maro$att_avg, gs_ala$att_avg, gs_fara$att_avg,
              gs_tol$att_avg, gs_fen$att_avg)
se_vals  <- c(get_se(gs_maro), get_se(gs_ala), get_se(gs_fara),
              get_se(gs_tol), get_se(gs_fen))
ci_lo_vals <- c(get_ci(gs_maro, "CI.lower"), get_ci(gs_ala, "CI.lower"),
                get_ci(gs_fara, "CI.lower"), get_ci(gs_tol, "CI.lower"),
                get_ci(gs_fen, "CI.lower"))
ci_hi_vals <- c(get_ci(gs_maro, "CI.upper"), get_ci(gs_ala, "CI.upper"),
                get_ci(gs_fara, "CI.upper"), get_ci(gs_tol, "CI.upper"),
                get_ci(gs_fen, "CI.upper"))
pval_vals  <- c(get_pv(gs_maro), get_pv(gs_ala), get_pv(gs_fara),
                get_pv(gs_tol), get_pv(gs_fen))

tibble(
  Observatory     = c("Marovoay (03)", "Alaotra (21)", "Farafangana (15) \u26a0",
                      "Toliara North (23)", "F\u00e9n\u00e9rive East (24)"),
  `PA type`       = c("Strict (II)", "Multipurpose (V)", "Multipurpose (V)",
                      "Mixed (V/II)", "Low-intensity (IV)"),
  `Treat yr`      = c(2003, 2008, 2006, 2007, 2007),
  `ATT (log pts)` = round(att_vals, 3),
  SE              = round(se_vals, 3),
  `95% CI`        = paste0("[", round(ci_lo_vals, 3), ", ", round(ci_hi_vals, 3), "]"),
  `ATT (%)`       = paste0(round((exp(att_vals) - 1) * 100, 1), "%"),
  `p-value`       = round(pval_vals, 3),
  `r*`            = c(gs_maro$r_cv, gs_ala$r_cv, gs_fara$r_cv, gs_tol$r_cv, gs_fen$r_cv),
  `N donors`      = c(gs_maro$N_co, gs_ala$N_co, gs_fara$N_co, gs_tol$N_co, gs_fen$N_co),
  `Post yrs`      = c(11L, 6L, 3L, 8L, 8L)
) |>
  gt() |>
  tab_header(
    title = md("**Five-Observatory Comparative Results**"),
    subtitle = "\u26a0 Farafangana: only 3 post-treatment years \u2014 treat as indicative"
  ) |>
  tab_style(style = cell_fill(color = "#FFCDD2"),
            locations = cells_body(rows = grepl("Marovoay", Observatory))) |>
  tab_style(style = cell_fill(color = "#BBDEFB"),
            locations = cells_body(rows = grepl("Alaotra", Observatory))) |>
  tab_style(style = cell_fill(color = "#DCEDC8"),
            locations = cells_body(rows = grepl("Farafangana", Observatory))) |>
  tab_style(style = cell_fill(color = "#FFE0B2"),
            locations = cells_body(rows = grepl("Toliara", Observatory))) |>
  tab_style(style = cell_fill(color = "#E8EAF6"),
            locations = cells_body(rows = grepl("n\u00e9rive", Observatory)))
```

```{r}
#| label: fig-att-all-overlay
#| fig-cap: "ATT paths for all five exposed observatories on the same relative-time axis. CI bands suppressed for legibility."
#| fig-height: 5.5
#| fig-width: 11

all_att <- bind_rows(
  att_maro_df |> mutate(obs = "Marovoay (strict, II)",        color_grp = "Marovoay (strict)"),
  att_ala_df  |> mutate(obs = "Alaotra (multipurpose, V)",    color_grp = "Alaotra (multipurpose)"),
  att_fara_df |> mutate(obs = "Farafangana (multipurpose, V)",color_grp = "Farafangana (multipurpose)"),
  att_tol_df  |> mutate(obs = "Toliara North (mixed, V/II)",  color_grp = "Toliara North (mixed)"),
  att_fen_df  |> mutate(obs = "F\u00e9n\u00e9rive East (low int., IV)",  color_grp = "F\u00e9n\u00e9rive East (low int.)")
)

ggplot(all_att, aes(rel_time, att, color = color_grp, linetype = color_grp)) +
  geom_hline(yintercept = 0, color = "grey60", linewidth = 0.4) +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey50", linewidth = 0.4) +
  annotate("rect", xmin = -0.5, xmax = Inf, ymin = -Inf, ymax = Inf,
           fill = "grey90", alpha = 0.18) +
  geom_line(linewidth = 0.85) +
  geom_point(size = 1.8) +
  scale_color_manual(
    values = c("Marovoay (strict)" = "#D32F2F",
               "Alaotra (multipurpose)" = "#1565C0",
               "Farafangana (multipurpose)" = "#2E7D32",
               "Toliara North (mixed)" = "#E65100",
               "F\u00e9n\u00e9rive East (low int.)" = "#6A1B9A"),
    name = NULL
  ) +
  scale_linetype_manual(
    values = c("Marovoay (strict)" = "solid",
               "Alaotra (multipurpose)" = "solid",
               "Farafangana (multipurpose)" = "dashed",
               "Toliara North (mixed)" = "solid",
               "F\u00e9n\u00e9rive East (low int.)" = "solid"),
    name = NULL
  ) +
  labs(x = "Years relative to exposure onset",
       y = "ATT (log equiv. income)",
       title = "ATT Paths: Five Conservation Governance Models",
       subtitle = "Dashed = Farafangana (indicative only, 3 post-treatment years)") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 8.5))
```

```{r}
#| label: fig-forest-plot
#| fig-cap: "Forest plot of average post-treatment ATTs with 95% parametric bootstrap CIs. Farafangana shown at reduced opacity."
#| fig-height: 4
#| fig-width: 8

forest_df <- tibble(
  obs   = c("Marovoay\n(strict, II, 2003)",
            "Alaotra\n(multipurpose, V, 2008)",
            "Farafangana \u26a0\n(multipurpose, V, 2006)",
            "Toliara North\n(mixed, V/II, 2007)",
            "F\u00e9n\u00e9rive East\n(low int., IV, 2007)"),
  att   = att_vals,
  ci_lo = ci_lo_vals,
  ci_hi = ci_hi_vals,
  col   = c("#D32F2F", "#1565C0", "#2E7D32", "#E65100", "#6A1B9A"),
  indicative = c(FALSE, FALSE, TRUE, FALSE, FALSE)
) |>
  arrange(att) |>
  mutate(obs = factor(obs, levels = obs))

ggplot(forest_df, aes(y = obs, x = att, xmin = ci_lo, xmax = ci_hi,
                      color = col, alpha = indicative)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_linerange(linewidth = 1.2) +
  geom_point(size = 3.5) +
  scale_color_identity() +
  scale_alpha_manual(values = c("FALSE" = 1, "TRUE" = 0.45), guide = "none") +
  labs(x = "ATT (log equiv. income)", y = NULL,
       title = "Forest Plot: Five-Observatory ATT Estimates",
       subtitle = "\u26a0 = Farafangana (3 post-treatment years, indicative only)") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank())
```

## Three patterns {#sec-patterns}

The five-case comparison suggests three broad patterns, each of which warrants careful interpretation given the small number of exposed units per model.

The first pattern is a large negative income gap under strict enforcement. Ankarafantsika (Marovoay) shows the largest and most persistent negative gap relative to its synthetic counterfactual, statistically significant and apparent through the 2025 resurvey. This is consistent with direct resource exclusion from a large, actively patrolled park. However, as discussed in @sec-maro-rcs and @sec-counterfactual, the within-Marovoay comparison between east-bank and west-bank villages finds no equivalent income difference, which raises the possibility that the gsynth is capturing regional economic decline — the long-run deterioration of the Marovoay irrigated rice economy — rather than a park-specific effect. The Marovoay result should therefore be treated as suggestive rather than conclusive evidence of a conservation-driven income loss.

The second pattern is near-zero estimated effects under participatory or multipurpose governance. Lac Alaotra, Toliara North, and tentatively Farafangana all show point estimates clustered near zero with no statistically significant departure. The community-based management framework at Alaotra, and the lighter regulatory touch at the other sites, appears consistent with no detectable aggregate income costs at the survey-year frequency — though low statistical power (one exposed unit per model) means that moderate effects cannot be ruled out.

The third pattern is that a small, research-oriented reserve with minimal enforcement leaves estimated income unchanged. Fénérive East shows a near-zero ATT with a moderately narrow confidence interval, which is the result one would expect from a 7.3 km² scientific forest that does not constrain community livelihoods at scale.


## Caveats

Several limitations temper interpretation of the extended comparison. With only three post-treatment years, the Farafangana estimate has low reliability and should not be compared numerically to the others. At Toliara North, the rapid succession of three protected areas in 2007–2008 means the estimated ATT captures a composite of overlapping events rather than a clean single intervention. Using only five donor observatories rather than the nine available in Chapter 4 reduces counterfactual precision, which is why ATT estimates for Marovoay and Alaotra differ slightly between the two chapters. Identification for the three new sites rests on decree dates and spatial proximity without the behavioural validation available for Marovoay and Alaotra, since no 2025 resurvey covered these observatories. Finally, each observatory-level model has exactly one exposed unit, which imposes a minimum permutation p-value of roughly $1/N_{\text{donors}}$; effects below around 20% cannot be reliably distinguished from chance.
